---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# CelliVerse Project

<!-- badges: start -->

[![CelliVerse R package](https://img.shields.io/badge/CelliVerse-R%20package-276DC3)](https://github.com/asalavaty/celliverse)
[![CRAN](https://www.r-pkg.org/badges/version/celliverse?color=blue)](https://cran.r-project.org/package=celliverse)
[![Zenodo](https://img.shields.io/badge/Zenodo-10.5281%2Fzenodo.20550512-1682D4)](https://doi.org/10.5281/zenodo.20550512)
[![Hugging Face Space](https://img.shields.io/badge/🤗%20Hugging%20Face-Space-yellow)](https://huggingface.co/spaces/asalavaty/celliverse)

<!-- badges: end -->

## Overview

This repository accompanies the **ClustoCell** study and contains resources required to reproduce the analyses and figures presented in the manuscript.

It provides:

- the prepared **CelliVerse MarkerDB** resources used for cell-type annotation;
- human and mouse positive and negative marker databases in both dense and sparse formats;
- analysis scripts used to generate the main, extended data, and supplementary figures presented in the study; and
- links to the `celliverse` R package and the publicly available datasets used in the analyses.

The associated [`celliverse`](https://github.com/asalavaty/celliverse) R package provides the implementations of **ClustoCell**, marker identification and evaluation methods, cell-type annotation workflows, visualization functions, and related tools used throughout the study.

---

## Repository structure

```text
CelliVerse-Project/
│
├── Marker Database/
│   ├── human_pos_marker_db.rds
│   ├── human_pos_marker_sparse_db.rds
│   ├── human_neg_marker_db.rds
│   ├── human_neg_marker_sparse_db.rds
│   ├── mouse_pos_marker_db.rds
│   ├── mouse_pos_marker_sparse_db.rds
│   ├── mouse_neg_marker_db.rds
│   └── mouse_neg_marker_sparse_db.rds
│
├── Scripts/
│   ├── Figures/
│   │   ├── Figure 2.Rmd
│   │   ├── Figure 3.Rmd
│   │   ├── Figure 4.Rmd
│   │   ├── Figure 5.Rmd
│   │   └── Figure 6.Rmd
│   │
│   ├── Extended Data Figures/
│   │   ├── Extended Data Figure 1.Rmd
│   │   ├── Extended Data Figure 2.Rmd
│   │   ├── Extended Data Figure 3.Rmd
│   │   └── Extended Data Figure 4.Rmd
│   │
│   └── Supplementary Figures/
│       ├── Supplementary Figure 1.Rmd
│       ├── ...
│       ├── Supplementary Figure 22.Rmd
│       ├── Supplementary Figure 23-24.Rmd
│       └── Supplementary Figure 25.Rmd
│
├── LICENSE
└── README.md
```

---

## CelliVerse MarkerDB

The **CelliVerse MarkerDB** is a harmonized cell-type marker resource used by `celliverse` for marker-based cell-type annotation.

This repository provides the prepared marker databases used in the study separately for:

| Species | Marker type | Dense | Sparse |
|:--|:--|:--|:--|
| Human | Positive | `human_pos_marker_db.rds` | `human_pos_marker_sparse_db.rds` |
| Human | Negative | `human_neg_marker_db.rds` | `human_neg_marker_sparse_db.rds` |
| Mouse | Positive | `mouse_pos_marker_db.rds` | `mouse_pos_marker_sparse_db.rds` |
| Mouse | Negative | `mouse_neg_marker_db.rds` | `mouse_neg_marker_sparse_db.rds` |

The same MarkerDB is also distributed as the `markerDB` data object within the [`celliverse`](https://github.com/asalavaty/celliverse) R package.

All publicly available marker resources and the HGNC gene nomenclature database used to construct the CelliVerse MarkerDB are reported in **Supplementary Data 1** of the manuscript.

### Loading a database

For example:

```{r, eval=FALSE}
human_pos_marker_db <- readRDS(
  "Marker Database/human_pos_marker_db.rds"
)

human_pos_marker_sparse_db <- readRDS(
  "Marker Database/human_pos_marker_sparse_db.rds"
)
```

---

## Reproducing manuscript analyses

The `Scripts` directory contains the R Markdown workflows used to generate the figures presented in the manuscript.

Scripts are organized according to their corresponding manuscript figure.

### Main figures

```text
Scripts/Figures/
```

contains the analysis workflows for **Figures 2–6**.

### Extended Data figures

```text
Scripts/Extended Data Figures/
```

contains the workflows for **Extended Data Figures 1–4**.

### Supplementary figures

```text
Scripts/Supplementary Figures/
```

contains the workflows for **Supplementary Figures 1–25**.

Each `.Rmd` file contains the analysis and visualization code corresponding to the indicated figure.

> **Note:** Some analyses are computationally intensive and may require substantial memory and runtime depending on the dataset and hardware. Scripts are provided to enable transparent inspection and reproduction of the analyses reported in the manuscript.

---

## Data availability

All publicly available bulk and single-cell RNA-seq datasets analyzed in the study are listed in **Supplementary Data 4** of the manuscript.

For reproducibility and long-term accessibility, the datasets used in the analyses are additionally archived on Zenodo:

**DOI:** [10.5281/zenodo.20550512](https://doi.org/10.5281/zenodo.20550512)

The datasets are not duplicated in this GitHub repository because of their size. They can be obtained from their original public repositories or from the Zenodo archive above.

---

## Installing CelliVerse

The analyses in this repository use functions implemented in the `celliverse` R package.

The current CRAN release can be installed using:

```{r, eval=FALSE}
install.packages("celliverse")
```

The development version can be installed directly from GitHub:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github(
  "asalavaty/celliverse",
  build_vignettes = TRUE
)
```

For complete package documentation, examples, and tutorials, see:

- [CelliVerse GitHub repository](https://github.com/asalavaty/celliverse)
- [CelliVerse CRAN page](https://cran.r-project.org/package=celliverse)
- [CelliVerse vignette](https://cran.r-project.org/web/packages/celliverse/vignettes/Introduction-to-CelliVerse.html)

---

## ClustoCell

**ClustoCell** is implemented within the `celliverse` R package and provides a data-driven framework for single-cell clustering and marker identification.

The package integrates ClustoCell with complementary functionality for:

- cluster and sub-cluster marker identification;
- positive and negative marker discovery;
- marker purity assessment;
- dataset-level feature selection;
- cell-type annotation using the CelliVerse MarkerDB;
- LLM-assisted cell-type annotation;
- generation of portable LLM annotation prompts; and
- visualization and downstream integration of clustering and annotation results.

A lightweight browser-based demonstration of ClustoCell outputs is also available through the [CelliVerse Hugging Face Space](https://huggingface.co/spaces/asalavaty/celliverse).

---

## CelliVerse Agent

The `celliverse` package also includes an **LLM-powered CelliVerse Agent** that provides a natural-language interface to CelliVerse functionality.

The Agent enables users to interact with their single-cell datasets and request CelliVerse analyses, including application of ClustoCell, through conversational instructions. This provides a convenient route for researchers who wish to use CelliVerse without directly writing R code.

The Agent supports both cloud-based language models and local models through providers such as Ollama and LM Studio. Full setup and usage instructions are provided in the [`celliverse` vignette](https://cran.r-project.org/web/packages/celliverse/vignettes/Introduction-to-CelliVerse.html).

---

## Interactive demonstration

A lightweight browser-based demonstration of selected CelliVerse functionality is available through Hugging Face Spaces:

[![Open CelliVerse Demo](https://img.shields.io/badge/🤗-Open%20CelliVerse%20Demo-FFD21E?style=for-the-badge)](https://huggingface.co/spaces/asalavaty/celliverse)

The demo provides an interactive ClustoCell PBMC3K example using precomputed outputs, allowing users to explore UMAP visualizations, clusters, sub-clusters, and marker results without installing the package locally.

---

## Reproducibility

To reproduce a particular manuscript result:

1. Install the required version of `celliverse` and any additional R packages used by the corresponding script.
2. Obtain the required dataset from its original source or the [Zenodo archive](https://doi.org/10.5281/zenodo.20550512).
3. Open the appropriate `.Rmd` file under `Scripts/`.
4. Update local data paths where required.
5. Run the analysis sections of the script to reproduce the corresponding results and figure.

Because individual scripts may use different public datasets and analysis-specific dependencies, the required inputs and processing steps should be followed directly from the corresponding figure workflow.

---

## Related resources

| Resource | Location |
|:--|:--|
| CelliVerse R package | [github.com/asalavaty/celliverse](https://github.com/asalavaty/celliverse) |
| CRAN package | [cran.r-project.org/package=celliverse](https://cran.r-project.org/package=celliverse) |
| Manuscript analysis repository | [github.com/asalavaty/CelliVerse-Project](https://github.com/asalavaty/CelliVerse-Project) |
| Archived study datasets | [Zenodo: 10.5281/zenodo.20550512](https://doi.org/10.5281/zenodo.20550512) |
| Interactive ClustoCell demo | [Hugging Face Space](https://huggingface.co/spaces/asalavaty/celliverse) |
| Package vignette | [CelliVerse vignette](https://cran.r-project.org/web/packages/celliverse/vignettes/Introduction-to-CelliVerse.html) |

---

## Citation

If you use **ClustoCell**, **CelliVerse MarkerDB**, the `celliverse` package, or resources from this repository, please cite the associated CelliVerse/ClustoCell publication.

The full citation will be added upon publication.

Citation information for the R package can also be accessed directly from R:

```{r citation, eval=FALSE}
citation("celliverse")
```

---

## Author

The CelliVerse project and `celliverse` R package were developed by [Adrian Salavaty](https://asalavaty.com/).

### Advisor

- Ramyar Molania

---

## License

Please see the [`LICENSE`](LICENSE) file in this repository for licensing information.

---

## Issues and questions

For issues related to the **CelliVerse R package**, please use the [`celliverse` issue tracker](https://github.com/asalavaty/celliverse/issues).

For questions specifically related to the manuscript reproduction scripts or resources contained in this repository, please open an issue in this repository.
