---
title: "Evaluating ClustoCell's Performance in finding novel biology"
author: "Adrian Salavaty"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(pROC)
library(umap)
library(cluster)
library(ggplot2)
library(dplyr)
library(splatter)
library(Seurat)
library(tidyverse)
library(scplotter)
library(scran)
library(scater)
library(Matrix)
library(SeuratDisk)
library(hdf5r)
library(patchwork)
library(paletteer)
library(hrbrthemes)
library(scales)

```

# Introduction

Script for evaluation of ClustoCell's Performance in finding novel biology.

# Function Defintion

We first define some helper functions.

```{r}

# Personalized Theme function

styled_theme <- function(
  style = c("academic", "minimal", "modern", "dark", "classic", "void"),
  border_mode = c("default", "all", "none"),
  title_align = c("center", "left", "right"),
  legends_to_drop = NULL,
  axes_title_drop = FALSE,
  legend_position = c("default", "top", "bottom", "left", "right", "inside"),
  axis_title_face = c("regular", "bold"),
  x_axis_text_angle = 45,
  x_axis_text_y = -0.3,
  remove_margin = c("none", "bottom", "all", "top", "left", "right")
) {
  style <- match.arg(style)
  border_mode <- match.arg(border_mode)
  title_align <- match.arg(title_align)
  legend_position <- match.arg(legend_position)
  axis_title_face <- match.arg(axis_title_face)
  remove_margin <- match.arg(remove_margin)

  base_font <- "Helvetica"
  align_map <- c(left = 0, center = 0.5, right = 1)
  hjust_val <- align_map[title_align]
  thin_size <- 0.3  

  # Decide axis title face
  axis_face_val <- ifelse(axis_title_face == "bold", "bold", "plain")

  # Base theme selection
  if (style == "void") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        axis.title.x = element_blank(),  
        axis.title.y = element_blank(),  
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "academic") {
    theme_out <- theme_minimal(base_size = 14, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = hjust_val),
        plot.subtitle = element_text(size = 13, hjust = hjust_val, color = "grey30"),
        plot.caption = element_text(size = 10, hjust = 1, color = "grey40"),
        axis.title = element_text(face = axis_face_val, size = 13),
        axis.text = element_text(size = 12, color = "black"),
        panel.grid.major = element_line(color = "grey80", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "minimal") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", size = thin_size),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.text = element_text(color = "black"),
        axis.title = element_text(face = axis_face_val),
        legend.position = "none",
        legend.title = element_text(face = "italic")
      )
  } else if (style == "modern") {
    theme_out <- theme_light(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.border = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "dark") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        plot.background = element_rect(fill = "black", color = NA),
        panel.background = element_rect(fill = "black", color = NA),
        panel.grid.major = element_line(color = "grey30", size = thin_size),
        panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 15, color = "white", hjust = hjust_val),
        plot.subtitle = element_text(color = "grey80", hjust = hjust_val),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white", face = axis_face_val),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white", face = "italic"),
        axis.line = element_line(color = "white", size = thin_size)
      )
  } else if (style == "classic") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  }

  # Apply border mode
  if (border_mode == "all") {
    theme_out <- theme_out + theme(panel.border = element_rect(color = "black", fill = NA, size = thin_size))
  } else if (border_mode == "none") {
    theme_out <- theme_out + theme(panel.border = element_blank(), axis.line = element_blank())
  }

  # Drop axes titles if requested
  if (axes_title_drop) {
    theme_out <- theme_out + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
  }

  # Override x-axis text angle and vertical position
  theme_out <- theme_out + theme(
    axis.text.x = element_text(
      angle = x_axis_text_angle,
      vjust = 1,
      hjust = 1,
      margin = margin(t = x_axis_text_y)
    )
  )

  # Handle margin removal
  if (remove_margin == "all") {
    theme_out <- theme_out + theme(plot.margin = margin(0, 0, 0, 0))
  } else if (remove_margin == "top") {
    theme_out <- theme_out + theme(plot.margin = margin(t = 0))
  } else if (remove_margin == "bottom") {
    theme_out <- theme_out + theme(plot.margin = margin(b = 0))
  } else if (remove_margin == "left") {
    theme_out <- theme_out + theme(plot.margin = margin(l = 0))
  } else if (remove_margin == "right") {
    theme_out <- theme_out + theme(plot.margin = margin(r = 0))
  }
  # if "none", do nothing

  # Legend positioning override
  if (legend_position != "default") {
    if (legend_position == "inside") {
      theme_out <- theme_out + theme(
        legend.position = c(0.98, 0.98),
        legend.justification = c("right", "top"),
        legend.direction = "vertical"
      )
    } else {
      horiz <- legend_position %in% c("top", "bottom")
      theme_out <- theme_out + theme(
        legend.position = legend_position,
        legend.direction = if (horiz) "horizontal" else "vertical"
      )
    }
  }

  # Return theme + guides if needed
  if (!is.null(legends_to_drop)) {
    drop_list <- lapply(legends_to_drop, function(x) "none")
    names(drop_list) <- legends_to_drop
    return(list(theme_out, guides(.list = drop_list)))
  } else {
    return(theme_out)
  }
}

#____________

# Label plot
label_plt <- function(label = "Label", size = 1) {
  ggplot(data.frame(x = 1, y = 1), aes(x = x, y = y)) +
    # Draw a single vertical bar
    geom_col(width = 0.6, fill = "lightgrey", color = "black") +
    # Add vertical text (angle=90)
    annotate("text", x = 1, y = 0.5, label = label, size = size,
             angle = 90, size = 6, vjust = 0.5, hjust = 0.5) +
    # Remove all extra space
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme_void(base_size = 14) +             # no axes, ticks, labels
    theme(
        plot.margin = margin(0, 0, 0, 0),      # remove outside margins
        panel.spacing = unit(0, "pt"),         # remove panel padding
        plot.background = element_blank(),     # no background
        panel.background = element_blank()
    )
}

#__________________________________________________________________

# Box plot with stats

make_facet_boxplot_with_stats <- function(
    df,
    x = "group",
    y = "value",
    fill = NULL,
    palette = NULL,
    facet.by = NULL,
    scales = "free_y",
    facet.nrow = NULL,       
    facet.ncol = NULL,      
    alpha = 0.8,
    xlab = NULL,
    ylab = NULL,
    subtitle = NULL,
    tag = NULL,
    jitter.width = 0.15,
    jitter.size = 0.35,
    jitter.alpha = 0.6,
    method = "wilcox.test",
    p.adjust.method = "none",
    hide.ns = FALSE,
    base_size = 10
) {

  library(dplyr)
  library(ggpubr)
  library(rstatix)
  library(ggplot2)

  # ----------------------------
  # 1. Statistical test per facet
  # ----------------------------
  stat.test <- df %>%
    group_by(!!sym(facet.by)) %>%
    rstatix::wilcox_test(as.formula(paste(y, "~", x))) %>%
    adjust_pvalue(method = p.adjust.method) %>%
    add_significance()

  # Add xy positions (critical for facetted layout)
  stat.test <- stat.test %>% add_xy_position(x = x)

  # ----------------------------
  # 2. Build ggpubr boxplot
  # ----------------------------
  p <- ggpubr::ggboxplot(
    df,
    x = x,
    y = y,
    fill = fill,
    alpha = alpha,
    palette = palette,
    facet.by = facet.by,
    scales = scales,
    nrow = facet.nrow,  
    ncol = facet.ncol,   
    outlier.shape = NA
  ) +
    geom_jitter(
      width = jitter.width,
      size = jitter.size,
      alpha = jitter.alpha
    ) +
    labs(
      x = xlab,
      y = ylab,
      subtitle = subtitle,
      tag = tag
    ) +
    theme_classic(base_size = base_size) +
    theme(
      plot.subtitle = element_text(size = 7),
      text = element_text(size = 6),
      strip.text = element_text(size = 5.5),
      plot.tag = element_text(size = 7, face = "bold"),
      axis.title.y = element_text(size = 7),
      plot.margin = margin(0, 5, 0, 5),
      legend.position = "none"
    )

  # ----------------------------------
  # 3. Add p-value stars to the plot
  # ----------------------------------
  p <- p + stat_pvalue_manual(
    stat.test,
    label = "p.adj.signif",
    hide.ns = hide.ns,
    size = 2.2,
    tip.length = 0.01
  )

  return(p)
}


```


```{r}

# Concordance = 100*Sum(rowMax)/N

library(paletteer)

paste(as.vector(col2rgb("#0000CCFF")), collapse = " ")

# https://pmassicotte.github.io/paletteer_gallery/

paletteer::paletteer_d("MetBrewer::Austria", 7)
#A40000FF #16317DFF #007E2FFF #FFCD12FF #B86092FF #721B3EFF #00B7A7FF 

plotthis::palette_list %>% names()

# Best pals from ggsci: pal_npg (10), pal_d3 (10), pal_ucscgb (26), pal_igv (51)

# "#FF0000FF" "#FF9900FF" "#FFCC00FF" "#00FF00FF" "#6699FFFF" "#CC33FFFF" "#99991EFF" "#999999FF" "#FF00CCFF" "#CC0000FF" "#FFCCCCFF" "#FFFF00FF" "#CCFF00FF" "#358000FF" "#0000CCFF" "#99CCFFFF" "#00FFFFFF"
# "#CCFFFFFF" "#9900CCFF" "#CC99FFFF" "#996600FF" "#666600FF" "#666666FF" "#CCCCCCFF" "#79CC3DFF" "#CCCC99FF"

# "#1F77B4FF" "#FF7F0EFF" "#2CA02CFF" "#D62728FF" "#9467BDFF" "#8C564BFF" "#E377C2FF" "#7F7F7FFF" "#BCBD22FF" "#17BECFFF"

# NPG colors:
# "#E64B35FF" "#4DBBD5FF" "#00A087FF" "#3C5488FF" "#F39B7FFF" "#8491B4FF" "#91D1C2FF" "#DC0000FF" "#7E6148FF" "#B09C85FF"

# "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF" "#924822FF" "#837B8DFF" "#C75127FF" "#D58F5CFF" "#7A65A5FF" "#E4AF69FF" "#3B1B53FF"
# "#CDDEB7FF" "#612A79FF" "#AE1F63FF" "#E7C76FFF" "#5A655EFF" "#CC9900FF" "#99CC00FF" "#A9A9A9FF" "#CC9900FF" "#99CC00FF" "#33CC00FF" "#00CC33FF" "#00CC99FF" "#0099CCFF" "#0A47FFFF" "#4775FFFF" "#FFC20AFF"
# "#FFD147FF" "#990033FF" "#991A00FF" "#996600FF" "#809900FF" "#339900FF" "#00991AFF" "#009966FF" "#008099FF" "#003399FF" "#1A0099FF" "#660099FF" "#990080FF" "#D60047FF" "#FF1463FF" "#00D68FFF"

 # "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF"

# Define the color palettes

npg_pal <- ggsci::pal_npg()(10)
npg_reverse_pal <- rev(ggsci::pal_npg("nrc")(10))

```

A function for converting count df to sparse matrix

```{r}

library(Rcpp)
library(RcppArmadillo)

Rcpp::sourceCpp(code = '
#define ARMA_64BIT_WORD 1
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

// [[Rcpp::export]]
Rcpp::S4 denseToSparse(Rcpp::DataFrame& df) {  // Pass by reference to allow in-place modification
  int nrows = df.nrows();
  int ncols = df.size();

  arma::sp_mat sparse(nrows, ncols);

  for (int col = 0; col < ncols; ++col) {
    arma::vec this_col = Rcpp::as<arma::vec>(df[col]);  // Wrap without copying
    sparse.col(col) = this_col;  // Assigns only non-zeros to sparse
    df[col] = R_NilValue;  // Free the column memory progressively
  }

  return Rcpp::wrap(sparse);  // Returns as dgCMatrix S4 object
}
')

```

Concordance = 100*Sum(rowMax)/N

# Breast Cancer Single-Cell Map (bcscm)

## Data preparation

The data was downloaded from https://lambrechtslab.sites.vib.be/en/single-cell
https://www.nature.com/articles/s41591-021-01323-8

### T-/NK cells

```{r}

bcscm_tcell_counts <- read_rds("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/1864-counts_tcell_cohort1.rds")
bcscm_tcell_meta <- fread("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/1870-BIOKEY_metaData_tcells_cohort1_web.csv")

ncol(bcscm_tcell_counts) == nrow(bcscm_tcell_meta) # TRUE
all(colnames(bcscm_tcell_counts) == bcscm_tcell_meta$Cell) # FALSE
all(colnames(bcscm_tcell_counts) %in% bcscm_tcell_meta$Cell) # TRUE
bcscm_tcell_meta <- bcscm_tcell_meta[match(colnames(bcscm_tcell_counts), bcscm_tcell_meta$Cell),]
all(colnames(bcscm_tcell_counts) == bcscm_tcell_meta$Cell) # TRUE

bcscm_tcell_so <- CreateSeuratObject(bcscm_tcell_counts, meta.data = bcscm_tcell_meta, min.cells = 3, min.features = 200)
qs2::qs_save(object = bcscm_tcell_so, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")

bcscm_tcell_so$cellSubType <- gsub("Proliferating", "Prolif", bcscm_tcell_so$cellSubType)

bcscm_tcell_so <- NormalizeData(bcscm_tcell_so, normalization.method = "LogNormalize", scale.factor = 10000)
bcscm_tcell_so <- FindVariableFeatures(bcscm_tcell_so, selection.method = "vst", nfeatures = 2000)

qs2::qs_save(object = bcscm_tcell_so, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")

bcscm_tcell_ClustoCell <- clustoCell(data = bcscm_tcell_so, sketch = T, sketch_ncells = 25000, refine_transferred_subClusters = TRUE, 
                                     high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)

qs2::qs_save(object = bcscm_tcell_ClustoCell, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_ClustoCell.qs2")

bcscm_tcell_so <- addClustoData(obj = bcscm_tcell_so, clustoCell = bcscm_tcell_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

# Finding the dataset-level markers and using them for dim reduction
bcscm_tcell_markers <- getDatasetMarkers(obj = bcscm_tcell_ClustoCell, 
                                         clusters = T, sub_clusters = T,
                                         positive_markers = T, negative_markers = T, 
                                         medium_markers = F, 
                                         thresh_mode = "n", pos_thresh = 200, neg_thresh = 50)

bcscm_tcell_so <- ScaleData(bcscm_tcell_so, features = bcscm_tcell_markers$combined_markers)
bcscm_tcell_so <- RunPCA(bcscm_tcell_so, features = bcscm_tcell_markers$combined_markers)
ElbowPlot(bcscm_tcell_so, ndims = 50)
# bcscm_tcell_so <- FindNeighbors(bcscm_tcell_so, dims = 1:16)
# bcscm_tcell_so <- FindClusters(bcscm_tcell_so, resolution = 0.5)
bcscm_tcell_so <- RunUMAP(bcscm_tcell_so, dims = 1:16)

qs2::qs_save(object = bcscm_tcell_so, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")

bcscm_tcell_so <- qs2::qs_read("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")
bcscm_tcell_ClustoCell <- qs2::qs_read("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_ClustoCell.qs2")

```

#### Inspecting the data

Checking the metadata and confusion tables

```{r}

head(bcscm_tcell_so@meta.data %>% dplyr::select(-Cell))

# ClustoCell clusters and subclusters
table(bcscm_tcell_so$ClustoCell)
table(bcscm_tcell_so$SubClustoCell)

# Author annotations
table(bcscm_tcell_so@meta.data$cellSubType)
table(bcscm_tcell_so@meta.data$expansion)
table(bcscm_tcell_so@meta.data$timepoint)
table(bcscm_tcell_so@meta.data$BC_type)

# Confusion of author annotations with ClustoCell
table(bcscm_tcell_so$cellSubType, bcscm_tcell_so$ClustoCell)
table(bcscm_tcell_so$ClustoCell, bcscm_tcell_so$expansion)
table(bcscm_tcell_so$ClustoCell, bcscm_tcell_so$BC_type)
table(bcscm_tcell_so$ClustoCell, bcscm_tcell_so$timepoint)

# Confusion of author annotations with SubClustoCell
table(bcscm_tcell_so$cellSubType, bcscm_tcell_so$SubClustoCell)
table(bcscm_tcell_so$SubClustoCell, bcscm_tcell_so$expansion)
table(bcscm_tcell_so$SubClustoCell, bcscm_tcell_so$BC_type)
table(bcscm_tcell_so$SubClustoCell, bcscm_tcell_so$timepoint)

```

Checking the markers of clusters and subclusters

```{r}

# Major Clusters
sapply(names(bcscm_tcell_ClustoCell$markers$major_clusters$cluster_specific$positive_markers), 
       function(i) { 
         tmp_markers <- bcscm_tcell_ClustoCell$markers$major_clusters$cluster_specific$positive_markers[[i]] %>% 
           dplyr::slice(1:30) %>% dplyr::pull(Feature)
         cat(paste0(i, ": "))
          cat(paste0(tmp_markers, collapse = ", "))
          cat("\n\n")
          }
       )

# Sub-clusters
sapply(names(bcscm_tcell_ClustoCell$markers$sub_clusters), 
       function(i) { 
         cat(paste0(i, ": "))
         cat("\n")

         sapply(names(bcscm_tcell_ClustoCell$markers$sub_clusters[[i]]$positive_markers), function(j) {
           tmp_markers <- (bcscm_tcell_ClustoCell$markers$sub_clusters[[i]]$positive_markers)[[j]] %>% 
             dplyr::slice(1:30) %>% dplyr::pull(Feature)
           cat(paste0(j, ": "))
           cat(paste0(tmp_markers, collapse = ", "))
           cat("\n\n")
           })
         cat("\n\n\n")
         }
       )

```

## Data exploration and clustering

```{r}
# Subsetting pre-treatment cells with known expansion labels
bcscm_tcell_pre <- subset(bcscm_tcell_so, subset = timepoint == "Pre" & expansion %in% c("E","NE"))

library(pROC)
library(glmnet)
library(caret)
library(pheatmap)

```

#### UMAP

```{r}

bcscm_tcell_umap <- CellDimPlot(bcscm_tcell_so, reduction = "umap", group_by = "SubClustoCell", 
                                  label = F, pt_size = 1.01, 
                                  theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                  legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = -50),
    plot.subtitle = element_text(size = 7, hjust = 0.05),
    legend.margin = margin(t = 0, r = 0, b = 5, l = -5),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  ) +
  labs(subtitle = "T/NK Data SubClustoCells", 
       color = NULL, x = NULL, y = NULL, tag = "a") +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2)))

```

#### Confusion Matrix

```{r}

# With Celltypes

bcscm_tcell_SubClustoCell_vs_manual_annot <- table(SubClustoCell = bcscm_tcell_pre$SubClustoCell,
                                                   manual_annot = bcscm_tcell_pre$cellSubType)

bcscm_tcell_SubClustoCell_vs_manual_annot <- prop.table(bcscm_tcell_SubClustoCell_vs_manual_annot, margin = 1)

bcscm_tcell_SubClustoCell_vs_manual_annot <- as.data.frame(bcscm_tcell_SubClustoCell_vs_manual_annot)
colnames(bcscm_tcell_SubClustoCell_vs_manual_annot) <- c("SubClustoCell", "Manual Annotations", "Fraction")


# ---- add colored y-axis tiles
bcscm_SubClustoCell_cat_df <- bcscm_tcell_SubClustoCell_vs_manual_annot |>
  distinct(`SubClustoCell`) |>
  mutate(`SubClustoCell` = factor(levels(`SubClustoCell`), 
                                       levels = levels(`SubClustoCell`))) %>% 
  mutate(color = ggsci::pal_igv()(20))

bcscm_tcell_SubClustoCell_vs_manual_annot_plt <- ggplot(bcscm_tcell_SubClustoCell_vs_manual_annot, 
                                                     aes(x = `Manual Annotations`, 
                                                         y = `SubClustoCell`,  color = Fraction)) +
  geom_point(size = bcscm_tcell_SubClustoCell_vs_manual_annot$Fraction) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(
    low = "white", 
    high = muted("#DC0000FF"),
    labels = label_number(scale_cut = cut_short_scale(), accuracy = 0.1)
  ) +
  
  ggnewscale::new_scale_fill() +                             # <- start fresh

  geom_tile(
  data = bcscm_SubClustoCell_cat_df,
  aes(y = `SubClustoCell`, x = -0.5, fill = color),
  inherit.aes = FALSE,
  height = 0.75
) +
  scale_fill_identity(guide = "none") +                    # use literal colors for strip
  scale_y_discrete(expand = expansion(add = 0.6)) +  # leave space at bottom
  coord_cartesian(clip = "off")  +      # prevent clipping of strip
  theme_minimal() +
  labs(title = NULL, tag = "b", subtitle = "Pre-treatment"
       ) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(plot.margin = margin(b = -50),
        text = element_text(size = 7), 
        # axis.title.y = element_blank(), 
        axis.text.y = element_text(size = 5), 
        axis.ticks.y = element_blank(),
        legend.position = "top", 
        legend.direction = "horizontal",
        legend.margin = margin(2, 2, 2, 2),      # Internal legend padding
        legend.box.margin = margin(t = 0, r = 0, b = 2, l = 0),  # External space to plot
        legend.box.spacing = unit(1, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm"),
        panel.grid.major.y = element_line(color = "grey", size = 0.3)) +
  guides(color = guide_colorbar(barheight = unit(0.2, "cm"), 
                                barwidth = unit(1.5, "cm")))

#____________________________________

# With expansion

bcscm_tcell_SubClustoCell_vs_expansion <- table(SubClustoCell = bcscm_tcell_pre$SubClustoCell[bcscm_tcell_pre$expansion != "n/a"],
                                                   expansion = bcscm_tcell_pre$expansion[bcscm_tcell_pre$expansion != "n/a"])

bcscm_tcell_SubClustoCell_vs_expansion <- as.data.frame(bcscm_tcell_SubClustoCell_vs_expansion)
colnames(bcscm_tcell_SubClustoCell_vs_expansion) <- c("SubClustoCell", "Expansion", "Count")

bcscm_tcell_SubClustoCell_vs_expansion_plt <- ggplot(bcscm_tcell_SubClustoCell_vs_expansion, 
                                                        aes(x = `Expansion`, 
                                                            y = `SubClustoCell`,  color = Count)) +
  geom_point(size = log10(bcscm_tcell_SubClustoCell_vs_expansion$Count)/3) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(
    low = "white", 
    high = muted("#DC0000FF"),
    labels = label_number(scale_cut = cut_short_scale(), accuracy = 1)
  ) +
    ggnewscale::new_scale_fill() +                             # <- start fresh

  geom_tile(
  data = bcscm_SubClustoCell_cat_df,
  aes(y = `SubClustoCell`, x = -0.5, fill = color),
  inherit.aes = FALSE,
  height = 0.75
) +
  scale_fill_identity(guide = "none") +                    # use literal colors for strip
  scale_y_discrete(expand = expansion(add = 0.6)) +  # leave space at bottom
  coord_cartesian(clip = "off")  +      # prevent clipping of strip
  theme_minimal() +
  labs(title = NULL, tag = "c", subtitle = "Pre-treatment"
  ) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(plot.margin = margin(b = -50),
        text = element_text(size = 7),
        axis.title.y = element_blank(), axis.text.y = element_text(size = 5), 
        # axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        legend.margin = margin(2, 2, 2, 2),      # Internal legend padding
        legend.box.margin = margin(t = 0, r = 0, b = 0, l = 5),  # External space to plot
        legend.box.spacing = unit(1, "pt"),
        legend.position = "top", 
        legend.direction = "horizontal",
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  guides(color = guide_colorbar(barheight = unit(0.2, "cm"), 
                                barwidth = unit(1.5, "cm")))


```

## First row figure

```{r}

# Combine plots

bcscm_first_row <- (
    bcscm_tcell_umap + 
    # patchwork::plot_spacer() +
    ((bcscm_tcell_SubClustoCell_vs_manual_annot_plt +
    bcscm_tcell_SubClustoCell_vs_expansion_plt) + plot_layout(widths = c(17, 3), heights = 1, guides = "keep", axes = "collect_y"))
) + plot_layout(widths = c(10, 23), heights = 1, guides = "keep") & theme(plot.margin = margin(t = 0, r = 4, b = 2, l = 2.9))


ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_first_row.png", 
       plot = bcscm_first_row, device = "png", width = 6.85, height = (1.762)*1.6, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_first_row.pdf", 
       plot = bcscm_first_row, device = "pdf", width = 6.85, height = (1.762)*1.6, units = "in", dpi = 300)


```

## Predictive cellSubTypes (based on the paper)

```{r}

patient_summary_CD8_CD4_EX_pre_manual <- bcscm_tcell_pre@meta.data %>%
  dplyr::filter(expansion %in% c("E","NE")) %>%
  group_by(patient_id) %>%
  summarise(
    expansion_status = unique(expansion),
    BC_type = unique(BC_type),
    n_cells = n(),
    frac_CD8_EX = mean(cellSubType == "CD8_EX"),
    frac_CD8_EX_Prolif = mean(cellSubType == "CD8_EX_Prolif"),
    frac_CD8_EX_Prolif = mean(cellSubType == "CD8_EX_Prolif"),
    frac_CD4_EX = mean(cellSubType == "CD4_EX"),
    frac_CD4_EX_Prolif = mean(cellSubType == "CD4_EX_Prolif"),
  ) %>%
  mutate(expansion_binary = ifelse(expansion_status == "E", 1, 0))

#  Box plot of (E vs NE)
patient_summary_CD8_CD4_EX_pre_manual_long <- patient_summary_CD8_CD4_EX_pre_manual %>%
  pivot_longer(starts_with("frac_"), names_to = "cellSubType", values_to = "fraction")

patient_summary_CD8_CD4_EX_pre_manual_long$cellSubType <- gsub("frac_", "", patient_summary_CD8_CD4_EX_pre_manual_long$cellSubType)

patient_summary_CD8_CD4_EX_pre_manual_long$cellSubType <- recode(patient_summary_CD8_CD4_EX_pre_manual_long$cellSubType,
  "CD8_EX" = "CD8-EX",
  "CD8_EX_Prolif" = "CD8-EX-Prolif",
  "CD4_EX" = "CD4-EX",
  "CD4_EX_Prolif" = "CD4-EX-Prolif"
)

patient_summary_CD8_CD4_EX_pre_boxplt <- make_facet_boxplot_with_stats(df = patient_summary_CD8_CD4_EX_pre_manual_long, 
                                                                  x = "expansion_status",
                                                                  y = "fraction",
                                                                  fill = "expansion_status", 
                                                                  palette = c("NE" = "gray80", "E" = "#E49E26"),
                                                                  facet.by = "cellSubType", facet.nrow = 2,
                                                                  subtitle = NULL, 
                                                                  ylab = "Per−patient cell fraction",
                                                                  tag = "d")

#______________________

# Elastic-net modelling
X_CD8_CD4_EX <- as.matrix(patient_summary_CD8_CD4_EX_pre_manual %>%
                       dplyr::mutate("log10_n_cells" = log10(n_cells)) %>% 
                       dplyr::select(starts_with("frac_"), log10_n_cells))

y_CD8_CD4_EX <- patient_summary_CD8_CD4_EX_pre_manual$expansion_binary

set.seed(1234)
cvfit_CD8_CD4_EX <- cv.glmnet(X_CD8_CD4_EX, y_CD8_CD4_EX, family = "binomial", alpha = 0.5, nfolds = 5)
plot(cvfit_CD8_CD4_EX)  # Supplementary Figure (if asked for)
coef(cvfit_CD8_CD4_EX, s = "lambda.min")

#_______________________

# Coefficient visualization
coefs_CD8_CD4_EX <- as.matrix(coef(cvfit_CD8_CD4_EX, s = "lambda.min"))[-1, , drop = FALSE]  # drop intercept
coefs_CD8_CD4_EX_df <- data.frame(
  term = rownames(coefs_CD8_CD4_EX),
  estimate = as.numeric(coefs_CD8_CD4_EX[,1])
)

coefs_CD8_CD4_EX_df$term <- recode(coefs_CD8_CD4_EX_df$term,
  "frac_CD8_EX" = "CD8-EX",
  "frac_CD8_EX_Prolif" = "CD8-EX-Prolif",
  "frac_CD4_EX" = "CD4-EX",
  "frac_CD4_EX_Prolif" = "CD4-EX-Prolif",
  "log10_n_cells" = "log10(total T cells)"
)

patient_summary_CD8_CD4_EX_pre_pred_direction_plt <- 
  ggplot(coefs_CD8_CD4_EX_df, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#1E88E5", "FALSE" = "gray60")) +
  theme_minimal(base_size = 6) +
  labs(x = NULL, y = "Coefficient (elastic-net, λmin)", tag = "e",
       title = "Predictive Directions") +
      theme(
    plot.subtitle = element_text(size = 7),
    plot.title = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(size = 6),
    axis.title.x = element_text(size = 7),
    strip.text = element_text(size = 5.5),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8), line = element_line(linewidth = 0.1), 
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 5, 0, 5),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 0),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 5), 
    legend.position = "top", legend.direction = "horizontal", 
    legend.key.width = unit(6, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )

#_______________________

# ROC analysis and ROC plot
pred_prob_CD8_CD4_EX <- predict(cvfit_CD8_CD4_EX, newx = X_CD8_CD4_EX, s = "lambda.min", type = "response")[,1]
roc_obj_CD8_CD4_EX <- roc(y_CD8_CD4_EX, pred_prob_CD8_CD4_EX)
auc_val_CD8_CD4_EX <- auc(roc_obj_CD8_CD4_EX)
cat("AUC =", auc_val_CD8_CD4_EX, "\n")

patient_summary_CD8_CD4_EX_pre_manual_roc_plt <- ggroc(roc_obj_CD8_CD4_EX, color = "#E64B35", size = 0.5) +
  theme_classic(base_size = 6) +
  annotate("text", x = 0.6, y = 0.1, label = paste0("AUROC = ", round(auc_val_CD8_CD4_EX, 2))) +
  labs(subtitle = "Elastic−net Model: Predictive Accuracy", x = "Specificity", 
       y = "Sensitivity", tag = "f") +
  theme(
    plot.subtitle = element_text(size = 7),
    plot.title = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(size = 6),
    strip.text = element_text(size = 5.5),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8), line = element_line(linewidth = 0.1), 
    axis.title = element_text(size = 7),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 5, 0, 5),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 0),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 5), 
    legend.position = "top", legend.direction = "horizontal", 
    legend.key.width = unit(6, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )

```

## Second row figure

```{r}

# Combine plots

bcscm_paperTypes_second_row <- (
    (wrap_elements(patient_summary_CD8_CD4_EX_pre_boxplt) +
       wrap_elements(patient_summary_CD8_CD4_EX_pre_pred_direction_plt) +
       wrap_elements(patient_summary_CD8_CD4_EX_pre_manual_roc_plt) +
       plot_layout(ncol = 3, widths = c(3, 3, 3))
    )
) + 
  plot_layout(heights = 1, guides = "keep")


ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_paperTypes_second_row.png", 
       plot = bcscm_paperTypes_second_row, device = "png", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_paperTypes_second_row.pdf", 
       plot = bcscm_paperTypes_second_row, device = "pdf", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)


```

## Evaluating the top sub-clusters

```{r}

library(pROC)
library(glmnet)

subClust_Exp_Type_df <- bcscm_tcell_pre@meta.data[, c("SubClustoCell", "expansion", "cellSubType")] %>% 
  mutate(expansion = ifelse(expansion == "E", 1, 0))

subClust_Exp_Type_df$SubClustoCell <- as.character(subClust_Exp_Type_df$SubClustoCell)

expansion_enriched_subs <- 
  subClust_Exp_Type_df %>%
  group_by(SubClustoCell) %>%
  summarise(
    n = n(),
    prop = mean(expansion)
  ) %>% 
  dplyr::filter(prop > mean(subClust_Exp_Type_df$expansion)) %>% 
  dplyr::pull(SubClustoCell) %>% as.character()

subClust_Exp_Type_df <- subClust_Exp_Type_df %>% dplyr::filter(SubClustoCell %in% expansion_enriched_subs)

#_________________

# Fisher exact test

SubClustoCells <- unique(subClust_Exp_Type_df$SubClustoCell)

set.seed(1234)
top_SubClustoCells <- lapply(SubClustoCells, function(c){
  sub_df <- data.frame(
    x = factor(subClust_Exp_Type_df$SubClustoCell == c, levels = c(TRUE, FALSE)),
    y = subClust_Exp_Type_df$expansion
  )
  test <- fisher.test(table(sub_df))
  data.frame(
    SubClustoCell = c,
    p_value = test$p.value,
    odds_ratio = test$estimate
  )
})

top_SubClustoCells <- do.call(rbind, top_SubClustoCells) %>% 
  dplyr::slice_min(p_value, n = 4)

```

### Evaluating the response-ready circuit

Evaluating the C4-Sub4, C5-Sub2, C5-Sub3, C2-Sub1 (top_SubClustoCells) among T/NK cells, identified through exploration of the table of clusters and cell types/states and their confusions.

- C5-Sub2 (terminally exhausted CD8 effectors with high inhibitory signaling),
- C5-Sub3 (interferon-licensed CD8 exhausted/effector progenitors),
- C4-Sub4 (CXCL13+ CD4 TFH/TEX-like checkpoint-rich effectors),
- C2-Sub1 (CD4 effector-memory state with GZMK/GZMA cytotoxicity and CXCL13-linked migratory potential),

##### Predictive Validity

```{r}

library(pROC)
library(glmnet)
library(caret)
library(pheatmap)

# Define predictive subclusters (from your discovery)
bcscm_tcell_predictive_subs <- c("C4-Sub4", "C5-Sub2", "C5-Sub3", "C2-Sub1")

# Computing patient level fractions
patient_summary_tcell_pre <- bcscm_tcell_pre@meta.data %>%
  dplyr::filter(expansion %in% c("E","NE")) %>%
  mutate(is_predictive = SubClustoCell %in% bcscm_tcell_predictive_subs) %>%
  group_by(patient_id) %>%
  summarise(
    expansion_status = unique(expansion),
    BC_type = unique(BC_type),
    n_cells = n(),
    frac_C4Sub4 = mean(SubClustoCell == "C4-Sub4"),
    frac_C5Sub2 = mean(SubClustoCell == "C5-Sub2"),
    frac_C5Sub3 = mean(SubClustoCell == "C5-Sub3"),
    frac_C2Sub1 = mean(SubClustoCell == "C2-Sub1")
  ) %>%
  mutate(expansion_binary = ifelse(expansion_status == "E", 1, 0))

#  Box plot of (E vs NE)
patient_summary_tcell_pre_long <- patient_summary_tcell_pre %>%
  pivot_longer(starts_with("frac_"), names_to = "subcluster", values_to = "fraction")

patient_summary_tcell_pre_long$subcluster <- recode(patient_summary_tcell_pre_long$subcluster,
  "frac_C4Sub4" = "C4-Sub4",
  "frac_C5Sub2" = "C5-Sub2",
  "frac_C5Sub3" = "C5-Sub3",
  "frac_C2Sub1" = "C2-Sub1",
  "log10_n_cells" = "log10(total T cells)"
)


patient_summary_tcell_pre_boxplt <- make_facet_boxplot_with_stats(df = patient_summary_tcell_pre_long, 
                                                                  x = "expansion_status",
                                                                  y = "fraction",
                                                                  fill = "expansion_status", 
                                                                  palette = c("NE" = "gray80", "E" = "#D81B60"),
                                                                  facet.by = "subcluster",
                                                                  subtitle = "Pre-treatment Frequencies of\nPredictive ClustoCell Subclusters", 
                                                                  ylab = "Per−patient cell fraction",
                                                                  tag = "g")

#_______________________

# Heatmap of gene signatures

# Defining your signatures
sig_panels_bcscm_tcell <- list(
  `C4-Sub4` = c("PDCD1", "LAG3", "CTLA4", "TIGIT", "CXCL13", "TNFRSF18", "CTSD", "PRDM1"),
  `C5-Sub2` = c("CD8A", "GZMB", "LAG3", "TIGIT", "CXCL13", "PRF1", "GZMH", "NKG7", "HLA-DRA", "HLA-DRB1", "HLA-DRB5"),
  `C5-Sub3` = c("CD8A", "GZMB", "LAG3", "CXCL13", "GNLY", "PRF1", "GZMH", "NKG7", "CCL4", "HLA-DRA", "HLA-DRB1", "HLA-DRB5"),
  `C2-Sub1` = c("GZMK", "KLRB1", "CXCL13", "ANXA1", "MTRNR2L8")
)

# Computing average expression per SubClustoCell
sig_panels_bcscm_tcell_avg_expr <- AverageExpression(
  bcscm_tcell_pre,
  group.by = "SubClustoCell",
  features = unlist(sig_panels_bcscm_tcell) %>% unique()
)$RNA

# Subsetting only predictive subclusters
predictive_subs <- c("C4-Sub4", "C5-Sub2", "C5-Sub3", "C2-Sub1")
sig_panels_bcscm_tcell_avg_expr <- sig_panels_bcscm_tcell_avg_expr[, colnames(sig_panels_bcscm_tcell_avg_expr) %in% predictive_subs]

# Create combined gene × signature labels to ensure uniqueness
sig_panels_bcscm_tcell_gene_order <- unlist(sig_panels_bcscm_tcell)
sig_panels_bcscm_tcell_group_labels <- rep(names(sig_panels_bcscm_tcell), lengths(sig_panels_bcscm_tcell))

# scale per gene (row)
sig_panels_bcscm_tcell_sig_scaled <- t(scale(t(sig_panels_bcscm_tcell_avg_expr[sig_panels_bcscm_tcell_gene_order, predictive_subs])))

# Make a data frame that keeps both
sig_panels_bcscm_tcell_ann_row <- data.frame(
  Gene = sig_panels_bcscm_tcell_gene_order,
  Signature = sig_panels_bcscm_tcell_group_labels
)

# Create unique rownames like "CXCL13 (C4-Sub4)" if duplicates exist
sig_panels_bcscm_tcell_ann_row <- sig_panels_bcscm_tcell_ann_row %>% 
  dplyr::distinct(Gene, .keep_all = T)

rownames(sig_panels_bcscm_tcell_ann_row) <- sig_panels_bcscm_tcell_ann_row$Gene

# Remove duplicates from sig_panels_bcscm_tcell_sig_scaled as well
sig_panels_bcscm_tcell_sig_scaled <- sig_panels_bcscm_tcell_sig_scaled[rownames(sig_panels_bcscm_tcell_ann_row),]

# shortenning label text for display
sig_panels_bcscm_tcell_ann_row$Signature <- factor(
  sig_panels_bcscm_tcell_ann_row$Signature,
  levels = names(sig_panels_bcscm_tcell)
)

library(ggalign)

# Correcting the orders as desired
sig_panels_bcscm_tcell_sig_scaled <- sig_panels_bcscm_tcell_sig_scaled[c(1,3,5,6,8,7,4,2,9:18, 21, 20, 22, 19),]

sig_panels_bcscm_tcell_ggheatmap <- 
  ggheatmap(sig_panels_bcscm_tcell_sig_scaled,
            # width = unit(6, "cm"), height = unit(10, "cm")
  ) +
  scale_fill_gradient2(
    low = "navy", mid = "white", high = "firebrick3", midpoint = 0,
    name = "Scaled\nExpr."
  ) +
  labs(subtitle = NULL, tag = "h") +
    theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 18, vjust = 0.7, face = "plain", color = "black"),
        axis.text = element_text(size = 6),
        text = element_text(size = 7),
        title = element_text(size = 7),
        plot.title = element_text(size = 7),
        plot.subtitle = element_text(size = 7, vjust = 0),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 6),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 0),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.4, "cm"))

#______________________

# Elastic-net modelling
X <- as.matrix(patient_summary_tcell_pre %>%
                  dplyr::mutate("log10_n_cells" = log10(n_cells)) %>% 
                  dplyr::select(starts_with("frac_"), log10_n_cells))

y <- patient_summary_tcell_pre$expansion_binary

cvfit <- cv.glmnet(X, y, family = "binomial", alpha = 0.5, nfolds = 5)
plot(cvfit)  # Supplementary Figure (if asked for)
coef(cvfit, s = "lambda.min")

#_______________________

# Coefficient visualization
coefs <- as.matrix(coef(cvfit, s = "lambda.min"))[-1, , drop = FALSE]  # drop intercept
coef_df <- data.frame(
  term = rownames(coefs),
  estimate = as.numeric(coefs[,1])
)

coef_df$term <- recode(coef_df$term,
  "frac_C4Sub4" = "C4-Sub4",
  "frac_C5Sub2" = "C5-Sub2",
  "frac_C5Sub3" = "C5-Sub3",
  "frac_C2Sub1" = "C2-Sub1",
  "log10_n_cells" = "log10(total T cells)"
)

patient_summary_tcell_pre_focus_pred_direction_plt <- 
  ggplot(coef_df, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#1E88E5", "FALSE" = "gray60")) +
  theme_minimal(base_size = 6) +
  labs(x = NULL, y = "Coefficient (elastic-net, λmin)", tag = "i",
       title = "Predictive Directions") +
      theme(
    plot.subtitle = element_text(size = 7),
    plot.title = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(size = 6),
    axis.title.x = element_text(size = 7),
    strip.text = element_text(size = 5.5),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8), line = element_line(linewidth = 0.1), 
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 5, 0, 5),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 0),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 5), 
    legend.position = "top", legend.direction = "horizontal", 
    legend.key.width = unit(6, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )

#_______________________

# ROC analysis and ROC plot
pred_prob <- predict(cvfit, newx = X, s = "lambda.min", type = "response")[,1]
roc_obj <- roc(y, pred_prob)
auc_val <- auc(roc_obj)
cat("AUC =", auc_val, "\n")

patient_summary_tcell_pre_roc_plt <- ggroc(roc_obj, color = "#E64B35", size = 0.5) +
  theme_classic(base_size = 6) +
  annotate("text", x = 0.6, y = 0.1, label = paste0("AUROC = ", round(auc_val, 2))) +
  labs(subtitle = "Elastic-net Model: Predictive Accuracy", x = "Specificity", 
       y = "Sensitivity", tag = "j") +
      theme(
    plot.subtitle = element_text(size = 7),
    plot.title = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(size = 6),
    strip.text = element_text(size = 5.5),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8), line = element_line(linewidth = 0.1), 
    axis.title = element_text(size = 7),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 5, 0, 5),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 0),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 5), 
    legend.position = "top", legend.direction = "horizontal", 
    legend.key.width = unit(6, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )


```

## Third row figure

```{r}

# Combine plots

bcscm_third_row <- (
wrap_elements(patient_summary_tcell_pre_boxplt) + 
  wrap_elements(ggalign::ggalignGrob(sig_panels_bcscm_tcell_ggheatmap)) +
  ((wrap_elements(patient_summary_tcell_pre_focus_pred_direction_plt) / patchwork::plot_spacer() / wrap_elements(patient_summary_tcell_pre_roc_plt)) +
     plot_layout(heights = c(13, 2, 14), widths = 1, guides = "keep") & theme(plot.margin = margin(0, 0, 0, 0)))
) + 
  plot_layout(widths = c(12, 12, 12), heights = 1, guides = "keep")


ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_third_row.png", 
       plot = bcscm_third_row, device = "png", width = 6.85, height = (1.762)*1.4, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_third_row.pdf", 
       plot = bcscm_third_row, device = "pdf", width = 6.85, height = (1.762)*1.4, units = "in", dpi = 300)


```

#### CD4_EM C2-Sub1

##### Predictive Validity

```{r}

patient_summary_CD4_EM_pre_manual <- bcscm_tcell_pre@meta.data %>%
  dplyr::filter(expansion %in% c("E","NE")) %>%
  group_by(patient_id) %>%
  summarise(
    expansion_status = unique(expansion),
    BC_type = unique(BC_type),
    n_cells = n(),
    frac_CD4_EM = mean(cellSubType == "CD4_EM")
  ) %>%
  mutate(expansion_binary = ifelse(expansion_status == "E", 1, 0))

#  Box plot of (E vs NE)
patient_summary_CD4_EM_pre_manual_long <- patient_summary_CD4_EM_pre_manual %>%
  pivot_longer(starts_with("frac_"), names_to = "cellSubType", values_to = "fraction")

patient_summary_CD4_EM_pre_manual_long$cellSubType <- gsub("frac_", "", patient_summary_CD4_EM_pre_manual_long$cellSubType)


patient_summary_CD4_EM_pre_boxplt <- make_facet_boxplot_with_stats(df = patient_summary_CD4_EM_pre_manual_long, 
                                                                  x = "expansion_status",
                                                                  y = "fraction",
                                                                  fill = "expansion_status", 
                                                                  palette = c("NE" = "gray80", "E" = "#E49E26"),
                                                                  facet.by = "cellSubType",
                                                                  subtitle = NULL, 
                                                                  ylab = "Per−patient cell fraction",
                                                                  tag = "k")

#__________________________________

# Computing patient level fractions
patient_summary_CD4_EM_C2Sub1_pre <- bcscm_tcell_pre@meta.data %>%
  dplyr::filter(expansion %in% c("E","NE"), cellSubType == "CD4_EM") %>%
  group_by(patient_id) %>%
  summarise(
    expansion_status = unique(expansion),
    BC_type = unique(BC_type),
    n_cells = n(),
    frac_C2Sub1 = mean(SubClustoCell == "C2-Sub1"),
  ) %>%
  mutate(expansion_binary = ifelse(expansion_status == "E", 1, 0))

#  Box plot of (E vs NE)
patient_summary_CD4_EM_C2Sub1_pre_long  <- patient_summary_CD4_EM_C2Sub1_pre %>%
  pivot_longer(starts_with("frac_"), names_to = "subcluster", values_to = "fraction")

patient_summary_CD4_EM_C2Sub1_pre_long$subcluster <- recode(patient_summary_CD4_EM_C2Sub1_pre_long$subcluster,
                                                           "frac_C2Sub1" = "C2-Sub1 CD4 EM"
)

patient_summary_CD4_EM_C2Sub1_pre_boxplt <- make_facet_boxplot_with_stats(df = patient_summary_CD4_EM_C2Sub1_pre_long, 
                                                                         x = "expansion_status",
                                                                         y = "fraction",
                                                                         fill = "expansion_status", 
                                                                         palette = c("NE" = "gray80", "E" = "#D81B60"),
                                                                         facet.by = "subcluster",
                                                                         subtitle = NULL, 
                                                                         ylab = "Per−patient cell fraction",
                                                                         tag = NULL)

#______________________

# Elastic-net modelling

patient_summary_CD4_EM_combined_C2Sub1_pre <- bcscm_tcell_pre@meta.data %>%
  dplyr::filter(expansion %in% c("E","NE")) %>%
  group_by(patient_id) %>%
  summarise(
    expansion_status = unique(expansion),
    BC_type = unique(BC_type),
    n_cells = n(),
    frac_CD8_EX = mean(cellSubType == "CD8_EX"),
    frac_CD8_EX_Prolif = mean(cellSubType == "CD8_EX_Prolif"),
    frac_CD4_EX = mean(cellSubType == "CD4_EX"),
    frac_CD4_EX_Prolif = mean(cellSubType == "CD4_EX_Prolif"),
    frac_C2Sub1_CD4_EM = mean(SubClustoCell == "C2-Sub1" & cellSubType %in% c("CD4_EM"))
  ) %>%
  mutate(expansion_binary = ifelse(expansion_status == "E", 1, 0))

X_CD4_EM_combined_C2Sub1 <- as.matrix(patient_summary_CD4_EM_combined_C2Sub1_pre %>%
                                        dplyr::mutate("log10_n_cells" = log10(n_cells)) %>% 
                                        dplyr::select(starts_with("frac_"), log10_n_cells))

y_CD4_EM_combined_C2Sub1 <- patient_summary_CD4_EM_combined_C2Sub1_pre$expansion_binary

set.seed(1234)
cvfit_CD4_EM_combined_C2Sub1 <- cv.glmnet(X_CD4_EM_combined_C2Sub1, y_CD4_EM_combined_C2Sub1, family = "binomial", alpha = 0.5, nfolds = 5)
plot(cvfit_CD4_EM_combined_C2Sub1)  # Supplementary Figure (if asked for)
coef(cvfit_CD4_EM_combined_C2Sub1, s = "lambda.min")

#_______________________

# Coefficient visualization

coefs_CD4_EM_combined_C2Sub1 <- as.matrix(coef(cvfit_CD4_EM_combined_C2Sub1, s = "lambda.min"))[-1, , drop = FALSE]  # drop intercept
coef_df_CD4_EM_combined_C2Sub1 <- data.frame(
  term = rownames(coefs_CD4_EM_combined_C2Sub1),
  estimate = as.numeric(coefs_CD4_EM_combined_C2Sub1[,1])
)



coef_df_CD4_EM_combined_C2Sub1$term <- recode(coef_df_CD4_EM_combined_C2Sub1$term,
                                              "frac_C2Sub1_CD4_EM" = "C2-Sub1_CD4_EM",
                                              "log10_n_cells" = "log10(total T cells)"
)

coef_df_CD4_EM_combined_C2Sub1$term <- gsub("frac_", "", coef_df_CD4_EM_combined_C2Sub1$term)

patient_summary_tcell_pre_focus_pred_CD4_EM_combined_C2Sub1_direction_plt <- 
  ggplot(coef_df_CD4_EM_combined_C2Sub1, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#1E88E5", "FALSE" = "gray60")) +
  theme_minimal(base_size = 6) +
  labs(x = NULL, y = "Coefficient (elastic-net, λmin)", tag = NULL,
       title = "Predictive Directions") +
  theme(
    plot.subtitle = element_text(size = 7),
    plot.title = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(size = 6),
    axis.title.x = element_text(size = 7),
    strip.text = element_text(size = 5.5),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8), line = element_line(linewidth = 0.1), 
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 5, 0, 5),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 0),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 5), 
    legend.position = "top", legend.direction = "horizontal", 
    legend.key.width = unit(6, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )

#_______________________

# ROC analysis and ROC plot
set.seed(1234)
pred_prob_CD4_EM_combined_C2Sub1 <- predict(cvfit_CD4_EM_combined_C2Sub1, newx = X_CD4_EM_combined_C2Sub1, s = "lambda.min", type = "response")[,1]
roc_obj_CD4_EM_combined_C2Sub1 <- roc(y_CD4_EM_combined_C2Sub1, pred_prob_CD4_EM_combined_C2Sub1)
auc_val_CD4_EM_combined_C2Sub1 <- auc(roc_obj_CD4_EM_combined_C2Sub1)
cat("AUC =", auc_val_CD4_EM_combined_C2Sub1, "\n")

patient_summary_CD4_EM_combined_C2Sub1_pre_roc_plt <- ggroc(roc_obj_CD4_EM_combined_C2Sub1, color = "#E64B35", size = 0.5) +
  theme_classic(base_size = 6) +
  annotate("text", x = 0.6, y = 0.1, label = paste0("AUROC = ", round(auc_val_CD4_EM_combined_C2Sub1, 2))) +
  labs(subtitle = "Predictive Cell Subtypes +\nC2-Sub1-CD4-EM", x = "Specificity", 
       y = "Sensitivity", tag = NULL) +
  theme(
    plot.subtitle = element_text(size = 7),
    plot.title = element_text(size = 7),
    legend.title = element_blank(),
    text = element_text(size = 6),
    strip.text = element_text(size = 5.5),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8), line = element_line(linewidth = 0.1), 
    axis.title = element_text(size = 7),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 5, 0, 5),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 0),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.2, "cm"),
    legend.text = element_text(size = 5), 
    legend.position = "top", legend.direction = "horizontal", 
    legend.key.width = unit(6, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )

```

#### Fourth row figure

```{r}

# Combine plots

bcscm_conflicting_row <- (
    wrap_plots(
      wrap_elements(
        wrap_plots(patient_summary_CD4_EM_pre_boxplt, 
                   patient_summary_CD4_EM_C2Sub1_pre_boxplt, 
                   ncol = 2, axes = "collect")),
    
    wrap_elements(
        patient_summary_tcell_pre_focus_pred_CD4_EM_combined_C2Sub1_direction_plt),
    
        wrap_elements(
            patient_summary_CD4_EM_combined_C2Sub1_pre_roc_plt),
    widths = c(1.8, 1.5, 1.6), heights = 1, guides = "keep"
    ) & theme(plot.margin = margin(0, 0, b = 3, l = 2.9))
) + plot_layout(widths = c(50), heights = 1, guides = "keep")


ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_conflicting_row.png", 
       plot = bcscm_conflicting_row, device = "png", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/bcscm/Final/bcscm_conflicting_row.pdf", 
       plot = bcscm_conflicting_row, device = "pdf", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)


```

# CD4-EM DE analysis

```{r}

# DE analysis between subsets

# CD4_EM
CD4_EM_C2Sub1_vs_C3Sub2 <- Seurat::FindMarkers(subset(bcscm_tcell_pre, subset = cellSubType == "CD4_EM"), 
                                                group.by = "SubClustoCell", 
                                                ident.1 = "C2-Sub1", 
                                                ident.2 = "C3-Sub2")

qs2::qs_save(object = CD4_EM_C2Sub1_vs_C3Sub2, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/CD4_EM_C2Sub1_vs_C3Sub2.qs2")

CD4_EM_C2Sub1_vs_C3Sub2 <- qs2::qs_read("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/CD4_EM_C2Sub1_vs_C3Sub2.qs2")

CD4_EM_C2Sub1_vs_C3Sub2_top_ups <- CD4_EM_C2Sub1_vs_C3Sub2 %>% dplyr::filter(p_val_adj < 0.05, avg_log2FC > 0.585) %>% rownames()

# GZMA and GZMK

CD4_EM_C2Sub1_vs_C3Sub2_top_downs <- CD4_EM_C2Sub1_vs_C3Sub2 %>% dplyr::filter(p_val_adj < 0.05, avg_log2FC < -0.585) %>% dplyr::slice_min(order_by = avg_log2FC, n = 20) %>% rownames()

# ST6GALNAC1

#______________________________

# Ploting the DE data

CD4_EM_C2Sub1_vs_C3Sub2$Log2FC <- CD4_EM_C2Sub1_vs_C3Sub2$avg_log2FC
CD4_EM_C2Sub1_vs_C3Sub2$padj <- CD4_EM_C2Sub1_vs_C3Sub2$p_val_adj

CD4_EM_C2Sub1_vs_C3Sub2$reg <- 
  ifelse(CD4_EM_C2Sub1_vs_C3Sub2$Log2FC > 0.5 & 
           CD4_EM_C2Sub1_vs_C3Sub2$padj < 0.05, "Up",
         ifelse(CD4_EM_C2Sub1_vs_C3Sub2$Log2FC < -0.5 & 
                  CD4_EM_C2Sub1_vs_C3Sub2$padj < 0.05, "Down", "NS"))

CD4_EM_C2Sub1_vs_C3Sub2$reg <- factor(CD4_EM_C2Sub1_vs_C3Sub2$reg, 
                                                             levels=c("Down","Up","NS"))

CD4_EM_C2Sub1_vs_C3Sub2$label <- ifelse(rownames(CD4_EM_C2Sub1_vs_C3Sub2) %in% c(
  CD4_EM_C2Sub1_vs_C3Sub2 %>% dplyr::filter(padj < 0.05) %>% rownames_to_column("Feature") %>% dplyr::slice_max(order_by = Log2FC, n = 5) %>% dplyr::pull(Feature),
  CD4_EM_C2Sub1_vs_C3Sub2 %>% dplyr::filter(padj < 0.05) %>% rownames_to_column("Feature") %>% dplyr::slice_min(order_by = Log2FC, n = 5) %>% dplyr::pull(Feature)
), rownames(CD4_EM_C2Sub1_vs_C3Sub2), NA)

#_______

CD4_EM_C2Sub1_vs_C3Sub2_plt <- 
  ggplot(CD4_EM_C2Sub1_vs_C3Sub2, 
         aes(Log2FC, -log10(padj), color = reg)) +
  
  geom_point(size = 1.3, alpha = 0.9) +
  
  ggrepel::geom_text_repel(
    aes(label = label),
    min.segment.length = 0.1,
    segment.size = 0.2,
    segment.alpha = 0.8, 
    colour = "black",
    size = 2.2,              
    max.overlaps = Inf,
    box.padding = 0.35,
    point.padding = 0.2,
    seed = 42
  ) +
  
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed") +
  
  scale_color_manual(
    values = c("Down" = "royalblue", "Up" = "red2", "NS" = "grey80"),
    labels = c("Down", "Up", "NS")
  ) +
  
  labs(tag = "a",
       caption = "FC cutoff = 0.5; padj cutoff = 0.05",
       title = "CD4 EM",
       subtitle = "C2-Sub1 vs. C3-Sub2",
       x = "Log2FC",
       y = "-Log10 padj",
       color = ""
  ) +
  
  theme_classic(base_size = 7) +
  theme(
    plot.tag = element_text(size = 7, face = "bold"),
    plot.caption = element_blank(),
    plot.title = element_text(size = 7, face = "bold"),
    plot.subtitle = element_text(size = 7),
    legend.position = "top",
    legend.text = element_text(size = 6),
    axis.text = element_text(size = 6),
    axis.title = element_text(size = 7),
    panel.grid = element_blank()
  ) +
  guides(color = guide_legend(override.aes = list(size = 2)))

#_________________________

# Plotting table of DEGs

CD4_EM_C2Sub1_vs_C3Sub2_top_ups_df <-
  data.frame(
    Rank1_8 = paste(1:8, CD4_EM_C2Sub1_vs_C3Sub2_top_ups[1:8], sep = " : ")
  )

CD4_EM_C2Sub1_vs_C3Sub2_top_ups_plt <- ggtexttable(
  CD4_EM_C2Sub1_vs_C3Sub2_top_ups_df,
  rows = NULL,
  cols = NULL,  
  theme = custom_tbl_theme_size8
)

# Add minimal lines: only top and bottom horizontal lines, no vertical lines for cleaner look
CD4_EM_C2Sub1_vs_C3Sub2_top_ups_plt <- CD4_EM_C2Sub1_vs_C3Sub2_top_ups_plt %>%
  # Removed vertical lines for minimalism
  tab_add_title(text = "Up DEGs", face = "bold", size = 6, padding = unit(1, "line")) +  # Reduced padding to eliminate top margin
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 6),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    legend.margin = margin(0, 0, 0, 0),      
    legend.box.margin = margin(0, 0, 0, 0),  
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    plot.margin = margin(t = -10, r = 0, b = 10, l = 0),  # remove extra vertical padding
  ) +
  theme_void()

#___________

CD4_EM_C2Sub1_vs_C3Sub2_top_downs_df <-
  data.frame(
    Rank1_10 = paste(1:10, CD4_EM_C2Sub1_vs_C3Sub2_top_downs[1:10], sep = " : "),
    Rank11_20 = paste(11:20, CD4_EM_C2Sub1_vs_C3Sub2_top_downs[11:20], sep = " : ")
  )

CD4_EM_C2Sub1_vs_C3Sub2_top_downs_plt <- ggtexttable(
  CD4_EM_C2Sub1_vs_C3Sub2_top_downs_df,
  rows = NULL,
  cols = NULL,  
  theme = custom_tbl_theme
)

# Add minimal lines: only top and bottom horizontal lines, no vertical lines for cleaner look
CD4_EM_C2Sub1_vs_C3Sub2_top_downs_plt <- CD4_EM_C2Sub1_vs_C3Sub2_top_downs_plt %>%
  tab_add_hline(at.row = c(1), row.side = "top", linewidth = 1.5, linetype = 1) %>%  # Thinner lines for elegance
  tab_add_hline(at.row = nrow(CD4_EM_C2Sub1_vs_C3Sub2_top_downs_df), row.side = "bottom", linewidth = 1.5, linetype = 1) %>%
  # Removed vertical lines for minimalism
  tab_add_title(text = "Top 20 Down DEGs", face = "bold", size = 6, padding = unit(1, "line")) +  # Reduced padding to eliminate top margin
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 6),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    legend.margin = margin(0, 0, 0, 0),      
    legend.box.margin = margin(0, 0, 0, 0),  
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    plot.margin = margin(t = -10, r = 0, b = 10, l = 0),  # remove extra vertical padding
  ) +
  theme_void()

```

## Survival analysis

#### Up DEGs plus CD4 EM signature

```{r}

# BiocManager::install(c("curatedTCGAData", "MultiAssayExperiment", "TCGAutils", "GSVA"), update = FALSE, ask = FALSE)

# Load libraries
library(curatedTCGAData)
library(survival)
library(survminer)
library(SummarizedExperiment)
library(singscore)

# Downloading TCGA-BRCA dataset (RNA-Seq normalized gene expression and clinical data)
brca_mae <- curatedTCGAData(diseaseCode = "BRCA",
                            version = "2.1.1",
                            assays = "RNASeq2GeneNorm",
                            dry.run = FALSE)

qs2::qs_save(object = brca_mae,
             file = "../../../Benchmarking/Test Datasets/TCGA/brca_mae.qs2")

brca_mae <- qs2::qs_read("../../../Benchmarking/Test Datasets/TCGA/brca_mae.qs2")

# Extract the gene expression data (SummarizedExperiment)
exp_se <- experiments(brca_mae)[["BRCA_RNASeq2GeneNorm-20160128"]]
colnames(exp_se) <- gsub("-01$", "", colnames(exp_se))

# Extract clinical data
clinical_data <- colData(brca_mae)

# Prepare expression matrix
expr_mat <- assay(exp_se)

# Prepare gene sets list
cd4_em_nonNK_nonCD8_signature <- c("CD4", "CD40LG", "IL7R", "ITGB1", "ANXA1")
up_genes_plus <- intersect(c(CD4_EM_C2Sub1_vs_C3Sub2_top_ups, cd4_em_nonNK_nonCD8_signature), rownames(expr_mat)) %>% unique()

down_genes <- intersect(CD4_EM_C2Sub1_vs_C3Sub2_top_downs, rownames(expr_mat)) %>% unique()

length(up_genes_plus)
length(down_genes)

# Compute Singscore (combined up/down)
# Rank genes per sample
ranked_mat <- rankGenes(expr_mat)

# Compute scores
plus_scores <- simpleScore(
  ranked_mat,
  upSet = up_genes_plus,
  downSet = down_genes,
  centerScore = TRUE
)

# Extract combined score
gene_set_plus_score <- plus_scores$TotalScore
names(gene_set_plus_score) <- colnames(expr_mat)

# Prepare survival data
# In TCGA, overall survival (OS) is typically constructed from:
# - vital_status: 0 (Alive) or 1 (Dead)
# - days_to_last_followup: for censored (alive)
# - days_to_death: for events (dead)

# Create OS time (in days)
os_time <- ifelse(clinical_data$vital_status == 1, 
                  clinical_data$days_to_death, 
                  clinical_data$days_to_last_followup)

# Create OS event (1 = dead, 0 = censored)
os_event <- ifelse(clinical_data$vital_status == 1, 1, 0)

# Censoring the follow-up time to 10 years (120 months; 3600 days)
## We censor all the patients with over 10 years survival (i.e. we set their survival/vital status to 0 up until 10 years and set their days (days_to_last_followup to 3600))
os_time[which(clinical_data$days_to_death > 3600 | clinical_data$days_to_last_followup > 3600)] <- 3600
os_event[which(clinical_data$days_to_death > 3600 | clinical_data$days_to_last_followup > 3600)] <- 0

# Ensure sample names match between expression and clinical
common_samples <- intersect(names(gene_set_plus_score), rownames(clinical_data))
length(common_samples)

# Subset to common samples
gene_set_plus_score <- gene_set_plus_score[common_samples]
os_time <- os_time[match(common_samples, rownames(clinical_data))]
os_event <- os_event[match(common_samples, rownames(clinical_data))]

# Remove NA in survival data
valid_idx <- !is.na(os_time) & !is.na(os_event) & !is.na(gene_set_plus_score)
os_time <- os_time[valid_idx]
os_event <- os_event[valid_idx]
gene_set_plus_score <- gene_set_plus_score[valid_idx]

#__________________________

# Converting os_time from days to months
os_time <- os_time/30 

# Create the modeling df
cox_plus_df <- tibble(os_time,
                 os_event,
                 `Gene-set Score` = gene_set_plus_score,
                 Age = clinical_data[names(gene_set_plus_score), "Age.at.Initial.Pathologic.Diagnosis"],
                 Gender = clinical_data[names(gene_set_plus_score), "gender"],
                 `Pathological Stage` = clinical_data[names(gene_set_plus_score), "pathologic_stage"]
)

cox_plus_df$Stage <- tolower(cox_plus_df$`Pathological Stage`)

cox_plus_df <- cox_plus_df %>% 
  dplyr::mutate(
    Stage = dplyr::case_when(
      grepl("i", Stage) & !grepl("ii|iii|iv", Stage) ~ "Stage I",
      grepl("ii", Stage) & !grepl("iii|iv", Stage) ~ "Stage II",
      grepl("iii", Stage) ~ "Stage III",
      grepl("iv", Stage) ~ "Stage IV",
      TRUE ~ NA_character_
    )
  )

```

##### Multi-variate Cox

```{r}

cox_plus_df_clean <- cox_plus_df %>%
  dplyr::filter(
    !is.na(os_time),
    !is.na(os_event),
    !is.na(`Gene-set Score`),
    !is.na(Age),
    !is.na(Gender),
    !is.na(`Pathological Stage`)
  )

cox_plus_df_clean <- cox_plus_df_clean %>% dplyr::filter(!is.na(Stage))
cox_plus_df_clean$Stage <- factor(cox_plus_df_clean$Stage)

table(cox_plus_df_clean$Stage, cox_plus_df_clean$os_event)
table(cox_plus_df_clean$Gender, cox_plus_df_clean$os_event) # Problematic (complete (or quasi) separation)

# Cox model
cox_plus_model <- coxph(
  Surv(os_time, os_event) ~ 
    `Gene-set Score` +
    Age +
    Stage,
  data = cox_plus_df_clean
)

summary(cox_plus_model)

cox_plus_plt <-
  forestmodel::forest_model(cox_plus_model, 
                            theme = theme_bw() + 
                              theme(panel.grid.major = element_blank(), 
                                    panel.grid.minor = element_blank(), 
                                    axis.title.y = element_blank(), 
                                    axis.title.x = element_blank(), 
                                    axis.text.y = element_blank(), 
                                    axis.ticks.y = element_blank(),
                                    text = element_text(size = 7),
                                    strip.text = element_text(size = 7),
                                    title = element_text(size = 7),
                                    plot.title = element_text(size = 7),
                                    plot.subtitle = element_text(size = 7, vjust = 0),
                                    plot.tag = element_text(size = 7, face = "bold"),
                                    legend.title = element_text(
                                      size = 6,
                                      margin = margin(b = 10)
                                    ),
                                    legend.margin = margin(2, 2, 2, 2),
                                    legend.box.margin = margin(2, 2, 2, 2),
                                    legend.box.spacing = unit(1, "pt"),
                                    legend.key.width = unit(0.5, "cm"),
                                    legend.key.height = unit(0.5, "cm")
                              )) +
  labs(subtitle = "Singscore: Upregulated Genes and CD4 EM\nSignature − Downregulated Genes", 
       y = NULL, 
       x = NULL, 
       tag = "e")

cox_plus_plt@layers[[ which(sapply(cox_plus_plt@layers, function(x) inherits(x$geom, "GeomText"))) ]]$aes_params$size <- 2

```

##### Univariate Cox

```{r}

uni_plus_results <- purrr::map_dfr(c("Gene-set Score", "Age", "Stage"), function(v) {
  
  df <- cox_plus_df %>%
    dplyr::select(os_time, os_event, all_of(v)) %>%
    dplyr::filter(
      !is.na(os_time),
      !is.na(os_event),
      !is.na(.data[[v]])
    )
  
  fit <- coxph(
    as.formula(paste0("Surv(os_time, os_event) ~ `", v, "`")),
    data = df
  )
  
  broom::tidy(fit, exponentiate = TRUE, conf.int = TRUE) %>%
    mutate(variable = v,
           n = nrow(df),
           events = sum(df$os_event))
})

uni_plus_results$term <- gsub("`", "", uni_plus_results$term)
uni_plus_results$term <- gsub("StageStage", "Stage", uni_plus_results$term)

#________________

uni_plus_results_plt <- 
  uni_plus_results %>%
  ggplot(aes(x = estimate, y = reorder(term, estimate), color = p.value)) +
  geom_point(size = 2) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  geom_vline(xintercept = 1, linetype = "dashed") +
  scale_x_log10() +
  scale_color_viridis_c() +
  labs(x = "Hazard Ratio (log scale)", y = NULL,
       subtitle = "Univariate Cox HR\nSingscore: Upregulated Genes and CD4 EM\nSignature − Downregulated Genes", color = "Pvalue", tag = "l") +
  theme_bw(base_size = 7) +
  theme(
    text = element_text(size = 7),
    title = element_text(size = 7),
    plot.title = element_text(size = 7),
    plot.subtitle = element_text(size = 7, vjust = 0),
    axis.title.y = element_blank(),
    plot.tag = element_text(size = 7, face = "bold"),
    legend.title = element_text(
      size = 6,
      margin = margin(b = 10)
    ),
    legend.margin = margin(0, 0, 0, 0),
    legend.box.margin = margin(0, 0, 0, 5),
    legend.box.spacing = unit(0, "pt"),
    legend.key.width = unit(0.25, "cm"),
    legend.key.height = unit(0.3, "cm")
  )

```

##### KM plot

```{r}

# KM plot

# Median split
group_plus <- ifelse(cox_plus_df$`Gene-set Score` > median(cox_plus_df$`Gene-set Score`, na.rm = TRUE), "High", "Low")
group_plus <- factor(group_plus, levels = c("Low", "High"))

surv_plus_obj <- Surv(cox_plus_df$os_time, cox_plus_df$os_event)

km_plus_fit <- survfit(surv_plus_obj ~ group_plus)

ggsurv_plus <- ggsurvplot(
  km_plus_fit,
  data = data.frame(group_plus = group_plus),
  pval = TRUE,
  conf.int = FALSE,
  size = 0.45,
  censor.size = 2.3,
  palette = c("#4DBBD5", "#E64B35"),
  legend.labs = c("Low", "High"),
  title = "KM Survival Curve\nSingscore: Upregulated Genes and CD4 EM\nSignature − Downregulated Genes",
  xlab = "Time (months)",
  ylab = "Overall Survival Probability",
  pval.coord = c(15, 0.2),
  ggtheme = theme(
                  text = element_text(size = 7),
                  strip.text = element_text(size = 7),
                  title = element_text(size = 7),
                  plot.title = element_text(size = 7),
                  plot.subtitle = element_text(size = 7, vjust = 0),
                  plot.tag = element_text(size = 7, face = "bold"),
                  legend.title = element_text(
                    size = 6,
                    margin = margin(b = 10)
                  ),
                  legend.margin = margin(2, 2, 2, 2),
                  legend.box.margin = margin(2, 2, 2, 2),
                  legend.box.spacing = unit(1, "pt"),
                  legend.key.width = unit(0.5, "cm"),
                  legend.key.height = unit(0.5, "cm")
  )
)

ggsurv_plus$plot$layers[[ which(sapply(ggsurv_plus$plot$layers,function(x) inherits(x$geom, "GeomText"))) ]]$aes_params$size <- 2.5

ggsurv_plus_plt <- ggsurv_plus$plot
ggsurv_plus_plt <- ggsurv_plus_plt +
  theme_minimal() +
  labs(subtitle = NULL, 
       y = "Overall Survival Probability", 
       x = "Time (months)", 
       tag = "m") + 
  theme(
        text = element_text(size = 7),
        strip.text = element_text(size = 7),
        title = element_text(size = 7),
        plot.title = element_text(size = 7),
        plot.subtitle = element_text(size = 7, vjust = 0),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(
          size = 6,
          margin = margin(b = 10)
        ),
        legend.margin = margin(2, 2, 2, 2),
        legend.box.margin = margin(2, 2, 2, 2),
        legend.box.spacing = unit(1, "pt"),
        legend.key.width = unit(0.5, "cm"),
        legend.key.height = unit(0.5, "cm")
  )

```

### Fifth row figure

```{r}

# Second row figure
## Combine the plots of the subset
cd4_em_second_row_figure <- 
  uni_plus_results_plt + patchwork::plot_spacer() + ggsurv_plus_plt + 
       plot_layout(widths = c(1.2, 0.7, 1), heights = c(1), guides = "keep", axes = "keep") &
  theme(plot.margin = margin(t = 2, r = 5, b = 2, l = 5))

ggsave(filename = "../../../Testing results/cd4_em_second_row_figure.png", 
       plot = cd4_em_second_row_figure, device = "png", width = 6.85, height = (1.762)*1.5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/cd4_em_second_row_figure.pdf", 
       plot = cd4_em_second_row_figure, device = "pdf", width = 6.85, height = (1.762)*1.5, units = "in", dpi = 300)

```

# Combine all rows

```{r}

bcscm_t_nk_paper_subs_combined_rows <- 
  wrap_plots(wrap_elements(bcscm_first_row) & theme(plot.margin = margin(b = -50)),
             wrap_elements(bcscm_paperTypes_second_row + theme(plot.margin = margin(t = -50))), 
             wrap_elements(bcscm_third_row), 
             # wrap_elements(bcscm_conflicting_row),
             wrap_elements(cd4_em_second_row_figure),
             heights = c(1.4, 0.9, 1.3, 1.2)) & theme(plot.margin = margin(0, 0, 0, 0))

ggsave(filename = "../../../Testing results/bcscm_t_nk_paper_subs_combined_rows.pdf", 
       plot = bcscm_t_nk_paper_subs_combined_rows, device = "pdf", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

# ggsave(filename = "../../../Testing results/bcscm_t_nk_paper_subs_combined_rows.png", 
#        plot = bcscm_t_nk_paper_subs_combined_rows, device = "png", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

```
