---
title: "Evaluating ClustoCell's Performance in finding novel biology, validation in other dataset(s)"
author: "Adrian Salavaty"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(pROC)
library(umap)
library(cluster)
library(ggplot2)
library(dplyr)
library(splatter)
library(Seurat)
library(tidyverse)
library(scplotter)
library(scran)
library(scater)
library(Matrix)
library(SeuratDisk)
library(hdf5r)
library(patchwork)
library(paletteer)
library(hrbrthemes)
library(glmnet)
library(DescTools)
library(ggalign)
library(ggpubr)   
library(rmda)
library(data.table)
library(caret)
library(monocle3)
library(SeuratWrappers)
library(scales)

```

# Introduction

Script for evaluation of ClustoCell's Performance in finding novel biology, validation in other dataset(s).

# Function Defintion

We first define some helper functions.

```{r}

# Personalized Theme function

styled_theme <- function(
  style = c("academic", "minimal", "modern", "dark", "classic", "void"),
  border_mode = c("default", "all", "none"),
  title_align = c("center", "left", "right"),
  legends_to_drop = NULL,
  axes_title_drop = FALSE,
  legend_position = c("default", "top", "bottom", "left", "right", "inside"),
  axis_title_face = c("regular", "bold"),
  x_axis_text_angle = 45,
  x_axis_text_y = -0.3,
  remove_margin = c("none", "bottom", "all", "top", "left", "right")
) {
  style <- match.arg(style)
  border_mode <- match.arg(border_mode)
  title_align <- match.arg(title_align)
  legend_position <- match.arg(legend_position)
  axis_title_face <- match.arg(axis_title_face)
  remove_margin <- match.arg(remove_margin)

  base_font <- "Helvetica"
  align_map <- c(left = 0, center = 0.5, right = 1)
  hjust_val <- align_map[title_align]
  thin_size <- 0.3  

  # Decide axis title face
  axis_face_val <- ifelse(axis_title_face == "bold", "bold", "plain")

  # Base theme selection
  if (style == "void") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        axis.title.x = element_blank(),  
        axis.title.y = element_blank(),  
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "academic") {
    theme_out <- theme_minimal(base_size = 14, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = hjust_val),
        plot.subtitle = element_text(size = 13, hjust = hjust_val, color = "grey30"),
        plot.caption = element_text(size = 10, hjust = 1, color = "grey40"),
        axis.title = element_text(face = axis_face_val, size = 13),
        axis.text = element_text(size = 12, color = "black"),
        panel.grid.major = element_line(color = "grey80", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "minimal") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", size = thin_size),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.text = element_text(color = "black"),
        axis.title = element_text(face = axis_face_val),
        legend.position = "none",
        legend.title = element_text(face = "italic")
      )
  } else if (style == "modern") {
    theme_out <- theme_light(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.border = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "dark") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        plot.background = element_rect(fill = "black", color = NA),
        panel.background = element_rect(fill = "black", color = NA),
        panel.grid.major = element_line(color = "grey30", size = thin_size),
        panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 15, color = "white", hjust = hjust_val),
        plot.subtitle = element_text(color = "grey80", hjust = hjust_val),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white", face = axis_face_val),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white", face = "italic"),
        axis.line = element_line(color = "white", size = thin_size)
      )
  } else if (style == "classic") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  }

  # Apply border mode
  if (border_mode == "all") {
    theme_out <- theme_out + theme(panel.border = element_rect(color = "black", fill = NA, size = thin_size))
  } else if (border_mode == "none") {
    theme_out <- theme_out + theme(panel.border = element_blank(), axis.line = element_blank())
  }

  # Drop axes titles if requested
  if (axes_title_drop) {
    theme_out <- theme_out + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
  }

  # Override x-axis text angle and vertical position
  theme_out <- theme_out + theme(
    axis.text.x = element_text(
      angle = x_axis_text_angle,
      vjust = 1,
      hjust = 1,
      margin = margin(t = x_axis_text_y)
    )
  )

  # Handle margin removal
  if (remove_margin == "all") {
    theme_out <- theme_out + theme(plot.margin = margin(0, 0, 0, 0))
  } else if (remove_margin == "top") {
    theme_out <- theme_out + theme(plot.margin = margin(t = 0))
  } else if (remove_margin == "bottom") {
    theme_out <- theme_out + theme(plot.margin = margin(b = 0))
  } else if (remove_margin == "left") {
    theme_out <- theme_out + theme(plot.margin = margin(l = 0))
  } else if (remove_margin == "right") {
    theme_out <- theme_out + theme(plot.margin = margin(r = 0))
  }
  # if "none", do nothing

  # Legend positioning override
  if (legend_position != "default") {
    if (legend_position == "inside") {
      theme_out <- theme_out + theme(
        legend.position = c(0.98, 0.98),
        legend.justification = c("right", "top"),
        legend.direction = "vertical"
      )
    } else {
      horiz <- legend_position %in% c("top", "bottom")
      theme_out <- theme_out + theme(
        legend.position = legend_position,
        legend.direction = if (horiz) "horizontal" else "vertical"
      )
    }
  }

  # Return theme + guides if needed
  if (!is.null(legends_to_drop)) {
    drop_list <- lapply(legends_to_drop, function(x) "none")
    names(drop_list) <- legends_to_drop
    return(list(theme_out, guides(.list = drop_list)))
  } else {
    return(theme_out)
  }
}

#____________

# Label plot
label_plt <- function(label = "Label", size = 1) {
  ggplot(data.frame(x = 1, y = 1), aes(x = x, y = y)) +
    # Draw a single vertical bar
    geom_col(width = 0.6, fill = "lightgrey", color = "black") +
    # Add vertical text (angle=90)
    annotate("text", x = 1, y = 0.5, label = label, size = size,
             angle = 90, size = 6, vjust = 0.5, hjust = 0.5) +
    # Remove all extra space
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme_void(base_size = 14) +             # no axes, ticks, labels
    theme(
        plot.margin = margin(0, 0, 0, 0),      # remove outside margins
        panel.spacing = unit(0, "pt"),         # remove panel padding
        plot.background = element_blank(),     # no background
        panel.background = element_blank()
    )
}

#__________________________________________________________________

# Box plot with stats

make_facet_boxplot_with_stats <- function(
    df,
    x = "group",
    y = "value",
    fill = NULL,
    palette = NULL,
    facet.by = NULL,
    scales = "free_y",
    facet.nrow = NULL,       
    facet.ncol = NULL,      
    alpha = 0.8,
    xlab = NULL,
    ylab = NULL,
    subtitle = NULL,
    tag = NULL,
    jitter.width = 0.15,
    jitter.size = 0.35,
    jitter.alpha = 0.6,
    method = "wilcox.test",
    p.adjust.method = "none",
    hide.ns = FALSE,
    base_size = 10
) {

  library(dplyr)
  library(ggpubr)
  library(rstatix)
  library(ggplot2)

  # ----------------------------
  # 1. Statistical test per facet
  # ----------------------------
  stat.test <- df %>%
    group_by(!!sym(facet.by)) %>%
    rstatix::wilcox_test(as.formula(paste(y, "~", x))) %>%
    adjust_pvalue(method = p.adjust.method) %>%
    add_significance()

  # Add xy positions (critical for facetted layout)
  stat.test <- stat.test %>% add_xy_position(x = x)

  # ----------------------------
  # 2. Build ggpubr boxplot
  # ----------------------------
  p <- ggpubr::ggboxplot(
    df,
    x = x,
    y = y,
    fill = fill,
    alpha = alpha,
    palette = palette,
    facet.by = facet.by,
    scales = scales,
    nrow = facet.nrow,  
    ncol = facet.ncol,   
    outlier.shape = NA
  ) +
    geom_jitter(
      width = jitter.width,
      size = jitter.size,
      alpha = jitter.alpha
    ) +
    labs(
      x = xlab,
      y = ylab,
      subtitle = subtitle,
      tag = tag
    ) +
    theme_classic(base_size = base_size) +
    theme(
      plot.subtitle = element_text(size = 7),
      text = element_text(size = 6),
      strip.text = element_text(size = 5.5),
      plot.tag = element_text(size = 7, face = "bold"),
      axis.title.y = element_text(size = 7),
      plot.margin = margin(0, 5, 0, 5),
      legend.position = "none"
    )

  # ----------------------------------
  # 3. Add p-value stars to the plot
  # ----------------------------------
  p <- p + stat_pvalue_manual(
    stat.test,
    label = "p.adj.signif",
    hide.ns = hide.ns,
    size = 2.2,
    tip.length = 0.01
  )

  return(p)
}

```


```{r}

# Concordance = 100*Sum(rowMax)/N

library(paletteer)

paste(as.vector(col2rgb("#0000CCFF")), collapse = " ")

# https://pmassicotte.github.io/paletteer_gallery/

paletteer::paletteer_d("MetBrewer::Austria", 7)
#A40000FF #16317DFF #007E2FFF #FFCD12FF #B86092FF #721B3EFF #00B7A7FF 

plotthis::palette_list %>% names()

# Best pals from ggsci: pal_npg (10), pal_d3 (10), pal_ucscgb (26), pal_igv (51)

# "#FF0000FF" "#FF9900FF" "#FFCC00FF" "#00FF00FF" "#6699FFFF" "#CC33FFFF" "#99991EFF" "#999999FF" "#FF00CCFF" "#CC0000FF" "#FFCCCCFF" "#FFFF00FF" "#CCFF00FF" "#358000FF" "#0000CCFF" "#99CCFFFF" "#00FFFFFF"
# "#CCFFFFFF" "#9900CCFF" "#CC99FFFF" "#996600FF" "#666600FF" "#666666FF" "#CCCCCCFF" "#79CC3DFF" "#CCCC99FF"

# "#1F77B4FF" "#FF7F0EFF" "#2CA02CFF" "#D62728FF" "#9467BDFF" "#8C564BFF" "#E377C2FF" "#7F7F7FFF" "#BCBD22FF" "#17BECFFF"

# NPG colors:
# "#E64B35FF" "#4DBBD5FF" "#00A087FF" "#3C5488FF" "#F39B7FFF" "#8491B4FF" "#91D1C2FF" "#DC0000FF" "#7E6148FF" "#B09C85FF"

# "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF" "#924822FF" "#837B8DFF" "#C75127FF" "#D58F5CFF" "#7A65A5FF" "#E4AF69FF" "#3B1B53FF"
# "#CDDEB7FF" "#612A79FF" "#AE1F63FF" "#E7C76FFF" "#5A655EFF" "#CC9900FF" "#99CC00FF" "#A9A9A9FF" "#CC9900FF" "#99CC00FF" "#33CC00FF" "#00CC33FF" "#00CC99FF" "#0099CCFF" "#0A47FFFF" "#4775FFFF" "#FFC20AFF"
# "#FFD147FF" "#990033FF" "#991A00FF" "#996600FF" "#809900FF" "#339900FF" "#00991AFF" "#009966FF" "#008099FF" "#003399FF" "#1A0099FF" "#660099FF" "#990080FF" "#D60047FF" "#FF1463FF" "#00D68FFF"

 # "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF"

# Define the color palettes

npg_pal <- ggsci::pal_npg()(10)
npg_reverse_pal <- rev(ggsci::pal_npg("nrc")(10))

```

A function for converting count df to sparse matrix

```{r}

library(Rcpp)
library(RcppArmadillo)

Rcpp::sourceCpp(code = '
#define ARMA_64BIT_WORD 1
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

// [[Rcpp::export]]
Rcpp::S4 denseToSparse(Rcpp::DataFrame& df) {  // Pass by reference to allow in-place modification
  int nrows = df.nrows();
  int ncols = df.size();

  arma::sp_mat sparse(nrows, ncols);

  for (int col = 0; col < ncols; ++col) {
    arma::vec this_col = Rcpp::as<arma::vec>(df[col]);  // Wrap without copying
    sparse.col(col) = this_col;  // Assigns only non-zeros to sparse
    df[col] = R_NilValue;  // Free the column memory progressively
  }

  return Rcpp::wrap(sparse);  // Returns as dgCMatrix S4 object
}
')

```

Concordance = 100*Sum(rowMax)/N

# Basal Cell Carcinoma

The data was downloaded from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE123813
https://pubmed.ncbi.nlm.nih.gov/31359002/

```{r}

bcc_count <- fread("../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/GSE123813_bcc_scRNA_counts.txt.gz")
bcc_count <- bcc_count %>% column_to_rownames("V1")

bcc_tcell_metadata <- fread("../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/GSE123813_bcc_tcell_metadata.txt.gz")
all(bcc_tcell_metadata$cell.id %in% colnames(bcc_count)) # TRUE

bcc_tcell_count <- bcc_count[,bcc_tcell_metadata$cell.id]
all(colnames(bcc_tcell_count) == bcc_tcell_metadata$cell.id) # TRUE
rm(bcc_count)

#________________________

# Preparing the clonal expansion labels

bcc_tcr_metadata <- fread("../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/GSE123813_bcc_tcr.txt.gz")
bcc_tcr_metadata[, clone_id := paste0(cdr3s_aa)]

# retaining only the ones that are present in the metadata
bcc_tcr_metadata <- bcc_tcr_metadata[which(bcc_tcr_metadata$V1 %in% bcc_tcell_metadata$cell.id),]
bcc_tcr_metadata$cell.id <- bcc_tcr_metadata$V1

bcc_tcr_metadata <- left_join(bcc_tcr_metadata, bcc_tcell_metadata, by = "cell.id")

# clone abundance per patient per timepoint
bcc_tcr_clone_counts <- bcc_tcr_metadata %>%
  group_by(patient, treatment, clone_id) %>%
  summarize(n = n(), .groups = "drop")

# total T-cells per patient per timepoint
bcc_tcr_clone_totals <- bcc_tcell_metadata %>%
  dplyr::filter(cell.id %in% bcc_tcr_metadata$cell.id) %>% 
  group_by(patient, treatment) %>%
  summarize(N = n(), .groups = "drop")

bcc_tcr_clone_counts <- bcc_tcr_clone_counts %>% 
  left_join(bcc_tcr_clone_totals, by=c("patient","treatment"))

bcc_tcr_clone_counts_wide <- bcc_tcr_clone_counts %>%
  pivot_wider(
    names_from = treatment,
    values_from = c(n, N),
    values_fill = 0
  )

# Apply Fisher’s exact test for each clone in each patient (as per the paper)
bcc_tcr_clone_counts_expand_table <- bcc_tcr_clone_counts_wide %>%
  rowwise() %>%
  mutate(
    fisher_p = fisher.test(
      matrix(
        c(n_pre, N_pre - n_pre,
          n_post, N_post - n_post),
        nrow = 2, byrow = TRUE
      )
    )$p.value,
    log2FC = log2((n_post / N_post + 1e-9) / (n_pre / N_pre + 1e-9)),
    expansion_status = case_when(
      fisher_p < 0.05 & log2FC > 0.5  ~ "Expanded",
      fisher_p < 0.05 & log2FC < -0.5 ~ "Contracted",
      n_pre > 0 & n_post > 0          ~ "Persistent",
      n_pre > 0 & n_post == 0         ~ "Pre_only",
      n_pre == 0 & n_post > 0         ~ "Post_only",
      TRUE                            ~ "Undetermined"
    )
  ) %>%
  ungroup()

# Assigning expansion labels back to each cell
bcc_tcr_clone_status <- bcc_tcr_clone_counts_expand_table %>%
  select(patient, clone_id, expansion_status)

bcc_tcell_metadata$clone_id <- "n/a"
bcc_tcell_metadata_clone_idx <- match(bcc_tcr_metadata$cell.id, bcc_tcell_metadata$cell.id)
bcc_tcell_metadata$clone_id[bcc_tcell_metadata_clone_idx] <- bcc_tcr_metadata$clone_id

bcc_tcell_metadata <- bcc_tcell_metadata %>%
  left_join(bcc_tcr_clone_status, by=c("patient","clone_id"))

bcc_tcell_metadata$expansion_status[is.na(bcc_tcell_metadata$expansion_status)] <- "n/a"

# Cell-level label: “expansion”
bcc_tcell_metadata$expansion_cell <- ifelse(
  bcc_tcell_metadata$expansion_status == "Expanded", "Expanded", "Non-expanded"
)

# Patient-level label: “E vs NE”
bcc_patient_expansion <- bcc_tcr_clone_counts_expand_table %>%
  group_by(patient) %>%
  summarize(
    has_expanded_clone = any(expansion_status == "Expanded")
  ) %>%
  mutate(expansion_group = ifelse(has_expanded_clone, "E", "NE"))

bcc_tcell_metadata <- left_join(
  bcc_tcell_metadata,
  bcc_patient_expansion,
  by="patient"
)

# bcc_tcell_so@meta.data <- cbind(bcc_tcell_so@meta.data, bcc_tcell_metadata[,c("clone_id", "expansion_status", "expansion_cell", "has_expanded_clone", "expansion_group")])

#________________________

bcc_tcell_so <- CreateSeuratObject(bcc_tcell_count, meta.data = bcc_tcell_metadata, min.cells = 3, min.features = 200)
bcc_tcell_so <- NormalizeData(bcc_tcell_so, normalization.method = "LogNormalize", scale.factor = 10000)

qs2::qs_save(object = bcc_tcell_so, 
             file = "../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/bcc_tcell_so.qs2")

bcc_tcell_so <- qs2::qs_read("../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/bcc_tcell_so.qs2")

#____________

bcc_tcell_so_ClustoCell <- clustoCell(data = bcc_tcell_so, 
                                            high_quantile = 0.25, low_quantile = 0.25, 
                                            subset_to_HVG = F, seed = 121)

bcc_tcell_so <- addClustoData(obj = bcc_tcell_so, clustoCell = bcc_tcell_so_ClustoCell, 
                                    major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

qs2::qs_save(object = bcc_tcell_so_ClustoCell, 
             file = "../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/bcc_tcell_so_ClustoCell.qs2")

bcc_tcell_so_ClustoCell <- qs2::qs_read("../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/bcc_tcell_so_ClustoCell.qs2")

#__________________

# Finding the dataset-level markers and using them for dim reduction
bcc_tcell_markers <- getDatasetMarkers(obj = bcc_tcell_so_ClustoCell, 
                                             clusters = T, sub_clusters = T,
                                             positive_markers = T, negative_markers = T, 
                                             medium_markers = F, 
                                             thresh_mode = "n", pos_thresh = 100, neg_thresh = 25)

bcc_tcell_so <- ScaleData(bcc_tcell_so, features = bcc_tcell_markers$combined_markers)
bcc_tcell_so <- RunPCA(bcc_tcell_so, features = bcc_tcell_markers$combined_markers)
ElbowPlot(bcc_tcell_so, ndims = 50)
# bcc_tcell_so <- FindNeighbors(bcc_tcell_so, dims = 1:20)
# bcc_tcell_so <- FindClusters(bcc_tcell_so, resolution = 0.5)
bcc_tcell_so <- RunUMAP(bcc_tcell_so, dims = 1:20)

qs2::qs_save(object = bcc_tcell_so, 
             file = "../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/bcc_tcell_so.qs2")

bcc_tcell_so <- qs2::qs_read("../../../Benchmarking/Test Datasets/Basal-Squamous Cell Carcinoma/bcc_tcell_so.qs2")

```

### Inspecting the data

Checking the metadata and confusion tables

```{r}

head(bcc_tcell_so@meta.data %>% dplyr::select(-cell.id))

# ClustoCell clusters and subclusters
table(bcc_tcell_so$ClustoCell)
table(bcc_tcell_so$SubClustoCell)

# Author annotations
table(bcc_tcell_so@meta.data$cluster)
table(bcc_tcell_so@meta.data$treatment)
table(bcc_tcell_so@meta.data$expansion_group)

# Confusion of author annotations with ClustoCell
table(bcc_tcell_so$cluster, bcc_tcell_so$ClustoCell)
table(bcc_tcell_so$ClustoCell, bcc_tcell_so$treatment)
table(bcc_tcell_so$ClustoCell, bcc_tcell_so$expansion_group)

# Confusion of author annotations with SubClustoCell
table(bcc_tcell_so$cluster, bcc_tcell_so$SubClustoCell)
table(bcc_tcell_so$SubClustoCell, bcc_tcell_so$treatment)
table(bcc_tcell_so$SubClustoCell, bcc_tcell_so$expansion_group)

```

Checking the markers of clusters and subclusters

```{r}

# Major Clusters
sapply(names(bcc_tcell_so_ClustoCell$markers$major_clusters$cluster_specific$positive_markers), 
       function(i) { 
         tmp_markers <- bcc_tcell_so_ClustoCell$markers$major_clusters$cluster_specific$positive_markers[[i]] %>% 
           dplyr::slice(1:30) %>% dplyr::pull(Feature)
         cat(paste0(i, ": "))
         cat(paste0(tmp_markers, collapse = ", "))
         cat("\n\n")
       }
)

# Sub-clusters
sapply(names(bcc_tcell_so_ClustoCell$markers$sub_clusters), 
       function(i) { 
         cat(paste0(i, ": "))
         cat("\n")
         
         sapply(names(bcc_tcell_so_ClustoCell$markers$sub_clusters[[i]]$positive_markers), function(j) {
           tmp_markers <- (bcc_tcell_so_ClustoCell$markers$sub_clusters[[i]]$positive_markers)[[j]] %>% 
             dplyr::slice(1:60) %>% dplyr::pull(Feature)
           cat(paste0(j, ": "))
           cat(paste0(tmp_markers, collapse = ", "))
           cat("\n\n")
         })
         cat("\n\n\n")
       }
)

#_________________________

# Medium Markers

# Major Clusters
sapply(names(bcc_tcell_so_ClustoCell$markers$major_clusters$cluster_specific$medium_markers), 
       function(i) { 
         tmp_markers <- bcc_tcell_so_ClustoCell$markers$major_clusters$cluster_specific$medium_markers[[i]] %>% 
           dplyr::slice(1:30) %>% dplyr::pull(Feature)
         cat(paste0(i, ": "))
         cat(paste0(tmp_markers, collapse = ", "))
         cat("\n\n")
       }
)

# Sub-clusters
sapply(names(bcc_tcell_so_ClustoCell$markers$sub_clusters), 
       function(i) { 
         cat(paste0(i, ": "))
         cat("\n")
         
         sapply(names(bcc_tcell_so_ClustoCell$markers$sub_clusters[[i]]$medium_markers), function(j) {
           if(!inherits((bcc_tcell_so_ClustoCell$markers$sub_clusters[[i]]$medium_markers)[[j]], "logMessage")) {
             tmp_markers <- (bcc_tcell_so_ClustoCell$markers$sub_clusters[[i]]$medium_markers)[[j]] %>% 
             dplyr::slice(1:30) %>% dplyr::pull(Feature)
             cat(paste0(j, ": "))
             cat(paste0(tmp_markers, collapse = ", "))
           } else {
             cat(paste0(j, ": No Medium Marker"))
           }
           cat("\n\n")
         })
         cat("\n\n\n")
       }
)

```

## Data exploration and clustering

```{r}
# Subsetting pre-treatment cells with known expansion labels
bcc_tcell_so@meta.data <- bcc_tcell_so@meta.data %>% 
  dplyr::mutate(timepoint = treatment,
         expansion = expansion_group)

bcc_tcell_pre <- subset(bcc_tcell_so, subset = timepoint == "pre" & expansion %in% c("E","NE"))

library(pROC)
library(glmnet)
library(caret)
library(pheatmap)

```

#### UMAP

```{r}

out_dir <- "../../../Testing results/bcc/Final/"

# UMAP of SubClustoCells

bcc_tcell_umap <- CellDimPlot(
  bcc_tcell_so,
  reduction = "umap",
  group_by = "SubClustoCell",
  label = FALSE,
  pt_size = 1.01,
  theme = "theme_blank",
  xlab = "",
  ylab = "",
  raster = TRUE,
  legend.position = "right",
  legend.direction = "vertical",
  palette = "igv"
) +
  coord_fixed(clip = "off", expand = FALSE) +
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    plot.margin = margin(t = 0, r = 0, b = 5, l = -5),
    plot.subtitle = element_text(size = 7, hjust = 0.05),
    legend.margin = margin(t = 0, r = 0, b = 5, l = -5),
    legend.box.margin = margin(0, 0, 0, 5),
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  ) +
  labs(
    subtitle = "BCC T Data SubClustoCells",
    color = NULL,
    x = NULL,
    y = NULL,
    tag = "a"
  ) +
  guides(color = guide_legend(ncol = 2, override.aes = list(size = 2)))

```

#### Confusion matrices

```{r}

# strip on y-axis for SubClustoCell
make_subclust_strip_df <- function(df, subclust_col = "SubClustoCell") {
  df |>
    dplyr::distinct(.data[[subclust_col]]) |>
    dplyr::rename(SubClustoCell = !!subclust_col) |>
    dplyr::mutate(
      SubClustoCell = factor(SubClustoCell, levels = levels(bcc_tcell_pre$SubClustoCell)),
      color = ggsci::pal_igv()(length(levels(bcc_tcell_pre$SubClustoCell)))
    )
}

# Generic confusion plot function (SubClustoCell vs X)
make_subclust_confusion_plot <- function(tab, x_lab, tag_letter, fraction = TRUE, showY_text = TRUE, accuracy = 1) {
  
  if(fraction) {
    df <- prop.table(tab, margin = 1) %>% as.data.frame()
  } else {
    df <- tab %>% as.data.frame()
  }
  colnames(df) <- c("SubClustoCell", x_lab, "Fraction")

  strip_df <- make_subclust_strip_df(df)

  curr_plt <- 
  ggplot(df, aes_string(x = x_lab, y = "SubClustoCell", color = "Fraction")) +
    geom_point(size = if(fraction) {df$Fraction } else {log10(df$Fraction) / 3.5}) +
    scale_color_gradient(
      low = "white",
      high = scales::muted("#DC0000FF"),
      labels = label_number(scale_cut = cut_short_scale(), accuracy = accuracy)
    ) +
    ggnewscale::new_scale_fill() +
    geom_tile(
      data = strip_df,
      aes(y = SubClustoCell, x = -0.5, fill = color),
      inherit.aes = FALSE,
      height = 0.75
    ) +
    scale_fill_identity(guide = "none") +
    scale_y_discrete(expand = expansion(add = 0.6)) +
    coord_cartesian(clip = "off") +
    theme_minimal() +
    labs(title = NULL, tag = tag_letter, subtitle = "Pre-treatment") +
    styled_theme(
      legend_position = "right",
      style = "classic",
      title_align = "left",
      border_mode = "all",
      legends_to_drop = "size"
    ) +
    theme(
      text = element_text(size = 7),
      axis.title.y = element_blank(),
      axis.text.y = element_text(size = 5.8),
      # axis.ticks.y = element_blank(),
      legend.margin = margin(2, 2, 2, 2),
      legend.box.margin = margin(t = 0, r = 0, b = 0, l = 5),
      legend.box.spacing = unit(1, "pt"),
      title = element_text(size = 7),
      axis.title = element_text(size = 7),
      plot.tag = element_text(size = 7, face = "bold"),
      legend.title = element_text(size = 5.5, margin = margin(b = 2)),
      legend.key.width = unit(0.25, "cm"),
      legend.key.height = unit(0.5, "cm")
    ) +
    guides(
      color = guide_colorbar(
        barheight = unit(1.5, "cm"),
        barwidth  = unit(0.2, "cm")
      )
    )
  
  if(!showY_text) {
    curr_plt <- curr_plt + 
      theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),

      )
  }
  
  if(!fraction) {
        curr_plt <- curr_plt + labs(color = "Count")
  }
  
  curr_plt
}

#_________________

# SubClustoCell vs author clusters
bcc_SubClustoCell_vs_cluster_tab <- table(
  SubClustoCell = bcc_tcell_pre$SubClustoCell,
  cluster       = bcc_tcell_pre$cluster
)

bcc_SubClustoCell_vs_cluster_plt <-
  make_subclust_confusion_plot(
    bcc_SubClustoCell_vs_cluster_tab,
    x_lab      = "cluster",
    tag_letter = "b", accuracy = 0.1, 
  ) + labs(x = "Manual Annotations") +
  theme(panel.grid.major.y = element_line(color = "grey", size = 0.3))

# SubClustoCell vs expansion_group (E vs NE)
bcc_SubClustoCell_vs_expansion_tab <- table(
  SubClustoCell   = bcc_tcell_pre$SubClustoCell,
  expansion_group = bcc_tcell_pre$expansion_group
)

bcc_SubClustoCell_vs_expansion_plt <-
  make_subclust_confusion_plot(
    bcc_SubClustoCell_vs_expansion_tab,
    x_lab      = "expansion_group",
    tag_letter = "c", fraction = F, showY_text = F
  ) + labs(x = "Expansion")

```

## First-row figure

```{r}

bcc_first_ext_row <- (
    bcc_tcell_umap +
    ((bcc_SubClustoCell_vs_cluster_plt +
        bcc_SubClustoCell_vs_expansion_plt) +
       plot_layout(widths = c(9, 2), heights = 1, guides = "keep", axes = "collect_y"))
) +
  plot_layout(widths = c(13, 23), heights = 1, guides = "keep") & theme(plot.margin = margin(t = 0, r = 5, b = 0, l = 5))

ggsave(
  filename = file.path(out_dir, "bcc_first_ext_row.png"),
  plot     = bcc_first_ext_row,
  device   = "png",
  width    = 6.85,
  height   = (1.762) * 1.3,
  units    = "in",
  dpi      = 300
)

ggsave(
  filename = file.path(out_dir, "bcc_first_ext_row.pdf"),
  plot     = bcc_first_ext_row,
  device   = "pdf",
  width    = 6.85,
  height   = (1.762) * 1.5,
  units    = "in"
)

```

## Evaluating the top sub-clusters

```{r}

library(pROC)
library(glmnet)

bcc_subClust_Exp_Type_df <- bcc_tcell_pre@meta.data[, c("SubClustoCell", "expansion", "cluster")] %>% 
  mutate(expansion = ifelse(expansion == "E", 1, 0))

bcc_subClust_Exp_Type_df$SubClustoCell <- as.character(bcc_subClust_Exp_Type_df$SubClustoCell)

expansion_enriched_subs <- 
  bcc_subClust_Exp_Type_df %>%
  group_by(SubClustoCell) %>%
  summarise(
    n = n(),
    prop = mean(expansion)
  ) %>% 
  dplyr::filter(prop > mean(bcc_subClust_Exp_Type_df$expansion)) %>% 
  dplyr::pull(SubClustoCell) %>% as.character()

bcc_subClust_Exp_Type_df <- bcc_subClust_Exp_Type_df %>% dplyr::filter(SubClustoCell %in% expansion_enriched_subs)

#_________________

# Fisher exact test

bcc_SubClustoCells <- unique(bcc_subClust_Exp_Type_df$SubClustoCell)

set.seed(1234)
bcc_top_SubClustoCells <- lapply(bcc_SubClustoCells, function(c){
  sub_df <- data.frame(
    x = factor(bcc_subClust_Exp_Type_df$SubClustoCell == c, levels = c(TRUE, FALSE)),
    y = bcc_subClust_Exp_Type_df$expansion
  )
  test <- fisher.test(table(sub_df))
  data.frame(
    SubClustoCell = c,
    p_value = test$p.value,
    odds_ratio = test$estimate
  )
})

bcc_top_SubClustoCells <- do.call(rbind, bcc_top_SubClustoCells) %>% 
  dplyr::slice_min(p_value, n = 3) # Enough to get an AUC of 1

```

### Evaluating the response-ready circuit

Evaluating the C5-Sub1, C3-Sub2, C4-Sub3 among T cells, identified through exploration of the table of clusters and cell types/states and their confusions.

#### Data Prep.

```{r}

# Setup and metadata preparation

# Ensure key metadata are factors
bcc_tcell_pre$ClustoCell    <- factor(bcc_tcell_pre$ClustoCell)
bcc_tcell_pre$SubClustoCell <- factor(bcc_tcell_pre$SubClustoCell)
bcc_tcell_pre$cluster       <- factor(bcc_tcell_pre$cluster)     # author T cell cluster

bcc_tcell_pre$expansion_group <- factor(bcc_tcell_pre$expansion_group) # "E", "NE"
bcc_tcell_pre$treatment     <- factor(bcc_tcell_pre$treatment)        # "pre", "post"
bcc_tcell_pre$patient       <- factor(bcc_tcell_pre$patient)

```

### Predictive validity (pre-treatment, E vs NE)

  "C5-Sub1",          
  "C3-Sub2",    
  "C4-Sub3"

```{r}

# Predictive subclusters (Yost analogues of Bassez C4-Sub4 / C5-Sub2/3 / C6-Sub3)

bcc_predictive_subs <- c(
  "C5-Sub1",          
  "C3-Sub2",    
  "C4-Sub3"
)

# Per-patient summary of predictive subcluster fractions
patient_summary_bcc_pre <- bcc_tcell_pre@meta.data %>%
  filter(expansion_group %in% c("E", "NE")) %>%
  group_by(patient) %>%
  summarise(
    expansion_status = unique(expansion_group),
    n_cells          = n(),
    !!!setNames(
      lapply(bcc_predictive_subs, function(sc) {
        rlang::quo(mean(SubClustoCell == !!sc))
      }),
      paste0("frac_", gsub("-", "", bcc_predictive_subs))
    ),
    .groups = "drop"
  ) %>%
  mutate(
    expansion_binary = ifelse(expansion_status == "E", 1, 0)
  )

# Long format for boxplots
patient_summary_bcc_pre_long <- patient_summary_bcc_pre %>%
  pivot_longer(
    cols      = starts_with("frac_"),
    names_to  = "subcluster",
    values_to = "fraction"
  )

# Clean labels, e.g. frac_C2Sub1 -> C2-Sub1
patient_summary_bcc_pre_long$subcluster_label <-
  gsub("^frac_", "", patient_summary_bcc_pre_long$subcluster) |>
  gsub("^(C\\d+)(Sub\\d+)$", "\\1-\\2", x = _)

patient_summary_bcc_pre_long$subcluster_label <-
  factor(patient_summary_bcc_pre_long$subcluster_label,
         levels = unique(patient_summary_bcc_pre_long$subcluster_label))

# Boxplot: E vs NE
patient_summary_bcc_pre_boxplt <-
  make_facet_boxplot_with_stats(df = patient_summary_bcc_pre_long, 
                                                                  x = "expansion_status",
                                                                  y = "fraction",
                                                                  fill = "expansion_status", 
                                                                  palette = c("NE" = "gray80", "E" = "#D81B60"),
                                                                  facet.by = "subcluster_label",
                                                                  subtitle = "Pre-treatment Frequencies of\nPredictive ClustoCell Subclusters", 
                                                                  ylab = "Per−patient cell fraction",
                                                                  tag = "d")


```

### Elastic-net model (patient-level prediction of expansion)

```{r}

# Elastic-net logistic regression

X_bcc <- as.matrix(
  patient_summary_bcc_pre %>%
    mutate(log10_n_cells = log10(n_cells)) %>%
    dplyr::select(starts_with("frac_"), log10_n_cells)
)

y_bcc <- patient_summary_bcc_pre$expansion_binary

set.seed(1234)
cvfit_bcc <- cv.glmnet(
  X_bcc, y_bcc,
  family = "binomial",
  alpha  = 0.5,
  nfolds = 5
)

# Coefficients (excluding intercept)
coefs_bcc <- as.matrix(coef(cvfit_bcc, s = "lambda.min"))[-1, , drop = FALSE]
coef_df_bcc <- data.frame(
  term     = rownames(coefs_bcc),
  estimate = as.numeric(coefs_bcc[, 1]),
  row.names = NULL
)

# Human-readable labels
coef_df_bcc$term <- gsub("^frac_", "", coef_df_bcc$term)
coef_df_bcc$term <- gsub("^(C\\d+)(Sub\\d+)$", "\\1-\\2", coef_df_bcc$term)

coef_df_bcc$term[coef_df_bcc$term == "log10_n_cells"] <- "log10(total T cells)"

patient_summary_bcc_pre_focus_pred_direction_plt <-
  ggplot(coef_df_bcc, aes(x = reorder(term, estimate), y = estimate, fill = estimate > 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_manual(values = c("TRUE" = "#1E88E5", "FALSE" = "gray60")) +
  theme_minimal(base_size = 6) +
  labs(
    x     = NULL,
    y     = "Coefficient (elastic-net, λmin)",
    tag   = "e",
    title = "Predictive Directions"
  ) +
  theme(
    plot.subtitle   = element_text(size = 7),
    plot.title      = element_text(size = 7),
    legend.title    = element_blank(),
    text            = element_text(size = 6),
    axis.title.x    = element_text(size = 7),
    strip.text      = element_text(size = 5.5),
    plot.tag        = element_text(size = 7, face = "bold"),
    title           = element_text(size = 8),
    line            = element_line(linewidth = 0.1),
    plot.margin     = margin(0, 5, 0, 5),
    legend.margin   = margin(0, 0, 0, 0),
    legend.box.margin = margin(0, 0, 0, 0),
    legend.box.spacing = unit(0, "pt"),
    legend.key.size   = unit(0.2, "cm"),
    legend.text       = element_text(size = 5),
    legend.position   = "top",
    legend.direction  = "horizontal",
    legend.key.width  = unit(6, "pt"),
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )


```

### ROC and AUROC

```{r}

# ROC analysis

set.seed(1234)
pred_prob_bcc <- predict(
  cvfit_bcc,
  newx = X_bcc,
  s    = "lambda.min",
  type = "response"
)[, 1]

roc_obj_bcc <- roc(y_bcc, pred_prob_bcc)
auc_val_bcc <- auc(roc_obj_bcc)
cat("BCC AUC =", auc_val_bcc, "\n")

patient_summary_bcc_pre_roc_plt <-
  ggroc(roc_obj_bcc, color = "#E64B35", size = 0.5) +
  theme_classic(base_size = 6) +
  annotate("text", x = 0.6, y = 0.1,
           label = paste0("AUROC = ", round(auc_val_bcc, 3))) +
  labs(
    subtitle = "Elastic-net Model: Predictive Accuracy",
    x        = "Specificity",
    y        = "Sensitivity",
    tag      = "f"
  ) +
  theme(
    plot.subtitle   = element_text(size = 7),
    plot.title      = element_text(size = 7),
    legend.title    = element_blank(),
    text            = element_text(size = 6),
    strip.text      = element_text(size = 5.5),
    plot.tag        = element_text(size = 7, face = "bold"),
    title           = element_text(size = 8),
    line            = element_line(linewidth = 0.1),
    axis.title      = element_text(size = 7),
    plot.margin     = margin(0, 5, 0, 5),
    legend.margin   = margin(0, 0, 0, 0),
    legend.box.margin = margin(0, 0, 0, 0),
    legend.box.spacing = unit(0, "pt"),
    legend.key.size   = unit(0.2, "cm"),
    legend.text       = element_text(size = 5),
    legend.position   = "top",
    legend.direction  = "horizontal",
    legend.key.width  = unit(6, "pt"),
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  )


```

### Signature heatmap (predictive-state gene panels)

```{r}

# Predictive-state gene signatures: BCC T-cell

bcc_predictive_subs <- c(
  "C5-Sub1",          
  "C3-Sub2",    
  "C4-Sub3"
)

# Signature panels:
sig_panels_bcc_tcell <- list(
  `C5-Sub1` = c("CD8A", "CD8B", "GZMB", "GZMK", "GNLY", "NKG7", "SLC7A5"), 
  `C3-Sub2` = c("FOXP3", "CTLA4", "IL2RA", "BATF", "TNFRSF4", "TNFRSF18", "CD27", "LTB"),
  `C4-Sub3` = c("RORA", "KLRB1", "BATF", "CXCL13", "BHLHE40", "ID2", "CXCR6", "TNFRSF18")
)

# Keep only genes present in this object
sig_genes_bcc <- unique(unlist(sig_panels_bcc_tcell))
sig_genes_bcc <- sig_genes_bcc[sig_genes_bcc %in% rownames(bcc_tcell_pre)]

# Average expression per SubClustoCell for these genes
sig_panels_bcc_avg_expr <- Seurat::AverageExpression(
  bcc_tcell_pre,
  group.by = "SubClustoCell",
  features = sig_genes_bcc
)$RNA

# Restrict to predictive subclusters (C5-Sub1, C3-Sub2, C4-Sub3)
sig_panels_bcc_avg_expr <- sig_panels_bcc_avg_expr[
  , colnames(sig_panels_bcc_avg_expr) %in% bcc_predictive_subs,
  drop = FALSE
]

# Ordering columns:
# TPEX-Prolif (C3-Sub2) -> TPEX-ISG (C4-Sub3) -> TEX-Cyto (C5-Sub1)
col_order_bcc <- intersect(c("C3-Sub2", "C4-Sub3", "C5-Sub1"),
                           colnames(sig_panels_bcc_avg_expr))
sig_panels_bcc_avg_expr <- sig_panels_bcc_avg_expr[, col_order_bcc, drop = FALSE]

# Order genes by signature groups
gene_order_bcc <- unlist(sig_panels_bcc_tcell)
gene_order_bcc <- gene_order_bcc[gene_order_bcc %in% rownames(sig_panels_bcc_avg_expr)]

# Scale per gene
sig_panels_bcc_scaled <- t(scale(t(sig_panels_bcc_avg_expr[gene_order_bcc, , drop = FALSE])))

# Row annotation: which gene belongs to which signature
ann_row_bcc <- data.frame(
  Gene      = rownames(sig_panels_bcc_scaled),
  Signature = NA_character_,
  stringsAsFactors = FALSE
)

for (sig_name in names(sig_panels_bcc_tcell)) {
  genes_in_sig <- sig_panels_bcc_tcell[[sig_name]]
  ann_row_bcc$Signature[ann_row_bcc$Gene %in% genes_in_sig] <- sig_name
}

ann_row_bcc <- ann_row_bcc %>%
  dplyr::distinct(Gene, .keep_all = TRUE)

rownames(ann_row_bcc) <- ann_row_bcc$Gene

# Ensure order is consistent
sig_panels_bcc_scaled <- sig_panels_bcc_scaled[rownames(ann_row_bcc), , drop = FALSE]

ann_row_bcc$Signature <- factor(
  ann_row_bcc$Signature,
  levels = names(sig_panels_bcc_tcell)
)

# Column annotation: subclusters (C5-Sub1, C3-Sub2, C4-Sub3)
ann_col_bcc <- data.frame(
  SubClustoCell = colnames(sig_panels_bcc_scaled)
)
rownames(ann_col_bcc) <- ann_col_bcc$SubClustoCell

# Heatmap 
sig_panels_bcc_ggheatmap <-
  ggalign::ggheatmap(sig_panels_bcc_scaled) +
  scale_fill_gradient2(
    low      = "navy",
    mid      = "white",
    high     = "firebrick3",
    midpoint = 0,
    name     = "Scaled\nExpr."
  ) +
  labs(
    subtitle = "Predictive T-cell State Signatures\nMean Expression per Subcluster",
    tag      = "g"
  ) +
    theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 25, vjust = 0.7, face = "plain", color = "black"),
        axis.text = element_text(size = 6),
        text = element_text(size = 7),
        title = element_text(size = 7),
        plot.title = element_text(size = 7),
        plot.subtitle = element_text(size = 7, vjust = 0),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 6),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 0),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.4, "cm"))

```

### Signature heatmap (expanded)

```{r}

# Predictive-state gene signatures: BCC T-cell

bcc_predictive_subs_exp <- c(
  "C5-Sub1",          
  "C3-Sub2",    
  "C4-Sub3",
  "C7-Sub3", "C2-Sub2", "C2-Sub1" # Supporters
)

# Signature panels:
sig_panels_bcc_tcell_exp <- list(
  `CD8 cytotoxic/effector core`        = c("CD8A","CD8B","GZMB","GZMK","GNLY",
                                            "NKG7","SLC7A5","TBX21","IFNG"),
  `MHC-II / activation`                = c("HLA-DRA","HLA-DRB1","HLA-DRB5","NR4A1"),
  `CD8 effector-memory survival`       = c("BCL2A1","TCF7","CCR2","CCR5"),
  `Treg core + memory coat`            = c("FOXP3","CTLA4","IL2RA","TNFRSF4",
                                            "TNFRSF18","BATF","CD27","LTB","LEF1",
                                            "BCL2","BCL2L1","ICOS","MAF","GATA3"),
  `Th17 lineage core`                  = c("RORC","RORA","IL17A","IL17F","IL22",
                                            "IL23R","CCR6","KLRB1","BHLHE40","ID2"),
  `Th17-like helper module`            = c("CXCL13","CXCR6","CD40LG","IL21"),
  `Exhaustion / TFH exclusion`         = c("PDCD1","TIGIT","LAG3","TOX","ENTPD1",
                                            "CXCR5","BCL6")
)

# Keep only genes present in this object
sig_genes_bcc_exp <- unique(unlist(sig_panels_bcc_tcell_exp))
sig_genes_bcc_exp <- sig_genes_bcc_exp[sig_genes_bcc_exp %in% rownames(bcc_tcell_pre)]

# Average expression per SubClustoCell for these genes
sig_panels_bcc_avg_expr_exp <- Seurat::AverageExpression(
  bcc_tcell_pre,
  group.by = "SubClustoCell",
  features = sig_genes_bcc_exp
)$RNA

# Restrict to predictive subclusters (C5-Sub1, C3-Sub2, C4-Sub3)
sig_panels_bcc_avg_expr_exp <- sig_panels_bcc_avg_expr_exp[
  , colnames(sig_panels_bcc_avg_expr_exp) %in% bcc_predictive_subs_exp,
  drop = FALSE
]

# Ordering columns:
# TPEX-Prolif (C3-Sub2) -> TPEX-ISG (C4-Sub3) -> TEX-Cyto (C5-Sub1)
col_order_bcc_exp <- intersect(c("C3-Sub2", "C4-Sub3", "C5-Sub1", "C7-Sub3", "C2-Sub2", "C2-Sub1"),
                               colnames(sig_panels_bcc_avg_expr_exp))
sig_panels_bcc_avg_expr_exp <- sig_panels_bcc_avg_expr_exp[, col_order_bcc_exp, drop = FALSE]

# Order genes by signature groups
gene_order_bcc_exp <- unlist(sig_panels_bcc_tcell_exp)
gene_order_bcc_exp <- gene_order_bcc_exp[gene_order_bcc_exp %in% rownames(sig_panels_bcc_avg_expr_exp)]

# Scale per gene
sig_panels_bcc_scaled_exp <- t(scale(t(sig_panels_bcc_avg_expr_exp[gene_order_bcc_exp, , drop = FALSE])))

# Row annotation: which gene belongs to which signature
ann_row_bcc_exp <- data.frame(
  Gene      = rownames(sig_panels_bcc_scaled_exp),
  Signature = NA_character_,
  stringsAsFactors = FALSE
)

for (sig_name in names(sig_panels_bcc_tcell_exp)) {
  genes_in_sig <- sig_panels_bcc_tcell_exp[[sig_name]]
  ann_row_bcc_exp$Signature[ann_row_bcc_exp$Gene %in% genes_in_sig] <- sig_name
}

ann_row_bcc_exp <- ann_row_bcc_exp %>%
  dplyr::distinct(Gene, .keep_all = TRUE)

rownames(ann_row_bcc_exp) <- ann_row_bcc_exp$Gene

# Ensure order is consistent
sig_panels_bcc_scaled_exp <- sig_panels_bcc_scaled_exp[rownames(ann_row_bcc_exp), , drop = FALSE]

ann_row_bcc_exp$Signature <- factor(
  ann_row_bcc_exp$Signature,
  levels = names(sig_panels_bcc_tcell_exp)
)

# Column annotation: subclusters (C5-Sub1, C3-Sub2, C4-Sub3)
ann_col_bcc_exp <- data.frame(
  SubClustoCell = colnames(sig_panels_bcc_scaled_exp)
)
rownames(ann_col_bcc_exp) <- ann_col_bcc_exp$SubClustoCell

# Heatmap 
sig_panels_bcc_ggheatmap_exp <-
  ggalign::ggheatmap(sig_panels_bcc_scaled_exp) +
  scale_fill_gradient2(
    low      = "navy",
    mid      = "white",
    high     = "firebrick3",
    midpoint = 0,
    name     = "Scaled\nExpr."
  ) +
  labs(
    subtitle = "Expanded canonical-marker panel across expansion-associated BCC subclusters\nand their up/down-stream pseudotime subclusters",
    tag      = NULL
  ) +
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0),
        axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 25, vjust = 0.7, face = "plain", color = "black"),
        axis.text = element_text(size = 6),
        text = element_text(size = 7),
        title = element_text(size = 7),
        plot.title = element_text(size = 7),
        plot.subtitle = element_text(size = 7, vjust = 0),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 6),
        legend.margin = margin(t = 0, r = 0, b = 0, l = 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 0),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.4, "cm"))

```

## Second row figure

```{r}

bcc_second_row <- (
    wrap_elements(patient_summary_bcc_pre_boxplt) +
    ((wrap_elements(patient_summary_bcc_pre_focus_pred_direction_plt) /
        patchwork::plot_spacer() /
        wrap_elements(patient_summary_bcc_pre_roc_plt)) +
       plot_layout(heights = c(13, 2, 14), widths = 1, guides = "keep") &
       theme(plot.margin = margin(0, 0, 0, 0))) +
    ggalign::ggalignGrob(sig_panels_bcc_ggheatmap)
) +
  plot_layout(widths = c(12, 10, 12), heights = 1, guides = "keep")

ggsave(
  filename = file.path(out_dir, "bcc_second_row.png"),
  plot     = bcc_second_row,
  device   = "png",
  width    = 6.85,
  height   = (1.762) * 1.25,
  units    = "in",
  dpi      = 300
)

ggsave(
  filename = file.path(out_dir, "bcc_second_row.pdf"),
  plot     = bcc_second_row,
  device   = "pdf",
  width    = 6.85,
  height   = (1.762) * 1.25,
  units    = "in"
)


```

### Mechanistic Trajectories

#### C5-Sub1 (TP-Prolif) -> C5-Sub3 (TEX-Cyto)

```{r}

bcc_cd8_cells <- subset(
  bcc_tcell_so,
  subset = grepl("CD8|Naive", cluster)
)

bcc_cd8_cds <- as.cell_data_set(bcc_cd8_cells)
bcc_cd8_cds <- monocle3::cluster_cells(bcc_cd8_cds)
bcc_cd8_cds <- monocle3::learn_graph(bcc_cd8_cds, use_partition = FALSE)

rowData(bcc_cd8_cds)$gene_name        <- rownames(bcc_cd8_cds)
rowData(bcc_cd8_cds)$gene_short_name  <- rowData(bcc_cd8_cds)$gene_name

# UMAP colored by subcluster
bcc_cd8_subclusters_plt <-
    DimPlot(bcc_cd8_cells, 
          group.by = "SubClustoCell", 
          cells.highlight = list(
            "C7-Sub3" = colnames(bcc_cd8_cells)[bcc_cd8_cells$SubClustoCell %in% c("C7-Sub3")],
            "C5-Sub1" = colnames(bcc_cd8_cells)[bcc_cd8_cells$SubClustoCell %in% c("C5-Sub1")],
            "C2-Sub2" = colnames(bcc_cd8_cells)[bcc_cd8_cells$SubClustoCell %in% c("C2-Sub2")]
            ),
          cols.highlight = list(
            "C7-Sub3" = ggsci::pal_igv()(32)[26], 
            "C5-Sub1" = ggsci::pal_igv()(32)[15],
            "C2-Sub2" = ggsci::pal_igv()(32)[5]
            ),
          sizes.highlight = 0.1,
          order = c("C7-Sub3", "C5-Sub1", "C2-Sub2"),
          # cols = setNames(ggsci::pal_igv()(31), levels(droplevels(bcc_cd8_cells$SubClustoCell))),
          pt.size = 0.1) +
  theme_classic(base_size = 7) +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    plot.tag = element_text(size = 7, face = "bold"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.subtitle = element_text(size = 6.5, hjust = 0),
    plot.title = element_text(size = 7, hjust = 0),
    legend.title = element_text(size = 6),
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 5.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-4, "pt"),
    # legend.key.height = unit(0.4, "cm"),
    # legend.key.width = unit(0.25, "cm"),
    legend.margin = margin(0, 0, 0, 0),
    plot.margin = margin(1, 1, 1, 1, "mm")
  ) +
  guides(color = guide_legend(ncol = 1, override.aes = list(size = 2))) +
  labs(subtitle = "CD8 Subset; Subclusters UMAP", title = NULL, tag = "h", color = NULL)

# Order cells along pseudotime (conceptually:
bcc_cd8_cds <- monocle3::order_cells(bcc_cd8_cds)

bcc_cd8_pseudotime_plt <-
  monocle3::plot_cells(
    bcc_cd8_cds,
    color_cells_by        = "pseudotime",
    show_trajectory_graph = FALSE,
    label_branch_points   = FALSE,
    label_groups_by_cluster = FALSE,
    label_leaves          = FALSE
  ) +
  scale_color_viridis_c(name = "Pseudotime", option = "magma") +
  labs(
    subtitle = "CD8 Axis Pseudotime:\nC5-Sub1 (CD8 cytotoxic Eff/Eff-Mem) ->\nC7-Sub3 (activated effector CD8) ->\nC2-Sub2(CD8 TEX)",
    colour   = "Pseudotime"
  ) +
  theme_classic(base_size = 7) +
  theme(
    axis.line        = element_blank(),
    axis.ticks       = element_blank(),
    axis.text        = element_blank(),
    axis.title       = element_blank(),
    plot.subtitle    = element_text(size = 6.5, hjust = 0),
    plot.title       = element_text(size = 7, hjust = 0),
    legend.title     = element_text(size = 6),
    legend.text      = element_text(size = 6),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.25, "cm"),
    legend.margin    = margin(0, 0, 0, 0),
    plot.margin      = margin(1, 1, 1, 1, "mm")
  )

```

#### CD4 axis C2-Sub1 (Tfh helper) -> C4-Sub3 (CXCL13+ Th17)

```{r}

# CD4 axis C2-Sub1 (Tfh helper) -> C4-Sub3 (CXCL13+ Th17)

bcc_cd4_cells <- subset(
  bcc_tcell_so,
  subset = !grepl("CD8|Naive", cluster)
)

bcc_cd4_cds <- as.cell_data_set(bcc_cd4_cells)
bcc_cd4_cds <- monocle3::cluster_cells(bcc_cd4_cds)
bcc_cd4_cds <- monocle3::learn_graph(bcc_cd4_cds, use_partition = FALSE)

rowData(bcc_cd4_cds)$gene_name       <- rownames(bcc_cd4_cds)
rowData(bcc_cd4_cds)$gene_short_name <- rowData(bcc_cd4_cds)$gene_name

bcc_cd4_subclusters_plt <-
    DimPlot(bcc_cd4_cells, 
          group.by = "SubClustoCell", 
          cells.highlight = list(
            "C4-Sub3" = colnames(bcc_cd4_cells)[bcc_cd4_cells$SubClustoCell %in% c("C4-Sub3")],
            "C2-Sub1" = colnames(bcc_cd4_cells)[bcc_cd4_cells$SubClustoCell %in% c("C2-Sub1")]
            ),
          cols.highlight = list(
            "C4-Sub3" = ggsci::pal_igv()(32)[14], 
            "C2-Sub1" = ggsci::pal_igv()(32)[4]
            ),
          sizes.highlight = 0.1,
          order = c("C4-Sub3", "C2-Sub1"),
          # cols = setNames(ggsci::pal_igv()(31), levels(droplevels(bcc_cd8_cells$SubClustoCell))),
          pt.size = 0.1) +
  theme_classic(base_size = 7) +
  theme(
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    plot.tag = element_text(size = 7, face = "bold"),
    axis.text = element_blank(),
    axis.title = element_blank(),
    plot.subtitle = element_text(size = 6.5, hjust = 0),
    plot.title = element_text(size = 7, hjust = 0),
    legend.title = element_text(size = 6),
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 5.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-4, "pt"),
    # legend.key.height = unit(0.4, "cm"),
    # legend.key.width = unit(0.25, "cm"),
    legend.margin = margin(0, 0, 0, 0),
    plot.margin = margin(1, 1, 1, 1, "mm")
  ) +
  guides(color = guide_legend(ncol = 1, override.aes = list(size = 2))) +
  labs(subtitle = "CD4 Subset; Subclusters UMAP", title = NULL, tag = "i", color = NULL)

# Pseudotime
bcc_cd4_cds <- monocle3::order_cells(bcc_cd4_cds)

bcc_cd4_pseudotime_plt <-
  monocle3::plot_cells(
    bcc_cd4_cds,
    color_cells_by        = "pseudotime",
    show_trajectory_graph = FALSE,
    label_branch_points   = FALSE,
    label_groups_by_cluster = FALSE,
    label_leaves          = FALSE
  ) +
  scale_color_viridis_c(name = "Pseudotime", option = "magma") +
  labs(
    subtitle = "CD4 Axis Pseudotime:\nTfh Progenitor -> CXCL13+ Th17",
    colour   = "Pseudotime"
  ) +
  theme_classic(base_size = 7) +
  theme(
    axis.line        = element_blank(),
    axis.ticks       = element_blank(),
    axis.text        = element_blank(),
    axis.title       = element_blank(),
    plot.subtitle    = element_text(size = 6.5, hjust = 0),
    plot.title       = element_text(size = 7, hjust = 0),
    legend.title     = element_text(size = 6),
    legend.text      = element_text(size = 6),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width  = unit(0.25, "cm"),
    legend.margin    = margin(0, 0, 0, 0),
    plot.margin      = margin(1, 1, 1, 1, "mm")
  )

```

## Third-row figure

```{r}

# combined CD8 + CD4 trajectories

bcc_third_row <-
    ((
      (bcc_cd8_subclusters_plt + patchwork::plot_spacer() +
        bcc_cd8_pseudotime_plt + plot_layout(widths = c(10, 1, 10), heights = 1, guides = "keep")) /
     (bcc_cd4_subclusters_plt + patchwork::plot_spacer() +
        bcc_cd4_pseudotime_plt + plot_layout(widths = c(10, 1, 10), heights = 1, guides = "keep"))
     ) + plot_layout(heights = c(2, 1.8), widths = 1, guides = "keep"))

ggsave(
  filename = file.path(out_dir, "bcc_third_row.png"),
  plot     = bcc_third_row,
  device   = "png",
  width    = 6.85,
  height   = (1.762) * 1.25,
  units    = "in",
  dpi      = 300
)

ggsave(
  filename = file.path(out_dir, "bcc_third_row.pdf"),
  plot     = bcc_third_row,
  device   = "pdf",
  width    = 6.85,
  height   = (1.762) * 1.25,
  units    = "in"
)


```

## Combining the plots for main figure

```{r}

bcc_resp_pred_validation_figure_combined <- wrap_plots(wrap_elements(bcc_first_ext_row),
                                                   wrap_elements(bcc_second_row), 
                                                   wrap_elements(bcc_third_row), 
                                                   ncol = 1,
                                                   heights = c(1.4, 1.1, 1.3)) & theme(plot.margin = margin(0, 0, 0, 0))

bcc_resp_pred_validation_figure_combined <- wrap_elements(bcc_resp_pred_validation_figure_combined) +
  # ggtitle("Supplementary Figure 24: Expansion-associated immune circuits and lineage trajectories in basal cell carcinoma") +
  theme(plot.title = element_text(size = 8))

ggsave(filename = "../../../Testing results/bcc_resp_pred_validation_figure_combined.pdf", 
       plot = bcc_resp_pred_validation_figure_combined, device = "pdf", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/bcc_resp_pred_validation_figure_combined.png", 
       plot = bcc_resp_pred_validation_figure_combined, device = "png", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)


```

## Saving the expanded marker panel

```{r}

sig_panels_bcc_ggheatmap_exp_final <- wrap_elements(ggalign::ggalignGrob(sig_panels_bcc_ggheatmap_exp)) +
  # ggtitle("Supplementary Figure 25: Expanded canonical-marker panel resolves the cell-state identity of \nthe three expansion-enriched predictive subclusters and supports their refined annotation.") +
  theme(plot.title = element_text(size = 8))

ggsave(filename = "../../../Testing results/Supplementary Figure 23.pdf", 
       plot = sig_panels_bcc_ggheatmap_exp_final, device = "pdf", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/Supplementary Figure 23.png", 
       plot = sig_panels_bcc_ggheatmap_exp_final, device = "png", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)


```