---
title: "Evaluating ClustoCell's Performance for feature extraction for dimension reduction"
author: "Adrian Salavaty"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}

Sys.setenv(
  RETICULATE_PYTHON = "/Users/salavatyadrian/miniforge3/envs/scanpy_env/bin/python",
  RETICULATE_AUTOCONFIGURE = "FALSE",
  OMP_NUM_THREADS = "1",
  MKL_NUM_THREADS = "1",
  VECLIB_MAXIMUM_THREADS = "1",
  OPENBLAS_NUM_THREADS = "1"
)

library(reticulate)

py_config()

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(pROC)
library(umap)
library(cluster)
library(ggplot2)
library(dplyr)
library(splatter)
library(Seurat)
library(tidyverse)
library(scplotter)
library(scran)
library(scater)
library(SingleCellExperiment)
library(patchwork)
library(ggpubr)
library(scales)
library(PMCMRplus)  # For Friedman post-hoc
library(reshape2)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(pROC)
library(umap)
library(cluster)
library(ggplot2)
library(dplyr)
library(splatter)
library(Seurat)
library(tidyverse)
library(scplotter)
library(scran)
library(scater)
library(Matrix)
library(SeuratDisk)
library(hdf5r)
library(patchwork)
library(paletteer)
library(hrbrthemes)
library(reticulate)
library(data.table)
library(Matrix)
library(dplyr)
library(purrr)
library(rlang)

```

# Benchmarking ClustoCell's marker prioritizations

## Introduction

Script for Benchmarking ClustoCell's marker prioritizations against contemporary methods.

## Function Defintion

We first define some helper functions.

```{r}

# Personalized Theme function

styled_theme <- function(
  style = c("academic", "minimal", "modern", "dark", "classic", "void"),
  border_mode = c("default", "all", "none"),
  title_align = c("center", "left", "right"),
  legends_to_drop = NULL,
  axes_title_drop = FALSE,
  legend_position = c("default", "top", "bottom", "left", "right", "inside"),
  axis_title_face = c("regular", "bold"),
  x_axis_text_angle = 45,
  x_axis_text_y = -0.3,
  remove_margin = c("none", "bottom", "all", "top", "left", "right")
) {
  style <- match.arg(style)
  border_mode <- match.arg(border_mode)
  title_align <- match.arg(title_align)
  legend_position <- match.arg(legend_position)
  axis_title_face <- match.arg(axis_title_face)
  remove_margin <- match.arg(remove_margin)

  base_font <- "Helvetica"
  align_map <- c(left = 0, center = 0.5, right = 1)
  hjust_val <- align_map[title_align]
  thin_size <- 0.3  

  # Decide axis title face
  axis_face_val <- ifelse(axis_title_face == "bold", "bold", "plain")

  # Base theme selection
  if (style == "void") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        axis.title.x = element_blank(),  
        axis.title.y = element_blank(),  
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "academic") {
    theme_out <- theme_minimal(base_size = 14, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = hjust_val),
        plot.subtitle = element_text(size = 13, hjust = hjust_val, color = "grey30"),
        plot.caption = element_text(size = 10, hjust = 1, color = "grey40"),
        axis.title = element_text(face = axis_face_val, size = 13),
        axis.text = element_text(size = 12, color = "black"),
        panel.grid.major = element_line(color = "grey80", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "minimal") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", size = thin_size),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.text = element_text(color = "black"),
        axis.title = element_text(face = axis_face_val),
        legend.position = "none",
        legend.title = element_text(face = "italic")
      )
  } else if (style == "modern") {
    theme_out <- theme_light(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.border = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "dark") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        plot.background = element_rect(fill = "black", color = NA),
        panel.background = element_rect(fill = "black", color = NA),
        panel.grid.major = element_line(color = "grey30", size = thin_size),
        panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 15, color = "white", hjust = hjust_val),
        plot.subtitle = element_text(color = "grey80", hjust = hjust_val),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white", face = axis_face_val),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white", face = "italic"),
        axis.line = element_line(color = "white", size = thin_size)
      )
  } else if (style == "classic") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  }

  # Apply border mode
  if (border_mode == "all") {
    theme_out <- theme_out + theme(panel.border = element_rect(color = "black", fill = NA, size = thin_size))
  } else if (border_mode == "none") {
    theme_out <- theme_out + theme(panel.border = element_blank(), axis.line = element_blank())
  }

  # Drop axes titles if requested
  if (axes_title_drop) {
    theme_out <- theme_out + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
  }

  # Override x-axis text angle and vertical position
  theme_out <- theme_out + theme(
    axis.text.x = element_text(
      angle = x_axis_text_angle,
      vjust = 1,
      hjust = 1,
      margin = margin(t = x_axis_text_y)
    )
  )

  # Handle margin removal
  if (remove_margin == "all") {
    theme_out <- theme_out + theme(plot.margin = margin(0, 0, 0, 0))
  } else if (remove_margin == "top") {
    theme_out <- theme_out + theme(plot.margin = margin(t = 0))
  } else if (remove_margin == "bottom") {
    theme_out <- theme_out + theme(plot.margin = margin(b = 0))
  } else if (remove_margin == "left") {
    theme_out <- theme_out + theme(plot.margin = margin(l = 0))
  } else if (remove_margin == "right") {
    theme_out <- theme_out + theme(plot.margin = margin(r = 0))
  }
  # if "none", do nothing

  # Legend positioning override
  if (legend_position != "default") {
    if (legend_position == "inside") {
      theme_out <- theme_out + theme(
        legend.position = c(0.98, 0.98),
        legend.justification = c("right", "top"),
        legend.direction = "vertical"
      )
    } else {
      horiz <- legend_position %in% c("top", "bottom")
      theme_out <- theme_out + theme(
        legend.position = legend_position,
        legend.direction = if (horiz) "horizontal" else "vertical"
      )
    }
  }

  # Return theme + guides if needed
  if (!is.null(legends_to_drop)) {
    drop_list <- lapply(legends_to_drop, function(x) "none")
    names(drop_list) <- legends_to_drop
    return(list(theme_out, guides(.list = drop_list)))
  } else {
    return(theme_out)
  }
}

#____________

# Label plot
label_plt <- function(label = "Label", size = 1) {
  ggplot(data.frame(x = 1, y = 1), aes(x = x, y = y)) +
    # Draw a single vertical bar
    geom_col(width = 0.6, fill = "lightgrey", color = "black") +
    # Add vertical text (angle=90)
    annotate("text", x = 1, y = 0.5, label = label, size = size,
             angle = 90, size = 6, vjust = 0.5, hjust = 0.5) +
    # Remove all extra space
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme_void(base_size = 14) +             # no axes, ticks, labels
    theme(
        plot.margin = margin(0, 0, 0, 0),      # remove outside margins
        panel.spacing = unit(0, "pt"),         # remove panel padding
        plot.background = element_blank(),     # no background
        panel.background = element_blank()
    )
}

## Theme function for the table plots
custom_tbl_theme <- ttheme(
  "blank",
  base_size = 5,  # Increased for scaling height
  padding = unit(c(1, 2), "mm"),  # Small horizontal, larger vertical padding to expand row heights; adjust vertical as needed
  tbody.style = tbody_style(
    color = "black",
    fill = cbind(rep(c("white", "grey95"), length.out = 10), rep(c("white", "grey95"), length.out = 10)),
    hjust = 0,  # Left align body text
    x = 0.02,   # Adjust horizontal position for left alignment
    fontsize = 5  # Increased for scaling height
  ),
  colnames.style = colnames_style(
    color = "transparent",  # Make column names invisible to remove header
    fill = "white",
    fontsize = 0  # Set font size to 0 to effectively hide
  )
)

custom_tbl_plusHeader_theme <- ttheme(
  "blank",
  base_size = 5,  # Increased for scaling height
  padding = unit(c(1, 2), "mm"),  # Small horizontal, larger vertical padding to expand row heights; adjust vertical as needed
  tbody.style = tbody_style(
    color = "black",
    fill = cbind(rep(c("white", "grey95"), length.out = 10), rep(c("white", "grey95"), length.out = 10)),
    hjust = 0,  # Left align body text
    x = 0.02,   # Adjust horizontal position for left alignment
    fontsize = 5  # Increased for scaling height
  ),
  colnames.style = colnames_style(
    # color = "transparent",  # Make column names invisible to remove header
    fill = "lightgrey",
    fontsize = 5  # Set font size to 0 to effectively hide
  )
)

#_____________________

# Min-Max Normalize function
min_max_norm <- function(x) {
  min_val <- min(x)
  max_val <- max(x)
  if (max_val > min_val) {
    return((x - min_val) / (max_val - min_val))
  } else {
    return(rep(0, length(x)))
  }
}

#______________________

# Create a custom formatter for bolding only Avg.[Score] in the heatmap plots
bold_labeller <- function(x) {
  lapply(x, function(val) {
    if (val == "Median Score") {
      expression(bold("Median Score"))[[1]]
    } else {
      parse(text = val)[[1]]
    }
  }) |> as.expression()
}

```


```{r}

# Concordance = 100*Sum(rowMax)/N

# Best pals from ggsci: pal_npg (10), pal_d3 (20), pal_ucscgb (26), pal_igv (51)

plotthis::palette_list %>% names()

# NPG colors:
# "#E64B35FF" "#4DBBD5FF" "#00A087FF" "#3C5488FF" "#F39B7FFF" "#8491B4FF" "#91D1C2FF" "#DC0000FF" "#7E6148FF" "#B09C85FF"

# "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF" "#924822FF" "#837B8DFF" "#C75127FF"
# "#D58F5CFF" "#7A65A5FF" "#E4AF69FF" "#3B1B53FF" "#CDDEB7FF" "#612A79FF" "#AE1F63FF" "#E7C76FFF" "#5A655EFF" "#CC9900FF" "#99CC00FF" "#A9A9A9FF" "#CC9900FF"
# "#99CC00FF" "#33CC00FF" "#00CC33FF" "#00CC99FF"

# Define the color palettes

npg_pal <- ggsci::pal_npg()(10)
npg_reverse_pal <- rev(ggsci::pal_npg("nrc")(10))

```

## Data preparation

## Peritumoral Normal Liver

Data was downloaded from https://explore.data.humancellatlas.org/projects/a27dd619-25ad-46a0-ae0c-5c4940a1139b
https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE153643
https://pubmed.ncbi.nlm.nih.gov/33654293/

```{r}

library(zellkonverter)

liver_tumor_adj <- readH5AD("../../../Benchmarking/Test Datasets/Healthy liver tissue adjacent to tumor/GSM4648565_liver_raw_counts.h5ad")
liver_tumor_adj_count <- liver_tumor_adj@assays@data$X %>% t()
liver_tumor_adj_meta <- fread("../../../Benchmarking/Test Datasets/Healthy liver tissue adjacent to tumor/GSM4648565_liver_celltypes.csv.gz", header = F)
colnames(liver_tumor_adj_meta) <- c("Cell", "Celltype")

all(liver_tumor_adj_meta$Cell %in% colnames(liver_tumor_adj_count)) # TRUE
liver_tumor_adj_count <- liver_tumor_adj_count[,liver_tumor_adj_meta$Cell]
all(liver_tumor_adj_meta$Cell == colnames(liver_tumor_adj_count)) # TRUE

liver_tumor_adj <- CreateSeuratObject(counts = liver_tumor_adj_count, meta.data = liver_tumor_adj_meta)
liver_tumor_adj$manual_annot <- liver_tumor_adj$Celltype
liver_tumor_adj$major_manual_annot <- sapply(liver_tumor_adj$manual_annot, function(i) unlist(str_split(i, " "))[1])

liver_tumor_adj@meta.data <- liver_tumor_adj@meta.data %>%
  dplyr::mutate(minor_manual_annot1 = case_when(
    as.character(manual_annot) %in% c("Epithelial (Cholangiocyte)") ~ "Cholangiocyte",
    as.character(manual_annot) %in% c("Epithelial (Hepatocyte)") ~ "Hepatocyte",
    as.character(manual_annot) %in% c("Immune (Lymphocyte)") ~ "Lymphocyte",
    as.character(manual_annot) %in% c("Immune (Macrophage)") ~ "Myeloid",
    as.character(manual_annot) %in% c("Immune (Myeloid)") ~ "Myeloid",
    as.character(manual_annot) %in% c("Stromal (Endothelial)") ~ "Endothelial",
    as.character(manual_annot) %in% c("Stromal (SMC)") ~ "SMC",
    .default = manual_annot
  ))

# Normalizing the data
liver_tumor_adj <- NormalizeData(liver_tumor_adj, normalization.method = "LogNormalize", scale.factor = 10000)

# Identification of highly variable features (feature selection)
liver_tumor_adj <- FindVariableFeatures(liver_tumor_adj, selection.method = "vst", nfeatures = 2000)

# Scaling the data
liver_tumor_adj <- ScaleData(liver_tumor_adj, features = VariableFeatures(liver_tumor_adj))

# Perform linear dimensional reduction
liver_tumor_adj <- RunPCA(liver_tumor_adj, features = VariableFeatures(object = liver_tumor_adj))

# Elbow plot
ElbowPlot(liver_tumor_adj, ndims = 50)

# Cluster the cells
# liver_tumor_adj <- FindNeighbors(liver_tumor_adj, dims = 1:20)
# liver_tumor_adj <- FindClusters(liver_tumor_adj, resolution = 0.5)

# Run non-linear dimensional reduction
liver_tumor_adj <- RunUMAP(liver_tumor_adj, dims = 1:20)

qs2::qs_save(object = liver_tumor_adj, file = "../../../Benchmarking/Test Datasets/Healthy liver tissue adjacent to tumor/liver_tumor_adj.qs2")

qs2::qs_save(object = liver_tumor_adj, file = "../../../Benchmarking/Test Datasets/Healthy liver tissue adjacent to tumor/liver_tumor_adj.qs2")

liver_tumor_adj <- qs2::qs_read("../../../Benchmarking/Test Datasets/Healthy liver tissue adjacent to tumor/liver_tumor_adj.qs2")
liver_tumor_adj_manual_annot_markoClust <- qs2::qs_read("../../../Benchmarking/Deconvolution/liver_tumor_adj_manual_annot_markoClust.qs2")

```

## Prostate Cell Atlas (EGAS00001005787)

The data was downloaded from https://www.prostatecellatlas.org/
https://ega-archive.org/studies/EGAS00001005787

```{r}

prostateimmuneatlas <- read_rds("../../../Benchmarking/Test Datasets/Prostate Cancer Cell Atlas/prostate_portal_300921.RDS")
prostateimmuneatlas <- Seurat::as.Seurat(prostateimmuneatlas)

prostateimmuneatlas$manual_annot <- as.character(prostateimmuneatlas$celltype)
prostateimmuneatlas$manual_annot[prostateimmuneatlas$manual_annot %in% c("LE-KLK3", "LE-KLK4")] <- "LE"

prostateimmuneatlas@meta.data <- prostateimmuneatlas@meta.data %>%
  dplyr::mutate(manual_annot = case_when(
    as.character(manual_annot) %in% c("Mast cell") ~ "Mast",
    as.character(manual_annot) %in% c("NK cell") ~ "NK",
    as.character(manual_annot) %in% c("T cell") ~ "T",
    as.character(manual_annot) %in% c("B cell") ~ "B",
    .default = manual_annot
  ))

# Normalizing the data
prostateimmuneatlas <- NormalizeData(prostateimmuneatlas, normalization.method = "LogNormalize", scale.factor = 10000)

# Identification of highly variable features (feature selection)
prostateimmuneatlas <- FindVariableFeatures(prostateimmuneatlas, selection.method = "vst", nfeatures = 2000)

# Scaling the data
prostateimmuneatlas <- ScaleData(prostateimmuneatlas, features = VariableFeatures(prostateimmuneatlas))

prostateimmuneatlas[["RNA"]] <- prostateimmuneatlas[["originalexp"]]
DefaultAssay(prostateimmuneatlas) <- "RNA"
prostateimmuneatlas[["originalexp"]] <- NULL

qs2::qs_save(object = prostateimmuneatlas, file = "../../../Benchmarking/Test Datasets/Prostate Cancer Cell Atlas/prostateimmuneatlas.qs2")

prostateimmuneatlas <- qs2::qs_read("../../../Benchmarking/Test Datasets/Prostate Cancer Cell Atlas/prostateimmuneatlas.qs2")
prostateimmuneatlas_manual_annot_markoClust <- qs2::qs_read("../../../Benchmarking/Deconvolution/prostateimmuneatlas_manual_annot_markoClust.qs2")

```

## Query Data Prep

### GSE115469

Data was downloaded from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE115469
https://www.ncbi.nlm.nih.gov/pubmed/30348985

```{r}

Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 100)

GSE115469_df <- read_csv(file = "../../../Benchmarking/Test Datasets/GSE115469/GSE115469_Data.csv.gz") # Data is already normalized and log-transformed
GSE115469_df <- GSE115469_df %>% column_to_rownames("...1")

GSE115469_meta_df <- read_delim(file = "../../../Benchmarking/Test Datasets/GSE115469/GSE115469_CellClusterType.txt.gz")

ncol(GSE115469_df) == nrow(GSE115469_meta_df) # T
identical(colnames(GSE115469_df), GSE115469_meta_df$CellName) # T

GSE115469_so <- CreateSeuratObject(counts = GSE115469_df, meta.data = GSE115469_meta_df)

GSE115469_so$CellType %>% table() %>% names() %>% cat(sep = ", ")

liver_tumor_adj$minor_manual_annot1 %>% table() %>% names() %>% cat(sep = ", ")

```

### GSE185344
Data was downloaded from https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE185344
https://www.ncbi.nlm.nih.gov/pubmed/36229464
https://www.nature.com/articles/s41467-022-33780-1

```{r}

GSE185344_so <- readRDS(
  gzcon(gzfile("../../../Benchmarking/Test Datasets/GSE185344/GSE185344_PH_scRNA.final.rds.gz"))
)

GSE185344_so <- GSE185344_so$obj

GSE185344_meta <- read.csv("../../../Benchmarking/Test Datasets/GSE185344/GSE185344_PH_scRNA.rename_cluster.csv")

nrow(GSE185344_meta) == ncol(GSE185344_so) # T
identical(GSE185344_meta$X, colnames(GSE185344_so)) #T

GSE185344_so@meta.data <- cbind(GSE185344_so@meta.data, GSE185344_meta)

GSE185344_so$cellactivity_clusters %>% table() %>% names() %>% cat(sep = ", ")

prostateimmuneatlas$manual_annot %>% table() %>% names() %>% cat(sep = ", ")

```

## Performing the benchmarking

### Input params and helper functions

```{r}

# ==============================================================================
# ClustoCell marker benchmarking for liver and prostate scRNA-seq datasets
# ==============================================================================
# Reference datasets (marker discovery / signature construction):
#   liver_tumor_adj       labels: minor_manual_annot1
#   prostateimmuneatlas   labels: manual_annot
#
# Query datasets (pseudobulk generation / ground-truth proportions):
#   GSE115469_so          labels: CellType -> benchmark_celltype
#   GSE185344_so          labels: cellactivity_clusters -> benchmark_celltype
#
# Benchmarks:
#   1. DWLS default Seurat marker signature vs ClustoCell-marker signature
#   2. BayesPrism default marker selection vs ClustoCell-marker gene space
#
# ============================================================================== 

# ------------------------------- 0. PACKAGES -----------------------------------
suppressPackageStartupMessages({
  library(Seurat)
  library(dplyr)
  library(tidyr)
  library(purrr)
  library(ggplot2)
  library(Matrix)
  library(DescTools)
  library(boot)
  library(patchwork)
})

# ----------------------------- 1. USER SETTINGS --------------------------------
DWLS_SOURCE <- "../../../Benchmarking/Deconvolution/DWLS/Deconvolution_functions.R"
BASE_OUT_DIR <- "../../../Benchmarking/Deconvolution/Liver_Prostate"

N_BULK     <- 100
CELLS_BULK <- 1000
BOOT_R      <- 5000
TOP_N       <- 10
SEED_GLOBAL <- 121

# BayesPrism settings
RUN_BAYESPRISM <- TRUE
SPECIES         <- "hs"
PSEUDO_COUNT    <- 0.1
CELL_COUNT_CUT  <- 50
PVAL_MAX        <- 0.01
LFC_MIN         <- 0.1
EXP_CELLS       <- 5
OUTLIER_CUT     <- 0.01
OUTLIER_FRAC    <- 0.1
N_CORES         <- max(1, parallel::detectCores() - 1)
SEED_BP         <- 2024

# Reproduce the ordering used in the supplied PBMC script.
# TRUE ranks significant default markers by decreasing adjusted P value.
# Set FALSE for the more conventional most-significant-first ordering.
DEFAULT_ORDER_PADJ_DECREASING <- TRUE

dir.create(BASE_OUT_DIR, recursive = TRUE, showWarnings = FALSE)

if (!file.exists(DWLS_SOURCE)) {
  stop("DWLS functions file not found: ", DWLS_SOURCE)
}
source(DWLS_SOURCE)

if (RUN_BAYESPRISM && !requireNamespace("BayesPrism", quietly = TRUE)) {
  stop(
    "RUN_BAYESPRISM is TRUE but BayesPrism is not installed. Install it with:\n",
    "remotes::install_github('Danko-Lab/BayesPrism/BayesPrism')\n",
    "or set RUN_BAYESPRISM <- FALSE."
  )
}

# --------------------------- 2. INPUT VALIDATION -------------------------------
required_objects <- c(
  "liver_tumor_adj",
  "GSE115469_so",
  "liver_tumor_adj_manual_annot_markoClust",
  "prostateimmuneatlas",
  "GSE185344_so",
  "prostateimmuneatlas_manual_annot_markoClust"
)
missing_objects <- required_objects[!vapply(required_objects, exists, logical(1), inherits = TRUE)]
if (length(missing_objects) > 0) {
  stop("Missing required object(s): ", paste(missing_objects, collapse = ", "))
}

# ------------------------------- 3. HELPERS ------------------------------------
get_counts <- function(obj, assay = "RNA") {
  out <- tryCatch(
    GetAssayData(obj, assay = assay, layer = "counts"),
    error = function(e) GetAssayData(obj, assay = assay, slot = "counts")
  )
  if (nrow(out) == 0 || ncol(out) == 0) stop("Empty count matrix.")
  out
}

clean_labels <- function(x) {
  x <- as.character(x)
  x[x == "" | is.na(x)] <- NA_character_
  x
}

scalar_ccc <- function(truth, pred) {
  keep <- is.finite(truth) & is.finite(pred)
  truth <- truth[keep]
  pred  <- pred[keep]
  if (length(truth) < 2 || stats::sd(truth) == 0 || stats::sd(pred) == 0) {
    return(NA_real_)
  }
  out <- tryCatch(DescTools::CCC(truth, pred)$rho.c, error = function(e) NA_real_)
  if (is.data.frame(out) || is.matrix(out)) out <- out[1, 1]
  as.numeric(out)[1]
}

metric_summary <- function(truth, pred) {
  keep <- is.finite(truth) & is.finite(pred)
  truth <- truth[keep]
  pred  <- pred[keep]
  if (length(truth) == 0) {
    return(tibble(Pearson = NA_real_, Spearman = NA_real_, CCC = NA_real_,
                  RMSE = NA_real_, MAE = NA_real_))
  }
  pearson <- if (length(truth) < 2 || sd(truth) == 0 || sd(pred) == 0) {
    NA_real_
  } else {
    suppressWarnings(cor(truth, pred, method = "pearson"))
  }
  spearman <- if (length(truth) < 2 || sd(truth) == 0 || sd(pred) == 0) {
    NA_real_
  } else {
    suppressWarnings(cor(truth, pred, method = "spearman"))
  }
  tibble(
    Pearson  = pearson,
    Spearman = spearman,
    CCC      = scalar_ccc(truth, pred),
    RMSE     = sqrt(mean((pred - truth)^2)),
    MAE      = mean(abs(pred - truth))
  )
}

celltype_metrics <- function(truth_mat, pred_mat, method_name) {
  stopifnot(identical(dim(truth_mat), dim(pred_mat)))
  bind_rows(lapply(colnames(truth_mat), function(ct) {
    metric_summary(truth_mat[, ct], pred_mat[, ct]) %>%
      mutate(CellType = ct, Method = method_name, .before = 1)
  }))
}

boot_diff_ci <- function(x, y, R = BOOT_R) {
  keep <- is.finite(x) & is.finite(y)
  d <- x[keep] - y[keep]
  if (length(d) < 2) {
    return(list(Mean = if (length(d)) mean(d) else NA_real_,
                Lower = NA_real_, Upper = NA_real_))
  }
  stat <- function(data, idx) mean(data[idx], na.rm = TRUE)
  b <- boot::boot(d, statistic = stat, R = R)
  ci <- tryCatch(boot::boot.ci(b, type = "perc"), error = function(e) NULL)
  if (is.null(ci) || is.null(ci$percent)) {
    qs <- quantile(b$t[, 1], c(0.025, 0.975), na.rm = TRUE)
    return(list(Mean = mean(d), Lower = unname(qs[1]), Upper = unname(qs[2])))
  }
  list(Mean = mean(d), Lower = ci$percent[4], Upper = ci$percent[5])
}

extract_top_clustocell_markers <- function(markoclust_obj, top_n = 10) {
  marker_tables <- markoclust_obj$markers$major_clusters$cluster_specific$positive_markers
  if (is.null(marker_tables) || !is.list(marker_tables)) {
    stop("Could not find markers$major_clusters$cluster_specific$positive_markers.")
  }
  out <- lapply(marker_tables, function(tbl) {
    if (!all(c("Rank", "Feature") %in% colnames(tbl))) {
      stop("Each marker table must contain Rank and Feature columns.")
    }
    tbl %>%
      filter(!is.na(Rank), Rank <= top_n, !is.na(Feature), Feature != "") %>%
      arrange(Rank) %>%
      pull(Feature) %>%
      unique()
  })
  out <- out[lengths(out) > 0]
  out[names(out) != "NA" & !is.na(names(out))]
}

combine_marker_classes <- function(markers_list, class_map) {
  # class_map must be a named character vector: old label -> harmonized label/NA.
  old_names <- names(markers_list)
  new_names <- unname(class_map[old_names])
  keep <- !is.na(new_names) & new_names != ""
  split_genes <- split(unlist(markers_list[keep], use.names = FALSE),
                       rep(new_names[keep], lengths(markers_list[keep])))
  lapply(split_genes, unique)
}

restrict_marker_list <- function(markers_list, allowed_celltypes, reference_genes) {
  markers_list <- markers_list[intersect(names(markers_list), allowed_celltypes)]
  markers_list <- lapply(markers_list, function(x) unique(intersect(x, reference_genes)))
  markers_list[lengths(markers_list) > 0]
}

make_pseudobulks <- function(query_counts, query_labels, celltypes,
                             n_bulk = N_BULK, cells_per_bulk = CELLS_BULK,
                             seed = SEED_GLOBAL) {
  query_labels <- clean_labels(query_labels)
  names(query_labels) <- colnames(query_counts)
  eligible <- names(query_labels)[query_labels %in% celltypes]
  if (length(eligible) == 0) stop("No eligible query cells after label harmonization.")
  if (cells_per_bulk > length(eligible)) {
    warning("CELLS_BULK reduced from ", cells_per_bulk, " to ", length(eligible),
            " because the query has fewer eligible cells.")
    cells_per_bulk <- length(eligible)
  }

  pseudobulk_list <- vector("list", n_bulk)
  truth_list <- vector("list", n_bulk)
  set.seed(seed)
  for (i in seq_len(n_bulk)) {
    set.seed(seed + i)
    sampled_cells <- sample(eligible, size = cells_per_bulk, replace = FALSE)
    pseudobulk_list[[i]] <- Matrix::rowSums(query_counts[, sampled_cells, drop = FALSE])
    tab <- table(factor(query_labels[sampled_cells], levels = celltypes))
    prop <- as.numeric(tab) / sum(tab)
    names(prop) <- celltypes
    truth_list[[i]] <- prop
  }
  truth_mat <- do.call(rbind, truth_list)
  rownames(truth_mat) <- paste0("pseudobulk_", seq_len(n_bulk))
  list(bulk_list = pseudobulk_list, truth = truth_mat,
       cells_per_bulk = cells_per_bulk)
}

build_signature_default <- function(obj, label_col, allowed_celltypes,
                                    diff.cutoff = 0.5, pval.cutoff = 0.01,
                                    order_padj_decreasing = DEFAULT_ORDER_PADJ_DECREASING) {
  labels <- clean_labels(obj[[label_col, drop = TRUE]])
  keep_cells <- colnames(obj)[labels %in% allowed_celltypes]
  obj <- subset(obj, cells = keep_cells)
  obj[[label_col]] <- factor(as.character(obj[[label_col, drop = TRUE]]),
                             levels = allowed_celltypes)
  Idents(obj) <- label_col
  counts <- get_counts(obj)
  labels <- as.character(obj[[label_col, drop = TRUE]])

  de_all <- FindAllMarkers(
    obj,
    only.pos = TRUE,
    test.use = "wilcox",
    logfc.threshold = diff.cutoff,
    min.pct = 0.1,
    verbose = FALSE
  )
  fc_col <- intersect(c("avg_log2FC", "avg_logFC"), colnames(de_all))[1]
  if (is.na(fc_col)) stop("Could not identify Seurat fold-change column.")
  de_filtered <- de_all[
    is.finite(de_all$p_val_adj) & de_all$p_val_adj < pval.cutoff &
      is.finite(de_all[[fc_col]]) & de_all[[fc_col]] > diff.cutoff,
    , drop = FALSE
  ]

  cluster_tables <- setNames(vector("list", length(allowed_celltypes)), allowed_celltypes)
  number_genes <- setNames(integer(length(allowed_celltypes)), allowed_celltypes)
  for (ct in allowed_celltypes) {
    x <- de_filtered[as.character(de_filtered$cluster) == ct, , drop = FALSE]
    if (nrow(x)) {
      x <- x[grep("^MIR|^Mir", x$gene, invert = TRUE), , drop = FALSE]
      x <- x[order(x$p_val_adj, decreasing = order_padj_decreasing), , drop = FALSE]
    }
    cluster_tables[[ct]] <- x
    number_genes[ct] <- nrow(x)
  }
  if (any(number_genes == 0)) {
    stop("No default DE markers for: ", paste(names(number_genes)[number_genes == 0], collapse = ", "))
  }

  max_g <- min(200, max(number_genes))
  min_g <- min(50, min(number_genes))
  if (max_g < 2) stop("Too few default DE genes to build a signature.")
  G_range <- seq.int(max(2, min_g), max_g)
  kappas <- rep(Inf, length(G_range))

  for (j in seq_along(G_range)) {
    G <- G_range[j]
    genes <- unique(unlist(lapply(allowed_celltypes, function(ct) {
      head(cluster_tables[[ct]]$gene, min(G, number_genes[ct]))
    })))
    genes <- intersect(genes, rownames(counts))
    if (length(genes) < 2) next
    sig <- sapply(allowed_celltypes, function(ct) {
      Matrix::rowMeans(counts[genes, labels == ct, drop = FALSE])
    })
    kappas[j] <- tryCatch(kappa(sig), error = function(e) Inf)
  }
  best <- which.min(kappas)
  if (!length(best) || !is.finite(kappas[best])) stop("No valid default signature matrix.")
  G <- G_range[best]
  genes <- unique(unlist(lapply(allowed_celltypes, function(ct) {
    head(cluster_tables[[ct]]$gene, min(G, number_genes[ct]))
  })))
  genes <- intersect(genes, rownames(counts))
  sig <- sapply(allowed_celltypes, function(ct) {
    Matrix::rowMeans(counts[genes, labels == ct, drop = FALSE])
  })
  rownames(sig) <- genes
  colnames(sig) <- allowed_celltypes
  attr(sig, "optimal_G") <- G
  sig
}

build_signature_clustocell <- function(ref_counts, ref_labels, markers_list,
                                       allowed_celltypes) {
  ref_labels <- clean_labels(ref_labels)
  markers_list <- restrict_marker_list(markers_list, allowed_celltypes, rownames(ref_counts))
  missing_ct <- setdiff(allowed_celltypes, names(markers_list))
  if (length(missing_ct)) {
    stop("No usable ClustoCell markers for: ", paste(missing_ct, collapse = ", "))
  }
  genes <- unique(unlist(markers_list[allowed_celltypes], use.names = FALSE))
  sig <- sapply(allowed_celltypes, function(ct) {
    cells <- which(ref_labels == ct)
    Matrix::rowMeans(ref_counts[genes, cells, drop = FALSE])
  })
  rownames(sig) <- genes
  colnames(sig) <- allowed_celltypes
  sig
}

safe_deconvolve_dwls <- function(sig, bulk_list, celltypes, method_label) {
  results <- matrix(NA_real_, nrow = length(bulk_list), ncol = length(celltypes),
                    dimnames = list(paste0("pseudobulk_", seq_along(bulk_list)), celltypes))
  for (i in seq_along(bulk_list)) {
    tr <- trimData(sig, bulk_list[[i]])
    sol <- tryCatch({
      if (nrow(tr$sig) < 2 || ncol(tr$sig) < 2 || length(tr$bulk) < 2) {
        stop("Too little data after trimData().")
      }
      capture.output(ans <- solveDampenedWLS(tr$sig, tr$bulk))
      ans
    }, error = function(e) {
      warning(sprintf("[%s] pseudobulk %d failed: %s", method_label, i, conditionMessage(e)))
      NULL
    })
    if (!is.null(sol)) {
      sol <- as.numeric(sol)
      names(sol) <- colnames(tr$sig)
      sol[!is.finite(sol)] <- 0
      sol[sol < 0] <- 0
      if (sum(sol) > 0) sol <- sol / sum(sol)
      results[i, intersect(names(sol), celltypes)] <- sol[intersect(names(sol), celltypes)]
      results[i, is.na(results[i, ])] <- 0
    }
  }
  results
}

compute_comparison <- function(truth, pred_default, pred_custom,
                               default_name, custom_name, out_dir) {
  n_bulk <- nrow(truth)
  bulk_metrics <- map_dfr(seq_len(n_bulk), function(i) {
    bind_rows(
      metric_summary(truth[i, ], pred_default[i, ]) %>%
        mutate(Method = default_name, Bulk = i),
      metric_summary(truth[i, ], pred_custom[i, ]) %>%
        mutate(Method = custom_name, Bulk = i)
    )
  })
  ct_metrics <- bind_rows(
    celltype_metrics(truth, pred_default, default_name),
    celltype_metrics(truth, pred_custom, custom_name)
  )

  metric_names <- c("Pearson", "Spearman", "CCC", "RMSE", "MAE")
  stat_results <- map_dfr(metric_names, function(m) {
    def <- bulk_metrics %>% filter(Method == default_name) %>% pull(all_of(m))
    cus <- bulk_metrics %>% filter(Method == custom_name) %>% pull(all_of(m))
    keep <- is.finite(def) & is.finite(cus)
    def2 <- def[keep]
    cus2 <- cus[keep]
    tt <- tryCatch(t.test(cus2, def2, paired = TRUE), error = function(e) NULL)
    wt <- tryCatch(wilcox.test(cus2, def2, paired = TRUE, exact = FALSE), error = function(e) NULL)
    bci <- boot_diff_ci(cus2, def2)
    tibble(
      Metric = m,
      Mean_Default = mean(def2, na.rm = TRUE),
      Mean_ClustoCell = mean(cus2, na.rm = TRUE),
      Diff_Mean = bci$Mean,
      CI_Lower = bci$Lower,
      CI_Upper = bci$Upper,
      t_pvalue = if (is.null(tt)) NA_real_ else tt$p.value,
      wilcox_pvalue = if (is.null(wt)) NA_real_ else wt$p.value
    )
  })

  write.csv(bulk_metrics, file.path(out_dir, "metrics_per_bulk.csv"), row.names = FALSE)
  write.csv(ct_metrics, file.path(out_dir, "metrics_per_celltype.csv"), row.names = FALSE)
  write.csv(stat_results, file.path(out_dir, "paired_statistics.csv"), row.names = FALSE)

  list(bulk_metrics = bulk_metrics, ct_metrics = ct_metrics,
       stat_results = stat_results)
}

make_comparison_plots <- function(result, default_name, custom_name, algorithm_label,
                                  out_dir) {
  bulk_metrics <- result$bulk_metrics
  stats <- result$stat_results %>%
    mutate(
      Metric = factor(Metric, levels = c("RMSE", "MAE", "CCC", "Pearson", "Spearman")),
      Label = case_when(
        wilcox_pvalue < 0.001 ~ "***",
        wilcox_pvalue < 0.01 ~ "**",
        wilcox_pvalue < 0.05 ~ "*",
        TRUE ~ "ns"
      )
    )

  p_forest <- ggplot(stats, aes(x = Diff_Mean, y = Metric)) +
    geom_vline(xintercept = 0, linetype = 2, colour = "grey60") +
    geom_errorbarh(aes(xmin = CI_Lower, xmax = CI_Upper), height = 0.18, linewidth = 0.5) +
    geom_point(size = 1.2) +
    geom_text(aes(label = Label, x = CI_Upper), nudge_x = 0.01, size = 2.5) +
    coord_flip() +
    labs(x = paste0("Mean(ClustoCell − ", algorithm_label, ")"), y = NULL) +
    theme_classic(base_size = 7) +
    theme(axis.text.x = element_text(angle = 25, vjust = 0.8))

  method_cols <- setNames(c("#E69F00", "#56B4E9"), c(default_name, custom_name))
  p_rmse <- ggplot(bulk_metrics, aes(x = Method, y = RMSE, group = Bulk)) +
    geom_line(colour = "grey65", alpha = 0.3) +
    geom_point(aes(fill = Method), shape = 21, size = 1.5, stroke = 0.3) +
    scale_fill_manual(values = method_cols) +
    labs(x = NULL, y = "RMSE") +
    theme_classic(base_size = 7) +
    theme(legend.position = "none", axis.text.x = element_text(angle = 25, vjust = 0.68))

  p_ccc <- ggplot(bulk_metrics, aes(x = Method, y = CCC, group = Bulk)) +
    geom_line(colour = "grey65", alpha = 0.3) +
    geom_point(aes(fill = Method), shape = 21, size = 1.5, stroke = 0.3) +
    scale_fill_manual(values = method_cols) +
    labs(x = NULL, y = "CCC") +
    theme_classic(base_size = 7) +
    theme(legend.position = "none", axis.text.x = element_text(angle = 25, vjust = 0.68))

  combined <- p_forest / (p_rmse | p_ccc) + plot_layout(heights = c(1.15, 1))
  ggsave(file.path(out_dir, "benchmark_summary.pdf"), combined,
         width = 7.2, height = 5.2, device = cairo_pdf)
  ggsave(file.path(out_dir, "benchmark_summary.png"), combined,
         width = 7.2, height = 5.2, dpi = 600)
  list(forest = p_forest, rmse = p_rmse, ccc = p_ccc, combined = combined)
}

# -------------------------- 4. BAYESPRISM HELPERS -------------------------------
prepare_bp_default <- function(sc_dat, labels) {
  sc_filt <- BayesPrism::cleanup.genes(
    input = sc_dat,
    input.type = "count.matrix",
    species = SPECIES,
    gene.group = c("Rb", "Mrp", "other_Rb", "chrM", "MALAT1", "chrX", "chrY"),
    exp.cells = EXP_CELLS
  )
  sc_pc <- BayesPrism::select.gene.type(sc_filt, gene.type = "protein_coding")
  keep <- colSums(sc_pc > 0) > 3
  stat <- BayesPrism::get.exp.stat(
    sc.dat = sc_pc[, keep, drop = FALSE],
    cell.type.labels = labels,
    cell.state.labels = labels,
    pseudo.count = PSEUDO_COUNT,
    cell.count.cutoff = CELL_COUNT_CUT,
    n.cores = N_CORES
  )
  BayesPrism::select.marker(sc.dat = sc_pc, stat = stat,
                            pval.max = PVAL_MAX, lfc.min = LFC_MIN)
}

prepare_bp_custom <- function(sc_dat, markers_list, celltypes) {
  genes <- unique(unlist(markers_list[celltypes], use.names = FALSE))
  genes <- intersect(genes, colnames(sc_dat))
  if (length(genes) < 2) stop("Too few ClustoCell markers in BayesPrism reference.")
  sc_dat[, genes, drop = FALSE]
}

run_bp <- function(ref_mat, labels, bulk_mat, celltypes, method_label, result_file) {
  shared <- intersect(colnames(ref_mat), colnames(bulk_mat))
  if (length(shared) < 2) stop("Too few shared genes for ", method_label)
  ref_mat <- ref_mat[, shared, drop = FALSE]
  bulk_mat <- bulk_mat[, shared, drop = FALSE]
  prism <- BayesPrism::new.prism(
    reference = ref_mat,
    mixture = bulk_mat,
    input.type = "count.matrix",
    cell.type.labels = labels,
    cell.state.labels = labels,
    key = NULL,
    outlier.cut = OUTLIER_CUT,
    outlier.fraction = OUTLIER_FRAC
  )
  set.seed(SEED_BP)
  fit <- BayesPrism::run.prism(prism = prism, n.cores = N_CORES)
  saveRDS(fit, result_file)
  theta <- BayesPrism::get.fraction(fit, which.theta = "final", state.or.type = "type")
  pred <- matrix(0, nrow = nrow(bulk_mat), ncol = length(celltypes),
                 dimnames = list(rownames(bulk_mat), celltypes))
  common <- intersect(colnames(theta), celltypes)
  pred[rownames(theta), common] <- theta[, common, drop = FALSE]
  pred
}

# ----------------------- 5. GENERIC DATASET BENCHMARK ---------------------------
run_dataset_benchmark <- function(dataset_name,
                                  reference_obj,
                                  reference_label_col,
                                  query_obj,
                                  query_label_col,
                                  markers_list,
                                  output_root) {
  out_dataset <- file.path(output_root, dataset_name)
  out_dwls <- file.path(out_dataset, "DWLS")
  out_bp <- file.path(out_dataset, "BayesPrism")
  dir.create(out_dwls, recursive = TRUE, showWarnings = FALSE)
  if (RUN_BAYESPRISM) dir.create(out_bp, recursive = TRUE, showWarnings = FALSE)

  ref_counts <- get_counts(reference_obj)
  qry_counts <- get_counts(query_obj)
  ref_labels <- clean_labels(reference_obj[[reference_label_col, drop = TRUE]])
  qry_labels <- clean_labels(query_obj[[query_label_col, drop = TRUE]])
  names(ref_labels) <- colnames(ref_counts)
  names(qry_labels) <- colnames(qry_counts)

  # Use only classes present in both datasets and represented by marker lists.
  celltypes <- sort(Reduce(intersect, list(
    unique(ref_labels[!is.na(ref_labels)]),
    unique(qry_labels[!is.na(qry_labels)]),
    names(markers_list)
  )))
  if (length(celltypes) < 2) stop(dataset_name, ": fewer than two shared classes.")

  cat("\n", strrep("=", 78), "\n", dataset_name, "\n", sep = "")
  cat("Benchmark classes: ", paste(celltypes, collapse = ", "), "\n", sep = "")
  cat("Reference counts:\n")
  print(table(factor(ref_labels, levels = celltypes), useNA = "ifany"))
  cat("Query counts:\n")
  print(table(factor(qry_labels, levels = celltypes), useNA = "ifany"))

  markers_list <- restrict_marker_list(markers_list, celltypes, rownames(ref_counts))
  marker_summary <- tibble(
    CellType = celltypes,
    N_requested = lengths(markers_list[celltypes]),
    N_in_reference = lengths(markers_list[celltypes]),
    Markers = vapply(markers_list[celltypes], paste, collapse = ";", FUN.VALUE = character(1))
  )
  write.csv(marker_summary, file.path(out_dataset, "clustocell_marker_summary.csv"), row.names = FALSE)

  pb <- make_pseudobulks(qry_counts, qry_labels, celltypes)
  truth <- pb$truth
  bulk_list <- pb$bulk_list
  saveRDS(list(truth = truth, bulk_list = bulk_list, celltypes = celltypes),
          file.path(out_dataset, "pseudobulks_and_truth.rds"))

  # DWLS
  cat("Building default DWLS signature...\n")
  sig_default <- build_signature_default(reference_obj, reference_label_col, celltypes)
  cat("Default signature: ", nrow(sig_default), " genes x ", ncol(sig_default),
      " classes; optimal G = ", attr(sig_default, "optimal_G"), "\n", sep = "")
  cat("Building ClustoCell DWLS signature...\n")
  sig_custom <- build_signature_clustocell(ref_counts, ref_labels, markers_list, celltypes)
  write.csv(sig_default, file.path(out_dwls, "signature_default.csv"))
  write.csv(sig_custom, file.path(out_dwls, "signature_clustocell.csv"))

  pred_default <- safe_deconvolve_dwls(sig_default, bulk_list, celltypes, "DWLS")
  pred_custom <- safe_deconvolve_dwls(sig_custom, bulk_list, celltypes, "ClustoCell-DWLS")
  saveRDS(list(truth = truth, DWLS = pred_default, ClustoCell_DWLS = pred_custom),
          file.path(out_dwls, "deconvolution_predictions.rds"))
  dwls_result <- compute_comparison(truth, pred_default, pred_custom,
                                    "DWLS", "ClustoCell-DWLS", out_dwls)
  dwls_plots <- make_comparison_plots(dwls_result, "DWLS", "ClustoCell-DWLS",
                                      "DWLS", out_dwls)

  bp_result <- NULL
  bp_plots <- NULL
  if (RUN_BAYESPRISM) {
    cat("Preparing BayesPrism matrices...\n")
    keep_ref <- ref_labels %in% celltypes
    sc_dat <- as.matrix(Matrix::t(ref_counts[, keep_ref, drop = FALSE]))
    bp_labels <- ref_labels[keep_ref]
    stopifnot(nrow(sc_dat) == length(bp_labels))
    bulk_mat <- do.call(rbind, bulk_list)
    rownames(bulk_mat) <- rownames(truth)
    storage.mode(bulk_mat) <- "double"

    cat("Running default BayesPrism arm...\n")
    bp_ref_default <- prepare_bp_default(sc_dat, bp_labels)
    writeLines(colnames(bp_ref_default), file.path(out_bp, "markers_default.txt"))
    bp_pred_default <- run_bp(bp_ref_default, bp_labels, bulk_mat, celltypes,
                              "BayesPrism", file.path(out_bp, "bp_default_fit.rds"))

    cat("Running ClustoCell BayesPrism arm...\n")
    bp_ref_custom <- prepare_bp_custom(sc_dat, markers_list, celltypes)
    writeLines(colnames(bp_ref_custom), file.path(out_bp, "markers_clustocell.txt"))
    bp_pred_custom <- run_bp(bp_ref_custom, bp_labels, bulk_mat, celltypes,
                             "ClustoCell-BayesPrism",
                             file.path(out_bp, "bp_clustocell_fit.rds"))

    saveRDS(list(truth = truth, BayesPrism = bp_pred_default,
                 ClustoCell_BayesPrism = bp_pred_custom),
            file.path(out_bp, "deconvolution_predictions.rds"))
    bp_result <- compute_comparison(truth, bp_pred_default, bp_pred_custom,
                                    "BayesPrism", "ClustoCell-BayesPrism", out_bp)
    bp_plots <- make_comparison_plots(bp_result, "BayesPrism",
                                      "ClustoCell-BayesPrism", "BayesPrism", out_bp)
  }

  saveRDS(list(celltypes = celltypes, DWLS = dwls_result, BayesPrism = bp_result),
          file.path(out_dataset, "benchmark_results.rds"))
  invisible(list(celltypes = celltypes, DWLS = dwls_result,
                 DWLS_plots = dwls_plots, BayesPrism = bp_result,
                 BayesPrism_plots = bp_plots))
}

```

### Liver prep

```{r}

# ==============================================================================
# 6. LIVER LABEL MAPPING AND MARKERS
# ==============================================================================
liver_map <- c(
  "alpha-beta_T_Cells"            = "Lymphocyte",
  "Central_venous_LSECs"         = "Endothelial",
  "Cholangiocytes"               = "Cholangiocyte",
  "Erythroid_Cells"              = NA_character_,
  "gamma-delta_T_Cells_1"        = "Lymphocyte",
  "gamma-delta_T_Cells_2"        = "Lymphocyte",
  "Hepatic_Stellate_Cells"       = "SMC",
  "Hepatocyte_1"                 = "Hepatocyte",
  "Hepatocyte_2"                 = "Hepatocyte",
  "Hepatocyte_3"                 = "Hepatocyte",
  "Hepatocyte_4"                 = "Hepatocyte",
  "Hepatocyte_5"                 = "Hepatocyte",
  "Hepatocyte_6"                 = "Hepatocyte",
  "Inflammatory_Macrophage"      = "Myeloid",
  "Mature_B_Cells"               = "Lymphocyte",
  "NK-like_Cells"                = "Lymphocyte",
  "Non-inflammatory_Macrophage"  = "Myeloid",
  "Periportal_LSECs"             = "Endothelial",
  "Plasma_Cells"                 = "Lymphocyte",
  "Portal_endothelial_Cells"     = "Endothelial"
)

GSE115469_so$benchmark_celltype <- unname(liver_map[as.character(GSE115469_so$CellType)])
cat("\nGSE115469 liver mapping:\n")
print(table(GSE115469_so$CellType, GSE115469_so$benchmark_celltype, useNA = "always"))

# Remove only genuinely unmatched cells (Erythroid_Cells).
GSE115469_so_benchmark <- subset(
  GSE115469_so,
  cells = colnames(GSE115469_so)[!is.na(GSE115469_so$benchmark_celltype)]
)

liver_markers <- extract_top_clustocell_markers(
  liver_tumor_adj_manual_annot_markoClust,
  top_n = TOP_N
)
# Marker-list names are expected to equal the six liver_tumor_adj labels.
liver_allowed <- c("Cholangiocyte", "Endothelial", "Hepatocyte",
                   "Lymphocyte", "Myeloid", "SMC")
liver_markers <- liver_markers[intersect(liver_allowed, names(liver_markers))]

```

### Prostate prep

```{r}

# ==============================================================================
# 7. PROSTATE LABEL MAPPING AND HARMONIZED REFERENCE (GSE185344)
# ==============================================================================

prostate_query_map <- c(
  "B cells"                    = "B",
  "Basal cells"                = "BE",
  "Ciliated cells"             = NA_character_,
  "Endothelial cells"          = "Endothelial",
  "Epithelial cells"           = "LE",
  "Fibroblasts"                = "Fibroblast",
  "Macrophages"                = "MNP",
  "Mast cells"                 = "Mast",
  "NK cells"                   = "NK",
  "Olfactory epithelial cells" = NA_character_,
  "Schwann cells"              = NA_character_,
  "Smooth muscle cells"        = NA_character_,
  "T cells"                    = "T"
)

GSE185344_so$benchmark_celltype <- unname(
  prostate_query_map[
    as.character(GSE185344_so$cellactivity_clusters)
  ]
)

cat("\nGSE185344 prostate mapping:\n")

print(
  table(
    GSE185344_so$cellactivity_clusters,
    GSE185344_so$benchmark_celltype,
    useNA = "always"
  )
)

# Remove only unmatched cell types
GSE185344_so_benchmark <- subset(
  GSE185344_so,
  cells = colnames(GSE185344_so)[
    !is.na(GSE185344_so$benchmark_celltype)
  ]
)

# ------------------------------------------------------------------------------
# Harmonize the Prostate Cell Atlas reference labels to the same resolution
# ------------------------------------------------------------------------------

prostate_ref_map <- c(
  "B"           = "B",
  "BE"          = "BE",
  "CE"          = NA_character_,
  "Endothelial" = "Endothelial",
  "Fibroblast"  = "Fibroblast",
  "HE"          = NA_character_,
  "LE"          = "LE",
  "Mast"        = "Mast",
  "MNP"         = "MNP",
  "NK"          = "NK",
  "Sperm"       = NA_character_,
  "T"           = "T"
)

prostateimmuneatlas$benchmark_celltype <- unname(
  prostate_ref_map[
    as.character(prostateimmuneatlas$manual_annot)
  ]
)

cat("\nProstate Cell Atlas reference harmonization:\n")

print(
  table(
    prostateimmuneatlas$manual_annot,
    prostateimmuneatlas$benchmark_celltype,
    useNA = "always"
  )
)

prostateimmuneatlas_benchmark <- subset(
  prostateimmuneatlas,
  cells = colnames(prostateimmuneatlas)[
    !is.na(prostateimmuneatlas$benchmark_celltype)
  ]
)

# ------------------------------------------------------------------------------
# ClustoCell markers
# ------------------------------------------------------------------------------

prostate_markers_original <- extract_top_clustocell_markers(
  prostateimmuneatlas_manual_annot_markoClust,
  top_n = TOP_N
)

# Merge BE, CE, HE and LE markers into one epithelial marker list
prostate_markers <- combine_marker_classes(
  prostate_markers_original,
  prostate_ref_map
)

```

### Run benchmarking

```{r}

# ==============================================================================
# 8. RUN BOTH DATASET BENCHMARKS
# ==============================================================================
liver_benchmark_results <- run_dataset_benchmark(
  dataset_name = "Liver_GSE115469",
  reference_obj = liver_tumor_adj,
  reference_label_col = "minor_manual_annot1",
  query_obj = GSE115469_so_benchmark,
  query_label_col = "benchmark_celltype",
  markers_list = liver_markers,
  output_root = BASE_OUT_DIR
)

saveRDS(liver_benchmark_results, "../../../Benchmarking/Deconvolution/Liver_Prostate/Liver_GSE115469/liver_benchmark_results.rds")

liver_benchmark_results <- readRDS("../../../Benchmarking/Deconvolution/Liver_Prostate/Liver_GSE115469/liver_benchmark_results.rds")

liver_BayesPrism_p_forest <- liver_benchmark_results$BayesPrism_plots$forest + 
  labs(tag = "a") + theme(plot.tag = element_text(size = 7, face = "bold"))

liver_BayesPrism_p_ccc <- liver_benchmark_results$BayesPrism_plots$ccc + 
  labs(tag = "b") + theme(plot.tag = element_text(size = 7, face = "bold"))

liver_BayesPrism_p_rmse <- liver_benchmark_results$BayesPrism_plots$rmse + 
  labs(tag = "c") + theme(plot.tag = element_text(size = 7, face = "bold"))

#__________

liver_DWLS_p_forest <- liver_benchmark_results$DWLS_plots$forest + 
  labs(tag = "d") + theme(plot.tag = element_text(size = 7, face = "bold"))

liver_DWLS_p_ccc <- liver_benchmark_results$DWLS_plots$ccc + 
  labs(tag = "e") + theme(plot.tag = element_text(size = 7, face = "bold"))

liver_DWLS_p_rmse <- liver_benchmark_results$DWLS_plots$rmse + 
  labs(tag = "f") + theme(plot.tag = element_text(size = 7, face = "bold"))

deconv_liver_row_combined <- 
  (liver_BayesPrism_p_forest + 
  liver_BayesPrism_p_ccc + 
  liver_BayesPrism_p_rmse) /
  (liver_DWLS_p_forest + 
  liver_DWLS_p_ccc + 
  liver_DWLS_p_rmse)

#______________________________________________

prostate_benchmark_results <- run_dataset_benchmark(
  dataset_name = "Prostate_GSE185344",
  reference_obj = prostateimmuneatlas_benchmark,
  reference_label_col = "benchmark_celltype",
  query_obj = GSE185344_so_benchmark,
  query_label_col = "benchmark_celltype",
  markers_list = prostate_markers,
  output_root = BASE_OUT_DIR
)

saveRDS(
  prostate_benchmark_results,
  "../../../Benchmarking/Deconvolution/Liver_Prostate/Prostate_GSE185344/prostate_benchmark_results.rds"
)

prostate_benchmark_results <- readRDS("../../../Benchmarking/Deconvolution/Liver_Prostate/Prostate_GSE185344/prostate_benchmark_results.rds")

prostate_BayesPrism_p_forest <- prostate_benchmark_results$BayesPrism_plots$forest + 
  labs(tag = "g") + theme(plot.tag = element_text(size = 7, face = "bold"))

prostate_BayesPrism_p_ccc <- prostate_benchmark_results$BayesPrism_plots$ccc + 
  labs(tag = "h") + theme(plot.tag = element_text(size = 7, face = "bold"))

prostate_BayesPrism_p_rmse <- prostate_benchmark_results$BayesPrism_plots$rmse + 
  labs(tag = "i") + theme(plot.tag = element_text(size = 7, face = "bold"))

#__________

prostate_DWLS_p_forest <- prostate_benchmark_results$DWLS_plots$forest + 
  labs(tag = "j") + theme(plot.tag = element_text(size = 7, face = "bold"))

prostate_DWLS_p_ccc <- prostate_benchmark_results$DWLS_plots$ccc + 
  labs(tag = "k") + theme(plot.tag = element_text(size = 7, face = "bold"))

prostate_DWLS_p_rmse <- prostate_benchmark_results$DWLS_plots$rmse + 
  labs(tag = "l") + theme(plot.tag = element_text(size = 7, face = "bold"))

deconv_prostate_row_combined <- 
  (prostate_BayesPrism_p_forest + 
     prostate_BayesPrism_p_ccc + 
     prostate_BayesPrism_p_rmse) /
  (prostate_DWLS_p_forest + 
     prostate_DWLS_p_ccc + 
     prostate_DWLS_p_rmse)


```

# Combine the plots

```{r}

deconv_comb_plt <- (wrap_elements(deconv_liver_row_combined) / wrap_elements(deconv_prostate_row_combined)) & theme(plot.margin = margin(0, 0, 0, 0))

ggsave(filename = "../../../Testing results/deconv_comb_plt.pdf", 
       plot = deconv_comb_plt, 
       device = "pdf", 
       width = 6.85, height = 1.762*5, units = "in")

```

