---
title: "Evaluating ClustoCell's Performance in finding novel biology"
author: "Adrian Salavaty"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(pROC)
library(umap)
library(cluster)
library(ggplot2)
library(dplyr)
library(splatter)
library(Seurat)
library(tidyverse)
library(scplotter)
library(scran)
library(scater)
library(Matrix)
library(SeuratDisk)
library(hdf5r)
library(patchwork)
library(paletteer)
library(hrbrthemes)
library(scales)

```

# Introduction

Script for evaluation of ClustoCell's Performance in finding novel biology.

# Function Defintion

We first define some helper functions.

```{r}

# Personalized Theme function

styled_theme <- function(
  style = c("academic", "minimal", "modern", "dark", "classic", "void"),
  border_mode = c("default", "all", "none"),
  title_align = c("center", "left", "right"),
  legends_to_drop = NULL,
  axes_title_drop = FALSE,
  legend_position = c("default", "top", "bottom", "left", "right", "inside"),
  axis_title_face = c("regular", "bold"),
  x_axis_text_angle = 45,
  x_axis_text_y = -0.3,
  remove_margin = c("none", "bottom", "all", "top", "left", "right")
) {
  style <- match.arg(style)
  border_mode <- match.arg(border_mode)
  title_align <- match.arg(title_align)
  legend_position <- match.arg(legend_position)
  axis_title_face <- match.arg(axis_title_face)
  remove_margin <- match.arg(remove_margin)

  base_font <- "Helvetica"
  align_map <- c(left = 0, center = 0.5, right = 1)
  hjust_val <- align_map[title_align]
  thin_size <- 0.3  

  # Decide axis title face
  axis_face_val <- ifelse(axis_title_face == "bold", "bold", "plain")

  # Base theme selection
  if (style == "void") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        axis.title.x = element_blank(),  
        axis.title.y = element_blank(),  
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "academic") {
    theme_out <- theme_minimal(base_size = 14, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = hjust_val),
        plot.subtitle = element_text(size = 13, hjust = hjust_val, color = "grey30"),
        plot.caption = element_text(size = 10, hjust = 1, color = "grey40"),
        axis.title = element_text(face = axis_face_val, size = 13),
        axis.text = element_text(size = 12, color = "black"),
        panel.grid.major = element_line(color = "grey80", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "minimal") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", size = thin_size),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.text = element_text(color = "black"),
        axis.title = element_text(face = axis_face_val),
        legend.position = "none",
        legend.title = element_text(face = "italic")
      )
  } else if (style == "modern") {
    theme_out <- theme_light(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.border = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "dark") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        plot.background = element_rect(fill = "black", color = NA),
        panel.background = element_rect(fill = "black", color = NA),
        panel.grid.major = element_line(color = "grey30", size = thin_size),
        panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 15, color = "white", hjust = hjust_val),
        plot.subtitle = element_text(color = "grey80", hjust = hjust_val),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white", face = axis_face_val),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white", face = "italic"),
        axis.line = element_line(color = "white", size = thin_size)
      )
  } else if (style == "classic") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  }

  # Apply border mode
  if (border_mode == "all") {
    theme_out <- theme_out + theme(panel.border = element_rect(color = "black", fill = NA, size = thin_size))
  } else if (border_mode == "none") {
    theme_out <- theme_out + theme(panel.border = element_blank(), axis.line = element_blank())
  }

  # Drop axes titles if requested
  if (axes_title_drop) {
    theme_out <- theme_out + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
  }

  # Override x-axis text angle and vertical position
  theme_out <- theme_out + theme(
    axis.text.x = element_text(
      angle = x_axis_text_angle,
      vjust = 1,
      hjust = 1,
      margin = margin(t = x_axis_text_y)
    )
  )

  # Handle margin removal
  if (remove_margin == "all") {
    theme_out <- theme_out + theme(plot.margin = margin(0, 0, 0, 0))
  } else if (remove_margin == "top") {
    theme_out <- theme_out + theme(plot.margin = margin(t = 0))
  } else if (remove_margin == "bottom") {
    theme_out <- theme_out + theme(plot.margin = margin(b = 0))
  } else if (remove_margin == "left") {
    theme_out <- theme_out + theme(plot.margin = margin(l = 0))
  } else if (remove_margin == "right") {
    theme_out <- theme_out + theme(plot.margin = margin(r = 0))
  }
  # if "none", do nothing

  # Legend positioning override
  if (legend_position != "default") {
    if (legend_position == "inside") {
      theme_out <- theme_out + theme(
        legend.position = c(0.98, 0.98),
        legend.justification = c("right", "top"),
        legend.direction = "vertical"
      )
    } else {
      horiz <- legend_position %in% c("top", "bottom")
      theme_out <- theme_out + theme(
        legend.position = legend_position,
        legend.direction = if (horiz) "horizontal" else "vertical"
      )
    }
  }

  # Return theme + guides if needed
  if (!is.null(legends_to_drop)) {
    drop_list <- lapply(legends_to_drop, function(x) "none")
    names(drop_list) <- legends_to_drop
    return(list(theme_out, guides(.list = drop_list)))
  } else {
    return(theme_out)
  }
}

#____________

# Label plot
label_plt <- function(label = "Label", size = 1) {
  ggplot(data.frame(x = 1, y = 1), aes(x = x, y = y)) +
    # Draw a single vertical bar
    geom_col(width = 0.6, fill = "lightgrey", color = "black") +
    # Add vertical text (angle=90)
    annotate("text", x = 1, y = 0.5, label = label, size = size,
             angle = 90, size = 6, vjust = 0.5, hjust = 0.5) +
    # Remove all extra space
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme_void(base_size = 14) +             # no axes, ticks, labels
    theme(
        plot.margin = margin(0, 0, 0, 0),      # remove outside margins
        panel.spacing = unit(0, "pt"),         # remove panel padding
        plot.background = element_blank(),     # no background
        panel.background = element_blank()
    )
}

#__________________________________________________________________

# Box plot with stats

make_facet_boxplot_with_stats <- function(
    df,
    x = "group",
    y = "value",
    fill = NULL,
    palette = NULL,
    facet.by = NULL,
    scales = "free_y",
    facet.nrow = NULL,       
    facet.ncol = NULL,      
    alpha = 0.8,
    xlab = NULL,
    ylab = NULL,
    subtitle = NULL,
    tag = NULL,
    jitter.width = 0.15,
    jitter.size = 0.35,
    jitter.alpha = 0.6,
    method = "wilcox.test",
    p.adjust.method = "none",
    hide.ns = FALSE,
    base_size = 10
) {

  library(dplyr)
  library(ggpubr)
  library(rstatix)
  library(ggplot2)

  # ----------------------------
  # 1. Statistical test per facet
  # ----------------------------
  stat.test <- df %>%
    group_by(!!sym(facet.by)) %>%
    rstatix::wilcox_test(as.formula(paste(y, "~", x))) %>%
    adjust_pvalue(method = p.adjust.method) %>%
    add_significance()

  # Add xy positions (critical for facetted layout)
  stat.test <- stat.test %>% add_xy_position(x = x)

  # ----------------------------
  # 2. Build ggpubr boxplot
  # ----------------------------
  p <- ggpubr::ggboxplot(
    df,
    x = x,
    y = y,
    fill = fill,
    alpha = alpha,
    palette = palette,
    facet.by = facet.by,
    scales = scales,
    nrow = facet.nrow,  
    ncol = facet.ncol,   
    outlier.shape = NA
  ) +
    geom_jitter(
      width = jitter.width,
      size = jitter.size,
      alpha = jitter.alpha
    ) +
    labs(
      x = xlab,
      y = ylab,
      subtitle = subtitle,
      tag = tag
    ) +
    theme_classic(base_size = base_size) +
    theme(
      plot.subtitle = element_text(size = 7),
      text = element_text(size = 6),
      strip.text = element_text(size = 5.5),
      plot.tag = element_text(size = 7, face = "bold"),
      axis.title.y = element_text(size = 7),
      plot.margin = margin(0, 5, 0, 5),
      legend.position = "none"
    )

  # ----------------------------------
  # 3. Add p-value stars to the plot
  # ----------------------------------
  p <- p + stat_pvalue_manual(
    stat.test,
    label = "p.adj.signif",
    hide.ns = hide.ns,
    size = 2.2,
    tip.length = 0.01
  )

  return(p)
}


```


```{r}

# Concordance = 100*Sum(rowMax)/N

library(paletteer)

paste(as.vector(col2rgb("#0000CCFF")), collapse = " ")

# https://pmassicotte.github.io/paletteer_gallery/

paletteer::paletteer_d("MetBrewer::Austria", 7)
#A40000FF #16317DFF #007E2FFF #FFCD12FF #B86092FF #721B3EFF #00B7A7FF 

plotthis::palette_list %>% names()

# Best pals from ggsci: pal_npg (10), pal_d3 (10), pal_ucscgb (26), pal_igv (51)

# "#FF0000FF" "#FF9900FF" "#FFCC00FF" "#00FF00FF" "#6699FFFF" "#CC33FFFF" "#99991EFF" "#999999FF" "#FF00CCFF" "#CC0000FF" "#FFCCCCFF" "#FFFF00FF" "#CCFF00FF" "#358000FF" "#0000CCFF" "#99CCFFFF" "#00FFFFFF"
# "#CCFFFFFF" "#9900CCFF" "#CC99FFFF" "#996600FF" "#666600FF" "#666666FF" "#CCCCCCFF" "#79CC3DFF" "#CCCC99FF"

# "#1F77B4FF" "#FF7F0EFF" "#2CA02CFF" "#D62728FF" "#9467BDFF" "#8C564BFF" "#E377C2FF" "#7F7F7FFF" "#BCBD22FF" "#17BECFFF"

# NPG colors:
# "#E64B35FF" "#4DBBD5FF" "#00A087FF" "#3C5488FF" "#F39B7FFF" "#8491B4FF" "#91D1C2FF" "#DC0000FF" "#7E6148FF" "#B09C85FF"

# "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF" "#924822FF" "#837B8DFF" "#C75127FF" "#D58F5CFF" "#7A65A5FF" "#E4AF69FF" "#3B1B53FF"
# "#CDDEB7FF" "#612A79FF" "#AE1F63FF" "#E7C76FFF" "#5A655EFF" "#CC9900FF" "#99CC00FF" "#A9A9A9FF" "#CC9900FF" "#99CC00FF" "#33CC00FF" "#00CC33FF" "#00CC99FF" "#0099CCFF" "#0A47FFFF" "#4775FFFF" "#FFC20AFF"
# "#FFD147FF" "#990033FF" "#991A00FF" "#996600FF" "#809900FF" "#339900FF" "#00991AFF" "#009966FF" "#008099FF" "#003399FF" "#1A0099FF" "#660099FF" "#990080FF" "#D60047FF" "#FF1463FF" "#00D68FFF"

 # "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF"

# Define the color palettes

npg_pal <- ggsci::pal_npg()(10)
npg_reverse_pal <- rev(ggsci::pal_npg("nrc")(10))

```

A function for converting count df to sparse matrix

```{r}

library(Rcpp)
library(RcppArmadillo)

Rcpp::sourceCpp(code = '
#define ARMA_64BIT_WORD 1
#include <RcppArmadillo.h>
// [[Rcpp::depends(RcppArmadillo)]]

// [[Rcpp::export]]
Rcpp::S4 denseToSparse(Rcpp::DataFrame& df) {  // Pass by reference to allow in-place modification
  int nrows = df.nrows();
  int ncols = df.size();

  arma::sp_mat sparse(nrows, ncols);

  for (int col = 0; col < ncols; ++col) {
    arma::vec this_col = Rcpp::as<arma::vec>(df[col]);  // Wrap without copying
    sparse.col(col) = this_col;  // Assigns only non-zeros to sparse
    df[col] = R_NilValue;  // Free the column memory progressively
  }

  return Rcpp::wrap(sparse);  // Returns as dgCMatrix S4 object
}
')

```

Concordance = 100*Sum(rowMax)/N

# Breast Cancer Single-Cell Map (bcscm)

## Data preparation

The data was downloaded from https://lambrechtslab.sites.vib.be/en/single-cell
https://www.nature.com/articles/s41591-021-01323-8

### T-/NK cells

```{r}

bcscm_tcell_counts <- read_rds("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/1864-counts_tcell_cohort1.rds")
bcscm_tcell_meta <- fread("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/1870-BIOKEY_metaData_tcells_cohort1_web.csv")

ncol(bcscm_tcell_counts) == nrow(bcscm_tcell_meta) # TRUE
all(colnames(bcscm_tcell_counts) == bcscm_tcell_meta$Cell) # FALSE
all(colnames(bcscm_tcell_counts) %in% bcscm_tcell_meta$Cell) # TRUE
bcscm_tcell_meta <- bcscm_tcell_meta[match(colnames(bcscm_tcell_counts), bcscm_tcell_meta$Cell),]
all(colnames(bcscm_tcell_counts) == bcscm_tcell_meta$Cell) # TRUE

bcscm_tcell_so <- CreateSeuratObject(bcscm_tcell_counts, meta.data = bcscm_tcell_meta, min.cells = 3, min.features = 200)
qs2::qs_save(object = bcscm_tcell_so, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")

bcscm_tcell_so$cellSubType <- gsub("Proliferating", "Prolif", bcscm_tcell_so$cellSubType)

bcscm_tcell_so <- NormalizeData(bcscm_tcell_so, normalization.method = "LogNormalize", scale.factor = 10000)
bcscm_tcell_so <- FindVariableFeatures(bcscm_tcell_so, selection.method = "vst", nfeatures = 2000)

qs2::qs_save(object = bcscm_tcell_so, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")

bcscm_tcell_ClustoCell <- clustoCell(data = bcscm_tcell_so, sketch = T, sketch_ncells = 25000, refine_transferred_subClusters = TRUE, 
                                     high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)

qs2::qs_save(object = bcscm_tcell_ClustoCell, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_ClustoCell.qs2")

bcscm_tcell_so <- addClustoData(obj = bcscm_tcell_so, clustoCell = bcscm_tcell_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

# Finding the dataset-level markers and using them for dim reduction
bcscm_tcell_markers <- getDatasetMarkers(obj = bcscm_tcell_ClustoCell, 
                                         clusters = T, sub_clusters = T,
                                         positive_markers = T, negative_markers = T, 
                                         medium_markers = F, 
                                         thresh_mode = "n", pos_thresh = 200, neg_thresh = 50)

bcscm_tcell_so <- ScaleData(bcscm_tcell_so, features = bcscm_tcell_markers$combined_markers)
bcscm_tcell_so <- RunPCA(bcscm_tcell_so, features = bcscm_tcell_markers$combined_markers)
ElbowPlot(bcscm_tcell_so, ndims = 50)
# bcscm_tcell_so <- FindNeighbors(bcscm_tcell_so, dims = 1:16)
# bcscm_tcell_so <- FindClusters(bcscm_tcell_so, resolution = 0.5)
bcscm_tcell_so <- RunUMAP(bcscm_tcell_so, dims = 1:16)

qs2::qs_save(object = bcscm_tcell_so, file = "../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")

bcscm_tcell_so <- qs2::qs_read("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_so.qs2")
bcscm_tcell_ClustoCell <- qs2::qs_read("../../../Benchmarking/Test Datasets/breast cancer single-cell Map/bcscm_tcell_ClustoCell.qs2")

```

#### Inspecting the data

Checking the metadata and confusion tables

```{r}

head(bcscm_tcell_so@meta.data %>% dplyr::select(-Cell))

# ClustoCell clusters and subclusters
table(bcscm_tcell_so$ClustoCell)
table(bcscm_tcell_so$SubClustoCell)

# Author annotations
table(bcscm_tcell_so@meta.data$cellSubType)
table(bcscm_tcell_so@meta.data$expansion)
table(bcscm_tcell_so@meta.data$timepoint)
table(bcscm_tcell_so@meta.data$BC_type)

# Confusion of author annotations with ClustoCell
table(bcscm_tcell_so$cellSubType, bcscm_tcell_so$ClustoCell)
table(bcscm_tcell_so$ClustoCell, bcscm_tcell_so$expansion)
table(bcscm_tcell_so$ClustoCell, bcscm_tcell_so$BC_type)
table(bcscm_tcell_so$ClustoCell, bcscm_tcell_so$timepoint)

# Confusion of author annotations with SubClustoCell
table(bcscm_tcell_so$cellSubType, bcscm_tcell_so$SubClustoCell)
table(bcscm_tcell_so$SubClustoCell, bcscm_tcell_so$expansion)
table(bcscm_tcell_so$SubClustoCell, bcscm_tcell_so$BC_type)
table(bcscm_tcell_so$SubClustoCell, bcscm_tcell_so$timepoint)

```

Checking the markers of clusters and subclusters

```{r}

# Major Clusters
sapply(names(bcscm_tcell_ClustoCell$markers$major_clusters$cluster_specific$positive_markers), 
       function(i) { 
         tmp_markers <- bcscm_tcell_ClustoCell$markers$major_clusters$cluster_specific$positive_markers[[i]] %>% 
           dplyr::slice(1:30) %>% dplyr::pull(Feature)
         cat(paste0(i, ": "))
          cat(paste0(tmp_markers, collapse = ", "))
          cat("\n\n")
          }
       )

# Sub-clusters
sapply(names(bcscm_tcell_ClustoCell$markers$sub_clusters), 
       function(i) { 
         cat(paste0(i, ": "))
         cat("\n")

         sapply(names(bcscm_tcell_ClustoCell$markers$sub_clusters[[i]]$positive_markers), function(j) {
           tmp_markers <- (bcscm_tcell_ClustoCell$markers$sub_clusters[[i]]$positive_markers)[[j]] %>% 
             dplyr::slice(1:30) %>% dplyr::pull(Feature)
           cat(paste0(j, ": "))
           cat(paste0(tmp_markers, collapse = ", "))
           cat("\n\n")
           })
         cat("\n\n\n")
         }
       )

```

```{r}

# Define expanded marker signatures (original + reviewer-suggested genes)
sig_panels_bcscm_tcell <- list(
  `C4-Sub4` = c("PDCD1", "LAG3", "CTLA4", "TIGIT", "CXCL13", "TNFRSF18", "CTSD", "PRDM1",
                "CXCR5", "BCL6", "CD40LG", "TNFRSF4", "IL21", "ICOS", "CCR2", "CCR5", "MAF"),
  `C5-Sub2` = c("CD8A", "GZMB", "LAG3", "TIGIT", "CXCL13", "PRF1", "GZMH", "NKG7",
                "HLA-DRA", "HLA-DRB1", "HLA-DRB5", "IL2RA", "TNFRSF4", "ICOS",
                "NR4A1", "BCL2L1", "BCL2A1", "BCL2", "TCF7", "LEF1"),
  `C5-Sub3` = c("CD8A", "GZMB", "LAG3", "CXCL13", "GNLY", "PRF1", "GZMH", "NKG7",
                "CCL4", "HLA-DRA", "HLA-DRB1", "HLA-DRB5", "TOX", "ENTPD1",
                "CD101", "CXCR6", "GZMK", "CX3CR1"),
  `C2-Sub1` = c("GZMK", "KLRB1", "CXCL13", "ANXA1", "MTRNR2L8",
                "CD4", "IL7R", "ITGB1", "CD40LG")
)

# Compute average expression of all markers in each subcluster
sig_genes <- unique(unlist(sig_panels_bcscm_tcell))
sig_panels_bcscm_tcell_avg_expr <- AverageExpression(
  bcscm_tcell_pre,
  group.by = "SubClustoCell",
  features = sig_genes
)$RNA

# Restrict to the four predictive subclusters
predictive_subs <- c("C4-Sub4", "C5-Sub2", "C5-Sub3", "C2-Sub1")
sig_panels_bcscm_tcell_avg_expr <- sig_panels_bcscm_tcell_avg_expr[, predictive_subs]

# Prepare row annotation for the heatmap
sig_panels_bcscm_tcell_gene_order <- unlist(sig_panels_bcscm_tcell)
sig_panels_bcscm_tcell_group_labels <- rep(names(sig_panels_bcscm_tcell),
                                           lengths(sig_panels_bcscm_tcell))
sig_panels_bcscm_tcell_ann_row <- data.frame(
  Gene = sig_panels_bcscm_tcell_gene_order,
  Signature = sig_panels_bcscm_tcell_group_labels
)
# Keep unique genes with their first-assigned signature
sig_panels_bcscm_tcell_ann_row <- sig_panels_bcscm_tcell_ann_row %>%
  dplyr::distinct(Gene, .keep_all = TRUE)
rownames(sig_panels_bcscm_tcell_ann_row) <- sig_panels_bcscm_tcell_ann_row$Gene

# Scale each gene (row) across subclusters
expr_matrix <- sig_panels_bcscm_tcell_avg_expr[sig_panels_bcscm_tcell_ann_row$Gene, ]
sig_panels_bcscm_tcell_sig_scaled <- t(scale(t(expr_matrix)))

# Ensure the scaled matrix rows match the annotation rows
sig_panels_bcscm_tcell_sig_scaled <- sig_panels_bcscm_tcell_sig_scaled[rownames(sig_panels_bcscm_tcell_ann_row), ]

# Convert signature to factor to preserve panel order
sig_panels_bcscm_tcell_ann_row$Signature <- factor(
  sig_panels_bcscm_tcell_ann_row$Signature,
  levels = names(sig_panels_bcscm_tcell)
)

# Load ggheatmap (from ggalign package) and plot
library(ggalign)
sig_panels_bcscm_tcell_ggheatmap <- ggheatmap(sig_panels_bcscm_tcell_sig_scaled) +
  scale_fill_gradient2(
    low = "navy", mid = "white", high = "firebrick3", midpoint = 0,
    name = "Scaled\nExpr."
  ) +
  labs(subtitle = "Selected Subcluster Markers (Expanded Panel)", tag = NULL) +
  theme(
    plot.margin = margin(t = 0, r = 0, b = 0, l = 0),
    axis.title.y = element_blank(),
    axis.text.x = element_text(angle = 25, vjust = 0.7, face = "plain", color = "black"),
    axis.text = element_text(size = 6),
    text = element_text(size = 7.5),
    title = element_text(size = 7.5),
    plot.title = element_text(size = 7.5),
    plot.subtitle = element_text(size = 7.5, vjust = 0),
    axis.title = element_text(size = 7),
    plot.tag = element_text(size = 7.5, face = "bold"),
    legend.title = element_text(size = 6),
    legend.margin = margin(t = 0, r = 0, b = 0, l = 0),
    legend.box.margin = margin(0, 0, 0, 0),
    legend.box.spacing = unit(0, "pt"),
    legend.key.width = unit(0.25, "cm"),
    legend.key.height = unit(0.4, "cm")
  )

```

## Combining the plots for main figure

```{r}

sig_panels_bcscm_tcell_ggheatmap_final <- wrap_elements(ggalign::ggalignGrob(sig_panels_bcscm_tcell_ggheatmap)) +
  # ggtitle("Supplementary Figure 23: Expanded canonical-marker panel resolves the cell-state identity of \nthe four expansion-enriched predictive subclusters and supports their refined annotation.") +
  theme(plot.title = element_text(size = 8))

ggsave(filename = "../../../Testing results/Supplementary Figure 21.pdf", 
       plot = sig_panels_bcscm_tcell_ggheatmap_final, device = "pdf", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/Supplementary Figure 21.png", 
       plot = sig_panels_bcscm_tcell_ggheatmap_final, device = "png", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)


```
