---
title: "Evaluating ClustoCell's Robustness to Data Normalization and Batch Effects"
author: "Adrian Salavaty"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(pROC)
library(umap)
library(cluster)
library(ggplot2)
library(dplyr)
library(splatter)
library(Seurat)
library(tidyverse)
library(scplotter)
library(scran)
library(scater)
library(paletteer)

```

# Introduction

Script for evaluation the tolerance of ClustoCell to data normalization and presence of batch effects and unwanted variations in the data.

# Function Defintion

We first define some helper functions.

```{r}

# Personalized Theme function

styled_theme <- function(
  style = c("academic", "minimal", "modern", "dark", "classic", "void"),
  border_mode = c("default", "all", "none"),
  title_align = c("center", "left", "right"),
  legends_to_drop = NULL,
  axes_title_drop = FALSE,
  legend_position = c("default", "top", "bottom", "left", "right", "inside"),
  axis_title_face = c("regular", "bold"),
  x_axis_text_angle = 45,
  x_axis_text_y = -0.3,
  remove_margin = c("none", "bottom", "all", "top", "left", "right")
) {
  style <- match.arg(style)
  border_mode <- match.arg(border_mode)
  title_align <- match.arg(title_align)
  legend_position <- match.arg(legend_position)
  axis_title_face <- match.arg(axis_title_face)
  remove_margin <- match.arg(remove_margin)

  base_font <- "Helvetica"
  align_map <- c(left = 0, center = 0.5, right = 1)
  hjust_val <- align_map[title_align]
  thin_size <- 0.3  

  # Decide axis title face
  axis_face_val <- ifelse(axis_title_face == "bold", "bold", "plain")

  # Base theme selection
  if (style == "void") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        axis.title.x = element_blank(),  
        axis.title.y = element_blank(),  
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank(),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "academic") {
    theme_out <- theme_minimal(base_size = 14, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 16, hjust = hjust_val),
        plot.subtitle = element_text(size = 13, hjust = hjust_val, color = "grey30"),
        plot.caption = element_text(size = 10, hjust = 1, color = "grey40"),
        axis.title = element_text(face = axis_face_val, size = 13),
        axis.text = element_text(size = 12, color = "black"),
        panel.grid.major = element_line(color = "grey80", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "right",
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "minimal") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        panel.grid = element_blank(),
        axis.line = element_line(color = "black", size = thin_size),
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.text = element_text(color = "black"),
        axis.title = element_text(face = axis_face_val),
        legend.position = "none",
        legend.title = element_text(face = "italic")
      )
  } else if (style == "modern") {
    theme_out <- theme_light(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.border = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed", size = thin_size),
        panel.grid.minor = element_blank(),
        legend.position = "bottom",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  } else if (style == "dark") {
    theme_out <- theme_minimal(base_size = 13, base_family = base_font) +
      theme(
        plot.background = element_rect(fill = "black", color = NA),
        panel.background = element_rect(fill = "black", color = NA),
        panel.grid.major = element_line(color = "grey30", size = thin_size),
        panel.grid.minor = element_blank(),
        plot.title = element_text(face = "bold", size = 15, color = "white", hjust = hjust_val),
        plot.subtitle = element_text(color = "grey80", hjust = hjust_val),
        axis.text = element_text(color = "white"),
        axis.title = element_text(color = "white", face = axis_face_val),
        legend.background = element_blank(),
        legend.key = element_blank(),
        legend.text = element_text(color = "white"),
        legend.title = element_text(color = "white", face = "italic"),
        axis.line = element_line(color = "white", size = thin_size)
      )
  } else if (style == "classic") {
    theme_out <- theme_classic(base_size = 13, base_family = base_font) +
      theme(
        plot.title = element_text(face = "bold", size = 15, hjust = hjust_val),
        axis.title = element_text(face = axis_face_val),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_text(face = "italic"),
        axis.line = element_line(color = "black", size = thin_size)
      )
  }

  # Apply border mode
  if (border_mode == "all") {
    theme_out <- theme_out + theme(panel.border = element_rect(color = "black", fill = NA, size = thin_size))
  } else if (border_mode == "none") {
    theme_out <- theme_out + theme(panel.border = element_blank(), axis.line = element_blank())
  }

  # Drop axes titles if requested
  if (axes_title_drop) {
    theme_out <- theme_out + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
  }

  # Override x-axis text angle and vertical position
  theme_out <- theme_out + theme(
    axis.text.x = element_text(
      angle = x_axis_text_angle,
      vjust = 1,
      hjust = 1,
      margin = margin(t = x_axis_text_y)
    )
  )

  # Handle margin removal
  if (remove_margin == "all") {
    theme_out <- theme_out + theme(plot.margin = margin(0, 0, 0, 0))
  } else if (remove_margin == "top") {
    theme_out <- theme_out + theme(plot.margin = margin(t = 0))
  } else if (remove_margin == "bottom") {
    theme_out <- theme_out + theme(plot.margin = margin(b = 0))
  } else if (remove_margin == "left") {
    theme_out <- theme_out + theme(plot.margin = margin(l = 0))
  } else if (remove_margin == "right") {
    theme_out <- theme_out + theme(plot.margin = margin(r = 0))
  }
  # if "none", do nothing

  # Legend positioning override
  if (legend_position != "default") {
    if (legend_position == "inside") {
      theme_out <- theme_out + theme(
        legend.position = c(0.98, 0.98),
        legend.justification = c("right", "top"),
        legend.direction = "vertical"
      )
    } else {
      horiz <- legend_position %in% c("top", "bottom")
      theme_out <- theme_out + theme(
        legend.position = legend_position,
        legend.direction = if (horiz) "horizontal" else "vertical"
      )
    }
  }

  # Return theme + guides if needed
  if (!is.null(legends_to_drop)) {
    drop_list <- lapply(legends_to_drop, function(x) "none")
    names(drop_list) <- legends_to_drop
    return(list(theme_out, guides(.list = drop_list)))
  } else {
    return(theme_out)
  }
}

#____________

# Label plot
label_plt <- function(label = "Label", size = 1) {
  ggplot(data.frame(x = 1, y = 1), aes(x = x, y = y)) +
    # Draw a single vertical bar
    geom_col(width = 0.6, fill = "lightgrey", color = "black") +
    # Add vertical text (angle=90)
    annotate("text", x = 1, y = 0.5, label = label, size = size,
             angle = 90, size = 6, vjust = 0.5, hjust = 0.5) +
    # Remove all extra space
    scale_x_continuous(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme_void(base_size = 14) +             # no axes, ticks, labels
    theme(
        plot.margin = margin(0, 0, 0, 0),      # remove outside margins
        panel.spacing = unit(0, "pt"),         # remove panel padding
        plot.background = element_blank(),     # no background
        panel.background = element_blank()
    )
}

```


```{r}

# Concordance = 100*Sum(rowMax)/N

plotthis::palette_list %>% names()

# https://pmassicotte.github.io/paletteer_gallery/

# Best pals from ggsci: pal_npg (10), pal_d3 (10), pal_ucscgb (26), pal_igv (51)

# "#FF0000FF" "#FF9900FF" "#FFCC00FF" "#00FF00FF" "#6699FFFF" "#CC33FFFF" "#99991EFF" "#999999FF" "#FF00CCFF" "#CC0000FF" "#FFCCCCFF" "#FFFF00FF" "#CCFF00FF" "#358000FF" "#0000CCFF" "#99CCFFFF" "#00FFFFFF"
# "#CCFFFFFF" "#9900CCFF" "#CC99FFFF" "#996600FF" "#666600FF" "#666666FF" "#CCCCCCFF" "#79CC3DFF" "#CCCC99FF"

# NPG colors:
# "#E64B35FF" "#4DBBD5FF" "#00A087FF" "#3C5488FF" "#F39B7FFF" "#8491B4FF" "#91D1C2FF" "#DC0000FF" "#7E6148FF" "#B09C85FF"

# "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF" "#924822FF" "#837B8DFF" "#C75127FF" "#D58F5CFF" "#7A65A5FF" "#E4AF69FF" "#3B1B53FF"
# "#CDDEB7FF" "#612A79FF" "#AE1F63FF" "#E7C76FFF" "#5A655EFF" "#CC9900FF" "#99CC00FF" "#A9A9A9FF" "#CC9900FF" "#99CC00FF" "#33CC00FF" "#00CC33FF" "#00CC99FF" "#0099CCFF" "#0A47FFFF" "#4775FFFF" "#FFC20AFF"
# "#FFD147FF" "#990033FF" "#991A00FF" "#996600FF" "#809900FF" "#339900FF" "#00991AFF" "#009966FF" "#008099FF" "#003399FF" "#1A0099FF" "#660099FF" "#990080FF" "#D60047FF" "#FF1463FF" "#00D68FFF"

 # "#5050FFFF" "#CE3D32FF" "#749B58FF" "#F0E685FF" "#466983FF" "#BA6338FF" "#5DB1DDFF" "#802268FF" "#6BD76BFF" "#D595A7FF"

# Define the color palettes

npg_pal <- ggsci::pal_npg()(10)
npg_reverse_pal <- rev(ggsci::pal_npg("nrc")(10))

```

Concordance = 100*Sum(rowMax)/N

# Normalization

```{r}

library(patchwork)
library(scplotter)

# SeuratData::InstallData("pbmc3k")
pbmc <- SeuratData::LoadData("pbmc3k")
pbmc$manual_annot <- pbmc$seurat_annotations

pbmc@meta.data <- pbmc@meta.data %>% 
  dplyr::mutate(major_manual_cellTypes = case_when(
    as.character(manual_annot) %in% c("Naive CD4 T", "Memory CD4 T") ~ "CD4 T",
    as.character(manual_annot) %in% c("B") ~ "B",
    as.character(manual_annot) %in% c("CD8 T", "NK") ~ "CL",
    as.character(manual_annot) %in% c("CD14+ Mono", "FCGR3A+ Mono", "DC") ~ "MNP",
    as.character(manual_annot) %in% c("Platelet") ~ "PLT",
    .default = NA
  ))

pbmc$major_manual_cellTypes <- factor(pbmc$major_manual_cellTypes, levels = c("CD4 T", "B", "CL", "MNP", "PLT"))

pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)
pbmc <- ScaleData(pbmc, features = VariableFeatures(pbmc))
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
pbmc <- RunUMAP(pbmc, dims = 1:10)

pbmc_ClustoCell <- clustoCell(data = pbmc, high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)
pbmc <- addClustoData(obj = pbmc, clustoCell = pbmc_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

# Best pals from ggsci: pal_npg (10), pal_d3 (20), pal_ucscgb (26), pal_igv (51)

# Concordance = 100*Sum(rowMax)/N

```

## Seurat LogNormalize

```{r}

## LogNormalize

pbmc_LogNorm_ClustoCell <- clustoCell(data = pbmc, layer = "data", log1p = F, high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)
pbmc <- addClustoData(obj = pbmc, clustoCell = pbmc_LogNorm_ClustoCell, major_cluster_name = "ClustoCell_LogNorm", sub_cluster_name = "SubClustoCell_LogNorm")

#________

pbmc_ClustoCell_vs_ClustoCell_LogNorm <- table(Raw = pbmc$ClustoCell, Norm = pbmc$ClustoCell_LogNorm)
pbmc_ClustoCell_vs_ClustoCell_LogNorm_maxOverlap <- round(((pbmc_ClustoCell_vs_ClustoCell_LogNorm %>% apply(1, max) %>% sum())/ncol(pbmc))*100)


pbmc_ClustoCell_vs_ClustoCell_LogNorm <- as.data.frame(pbmc_ClustoCell_vs_ClustoCell_LogNorm)
colnames(pbmc_ClustoCell_vs_ClustoCell_LogNorm) <- c("Raw Data", "Seurat LogNorm", "Count")

pbmc_ClustoCell_vs_ClustoCell_LogNorm_plt <- ggplot(pbmc_ClustoCell_vs_ClustoCell_LogNorm, 
                                      aes(x = `Raw Data`, 
                                          y = `Seurat LogNorm`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_vs_ClustoCell_LogNorm$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_vs_ClustoCell_LogNorm_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "a") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                   barwidth = unit(0.2, "cm")))
#__________________________

pbmc_ClustoCell_SeuLogNormvs_manual_annot <- table(Raw = pbmc$ClustoCell_LogNorm, Norm = pbmc$major_manual_cellTypes)
pbmc_ClustoCell_SeuLogNormvs_manual_annot_maxOverlap <- round(((pbmc_ClustoCell_SeuLogNormvs_manual_annot %>% apply(1, max) %>% sum())/ncol(pbmc))*100)


pbmc_ClustoCell_SeuLogNormvs_manual_annot <- as.data.frame(pbmc_ClustoCell_SeuLogNormvs_manual_annot)
colnames(pbmc_ClustoCell_SeuLogNormvs_manual_annot) <- c("Seurat LogNorm", "Manual Annotation", "Count")

pbmc_ClustoCell_SeuLogNormvs_manual_annot_plt <- ggplot(pbmc_ClustoCell_SeuLogNormvs_manual_annot, 
                                                        aes(x = `Manual Annotation`, 
                                                            y = `Seurat LogNorm`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_SeuLogNormvs_manual_annot$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_SeuLogNormvs_manual_annot_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "e") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))

```

## Scran logNormCounts

```{r}

# Scran

library(Seurat)
library(scran)
library(scater)

## Margin 2
### Scran does not distort the structure of the data in terms of expressed and non-expressed genes.

pbmc_ScranLogNorm_counts <- GetAssayData(pbmc, slot = "counts")
pbmc_ScranLogNorm <- SingleCellExperiment(assays = list(counts = pbmc_ScranLogNorm_counts))
pbmc_ScranLogNorm_clusters <- quickCluster(pbmc_ScranLogNorm)
pbmc_ScranLogNorm <- computeSumFactors(pbmc_ScranLogNorm, clusters = pbmc_ScranLogNorm_clusters)
pbmc_ScranLogNorm <- logNormCounts(pbmc_ScranLogNorm)
pbmc_ScranLogNorm_norm_counts <- assay(pbmc_ScranLogNorm, "logcounts")
pbmc_ScranLogNorm <- CreateSeuratObject(counts = pbmc_ScranLogNorm_counts, assay = "RNA", meta.data = pbmc@meta.data)
pbmc_ScranLogNorm[["RNA"]]$data <- pbmc_ScranLogNorm_norm_counts

pbmc_ScranLogNorm_ClustoCell <- clustoCell(data = pbmc_ScranLogNorm, layer = "data", log1p = F, high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)
pbmc_ScranLogNorm <- addClustoData(obj = pbmc_ScranLogNorm, clustoCell = pbmc_ScranLogNorm_ClustoCell, major_cluster_name = "ClustoCell_ScranLogNorm", sub_cluster_name = "SubClustoCell_ScranLogNorm")

#________

pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm <- table(Norm = pbmc_ScranLogNorm$ClustoCell_ScranLogNorm, Raw = pbmc$ClustoCell)
pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm_maxOverlap <- round(((pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm %>% apply(1, max) %>% sum())/ncol(pbmc))*100)

pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm <- as.data.frame(pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm)
colnames(pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm) <- c("Scran LogNorm", "Raw Data", "Count")

pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm_plt <- ggplot(pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm, 
                                                            aes(x = `Raw Data`, 
                                                                y = `Scran LogNorm`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "b") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))

#__________________________

pbmc_ClustoCell_ScranLogNormvs_manual_annot <- table(Raw = pbmc_ScranLogNorm$ClustoCell_ScranLogNorm, Norm = pbmc$major_manual_cellTypes)
pbmc_ClustoCell_ScranLogNormvs_manual_annot_maxOverlap <- round(((pbmc_ClustoCell_ScranLogNormvs_manual_annot %>% apply(1, max) %>% sum())/ncol(pbmc))*100)


pbmc_ClustoCell_ScranLogNormvs_manual_annot <- as.data.frame(pbmc_ClustoCell_ScranLogNormvs_manual_annot)
colnames(pbmc_ClustoCell_ScranLogNormvs_manual_annot) <- c("Scran LogNorm", "Manual Annotation", "Count")

pbmc_ClustoCell_ScranLogNormvs_manual_annot_plt <- ggplot(pbmc_ClustoCell_ScranLogNormvs_manual_annot, 
                                                        aes(x = `Manual Annotation`, 
                                                            y = `Scran LogNorm`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_ScranLogNormvs_manual_annot$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_ScranLogNormvs_manual_annot_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "f") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))

```

## CLR Margin 2

```{r}

# CLR

## Margin 2
### CLR does not distort the structure of the data in terms of expressed and non-expressed genes.

pbmc_CLRNorm_margin2 <- pbmc 
pbmc_CLRNorm_margin2 <- NormalizeData(pbmc_CLRNorm_margin2, normalization.method = "CLR", scale.factor = 10000, margin = 2)

pbmc_CLRNorm_margin2_ClustoCell <- clustoCell(data = pbmc_CLRNorm_margin2, layer = "data", log1p = F, high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)
pbmc_CLRNorm_margin2 <- addClustoData(obj = pbmc_CLRNorm_margin2, clustoCell = pbmc_CLRNorm_margin2_ClustoCell, major_cluster_name = "ClustoCell_CLRNorm_margin2", sub_cluster_name = "SubClustoCell_CLRNorm_margin2")

#________

pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2 <- table(Norm = pbmc_CLRNorm_margin2$ClustoCell_CLRNorm_margin2, Raw = pbmc$ClustoCell)
pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2_maxOverlap <- round(((pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2 %>% apply(1, max) %>% sum())/ncol(pbmc))*100)

pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2 <- as.data.frame(pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2)
colnames(pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2) <- c("CLR (Margin: 2)", "Raw Data", "Count")

pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2_plt <- ggplot(pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2, 
                                                            aes(x = `Raw Data`, 
                                                                y = `CLR (Margin: 2)`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "c") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                   barwidth = unit(0.2, "cm")))

#__________________________

pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot <- table(Raw = pbmc_CLRNorm_margin2$ClustoCell_CLRNorm_margin2, Norm = pbmc$major_manual_cellTypes)
pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot_maxOverlap <- round(((pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot %>% apply(1, max) %>% sum())/ncol(pbmc))*100)


pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot <- as.data.frame(pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot)
colnames(pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot) <- c("CLR (Margin: 2)", "Manual Annotation", "Count")

pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot_plt <- ggplot(pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot, 
                                                          aes(x = `Manual Annotation`, 
                                                              y = `CLR (Margin: 2)`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "g") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))


```

## SCTransform

```{r}

# SCTransform

pbmc <- SCTransform(
  pbmc,
  verbose = TRUE
)

#________

pbmc_SCTransform_ClustoCell <- clustoCell(data = pbmc, assay = "SCT", layer = "counts", log1p = T, 
                                          high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)

pbmc <- addClustoData(obj = pbmc, clustoCell = pbmc_SCTransform_ClustoCell, major_cluster_name = "ClustoCell_SCTransform", sub_cluster_name = "SubClustoCell_SCTransform")

#________

pbmc_ClustoCell_vs_ClustoCell_SCTransform <- table(Norm = pbmc$ClustoCell_SCTransform, Raw = pbmc$ClustoCell)
pbmc_ClustoCell_vs_ClustoCell_SCTransform_maxOverlap <- round(((pbmc_ClustoCell_vs_ClustoCell_SCTransform %>% apply(1, max) %>% sum())/ncol(pbmc))*100)

pbmc_ClustoCell_vs_ClustoCell_SCTransform <- as.data.frame(pbmc_ClustoCell_vs_ClustoCell_SCTransform)
colnames(pbmc_ClustoCell_vs_ClustoCell_SCTransform) <- c("SCTransform", "Raw Data", "Count")

pbmc_ClustoCell_vs_ClustoCell_SCTransform_plt <- ggplot(pbmc_ClustoCell_vs_ClustoCell_SCTransform, 
                                                    aes(x = `Raw Data`, 
                                                        y = `SCTransform`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_vs_ClustoCell_SCTransform$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_vs_ClustoCell_SCTransform_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "d") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                   barwidth = unit(0.2, "cm")))

#__________________________

pbmc_ClustoCell_SCTransformvs_manual_annot <- table(Raw = pbmc$ClustoCell_SCTransform, Norm = pbmc$major_manual_cellTypes)
pbmc_ClustoCell_SCTransformvs_manual_annot_maxOverlap <- round(((pbmc_ClustoCell_SCTransformvs_manual_annot %>% apply(1, max) %>% sum())/ncol(pbmc))*100)


pbmc_ClustoCell_SCTransformvs_manual_annot <- as.data.frame(pbmc_ClustoCell_SCTransformvs_manual_annot)
colnames(pbmc_ClustoCell_SCTransformvs_manual_annot) <- c("SCTransform", "Manual Annotation", "Count")

pbmc_ClustoCell_SCTransformvs_manual_annot_plt <- ggplot(pbmc_ClustoCell_SCTransformvs_manual_annot, 
                                                          aes(x = `Manual Annotation`, 
                                                              y = `SCTransform`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_SCTransformvs_manual_annot$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_SCTransformvs_manual_annot_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "h") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))


```

## Combine normalization plots

```{r}

# Combine normalization plots

pbmc_combined_norm <- 
    (wrap_plots(pbmc_ClustoCell_vs_ClustoCell_LogNorm_plt,
                pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm_plt,
               pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2_plt, 
               pbmc_ClustoCell_vs_ClustoCell_SCTransform_plt,
               guides = "keep", axes = "collect_x", widths = c(1, 1, 1, 1)
               ) + plot_annotation(theme = theme(plot.title = element_text(size = 9)))) /
      (wrap_plots(pbmc_ClustoCell_SeuLogNormvs_manual_annot_plt,
               pbmc_ClustoCell_ScranLogNormvs_manual_annot_plt, 
               pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot_plt, 
               pbmc_ClustoCell_SCTransformvs_manual_annot_plt,
               guides = "keep", axes = "collect_x", widths = c(1, 1, 1, 1)
               ) + plot_annotation(theme = theme(plot.title = element_text(size = 9))))

saveRDS(object = pbmc_combined_norm, file = "../../../Testing results/Normalization/pbmc_combined_norm.rds")

ggsave(filename = "../../../Testing results/Normalization/pbmc_combined_norm.png", 
       plot = pbmc_combined_norm, device = "png", width = 6.85, height = (1.762)*2, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/Normalization/pbmc_combined_norm.pdf", 
       plot = pbmc_combined_norm, device = "pdf", width = 6.85, height = (1.762)*2, units = "in", dpi = 300)

#__________________________________

# Updated version

combined_norm_res_tbl <- data.frame(
    Method = c(
        "Seurat LogNorm", 
        "Seurat LogNorm",
        "Scran LogNorm",
        "Scran LogNorm",
        "CLR (Margin: 2)",
        "CLR (Margin: 2)",
        "SCTransform",
        "SCTransform"
        ),
    Concordance = c(
        pbmc_ClustoCell_vs_ClustoCell_LogNorm_maxOverlap,
        pbmc_ClustoCell_SeuLogNormvs_manual_annot_maxOverlap,
        pbmc_ClustoCell_vs_ClustoCell_ScranLogNorm_maxOverlap,
        pbmc_ClustoCell_ScranLogNormvs_manual_annot_maxOverlap,
        pbmc_ClustoCell_vs_ClustoCell_CLRNorm_margin2_maxOverlap,
        pbmc_ClustoCell_CLRNorm_margin2vs_manual_annot_maxOverlap,
        pbmc_ClustoCell_vs_ClustoCell_SCTransform_maxOverlap,
        pbmc_ClustoCell_SCTransformvs_manual_annot_maxOverlap
        ),
    Comparison = c(
        "Raw Data",
        "Manual Annotation",
        "Raw Data",
        "Manual Annotation",
        "Raw Data",
        "Manual Annotation",
        "Raw Data",
        "Manual Annotation"
    )
)

qs2::qs_save(object = pbmc, "../../../../Benchmarking/pbmc_for_nrom_eval.qs2")
qs2::qs_save(object = combined_norm_res_tbl, "../../../../Benchmarking/combined_norm_res_tbl.qs2")

pbmc <- qs2::qs_read("../../../../Benchmarking/pbmc_for_nrom_eval.qs2")
combined_norm_res_tbl <- qs2::qs_read("../../../../Benchmarking/combined_norm_res_tbl.qs2")

#_______

combined_norm_res_plt <-
  combined_norm_res_tbl %>%
    pivot_wider(names_from = Comparison,
                values_from = Concordance) %>%
    ggplot(aes(y = reorder(Method, `Raw Data`))) +
    geom_segment(aes(x = `Manual Annotation`,
                     xend = `Raw Data`,
                     yend = Method),
                 colour = "grey75",
                 linewidth = 1.2) +
    geom_point(aes(x = `Raw Data`,
                   colour = "Raw Data"),
               size = 2.5) +
    geom_point(aes(x = `Manual Annotation`,
                   colour = "Manual Annotation"),
               size = 2.5) +
    scale_colour_manual(values = c("#0072B2", "#BC3754")) +
    labs(x = "Concordance (%)",
         y = NULL,
         colour = NULL,
         tag = "a") +
    theme_classic() + coord_flip() +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm"))
```

# UV and Batch effect

## Sampled PBMC

```{r}

# UV and Batch Effect

# Two Sample PBMC

DefaultAssay(pbmc) <- "RNA"

pbmc_batch_twoSample <- pbmc
pbmc_batch_twoSample$seurat_clusters <- NULL

# Creating artificial batches
set.seed(121)
pbmc_cells <- colnames(pbmc_batch_twoSample)
pbmc_batch_twoSample_assign <- sample(c("Sample1","Sample2"), size = length(pbmc_cells), replace = TRUE)
pbmc_batch_twoSample$sample <- pbmc_batch_twoSample_assign

table(pbmc_batch_twoSample$sample, pbmc_batch_twoSample$manual_annot)

pbmc_batch_twoSample.list <- SplitObject(pbmc_batch_twoSample, split.by = "sample")

#_____________________

pbmc_batch_twoSample_ClustoCell.list <- lapply(names(pbmc_batch_twoSample.list), function(i) {
  curr_ClustoCell <- clustoCell(data = pbmc_batch_twoSample.list[[i]], high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)
  pbmc_batch_twoSample.list[[i]] <<- addClustoData(obj = pbmc_batch_twoSample.list[[i]], 
                                                  clustoCell = curr_ClustoCell, 
                                                  major_cluster_name = paste0("ClustoCell", "_", i), 
                                                  sub_cluster_name = paste0("SubClustoCell", "_", i))
  curr_ClustoCell
})

names(pbmc_batch_twoSample_ClustoCell.list) <- names(pbmc_batch_twoSample.list)

#__________________________

pbmc_ClustoCell_Sample1vs_manual_annot <- table(Raw = pbmc_batch_twoSample.list[["Sample1"]]$ClustoCell_Sample1, Norm = pbmc_batch_twoSample.list[["Sample1"]]$major_manual_cellTypes)
pbmc_ClustoCell_Sample1vs_manual_annot_maxOverlap <- round(((pbmc_ClustoCell_Sample1vs_manual_annot %>% apply(1, max) %>% sum())/ncol(pbmc_batch_twoSample.list[["Sample1"]]))*100)


pbmc_ClustoCell_Sample1vs_manual_annot <- as.data.frame(pbmc_ClustoCell_Sample1vs_manual_annot)
colnames(pbmc_ClustoCell_Sample1vs_manual_annot) <- c("Sample1", "Manual Annotation", "Count")

pbmc_ClustoCell_Sample1vs_manual_annot_plt <- ggplot(pbmc_ClustoCell_Sample1vs_manual_annot, 
                                                    aes(x = `Manual Annotation`, 
                                                        y = `Sample1`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_Sample1vs_manual_annot$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_Sample1vs_manual_annot_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "b") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))

#__________________________

pbmc_ClustoCell_Sample1vs_complete_data <- table(Raw = pbmc_batch_twoSample.list[["Sample1"]]$ClustoCell_Sample1, Norm = pbmc_batch_twoSample.list[["Sample1"]]$ClustoCell)
pbmc_ClustoCell_Sample1vs_complete_data_maxOverlap <- round(((pbmc_ClustoCell_Sample1vs_complete_data %>% apply(1, max) %>% sum())/ncol(pbmc_batch_twoSample.list[["Sample1"]]))*100)


pbmc_ClustoCell_Sample1vs_complete_data <- as.data.frame(pbmc_ClustoCell_Sample1vs_complete_data)
colnames(pbmc_ClustoCell_Sample1vs_complete_data) <- c("Sample1", "Complete Data", "Count")

pbmc_ClustoCell_Sample1vs_complete_data_plt <- ggplot(pbmc_ClustoCell_Sample1vs_complete_data, 
                                                     aes(x = `Complete Data`, 
                                                         y = `Sample1`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_Sample1vs_complete_data$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_Sample1vs_complete_data_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = NULL) + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                   barwidth = unit(0.2, "cm")))

#__________________________

pbmc_ClustoCell_Sample2vs_manual_annot <- table(Raw = pbmc_batch_twoSample.list[["Sample2"]]$ClustoCell_Sample2, Norm = pbmc_batch_twoSample.list[["Sample2"]]$major_manual_cellTypes)
pbmc_ClustoCell_Sample2vs_manual_annot_maxOverlap <- round(((pbmc_ClustoCell_Sample2vs_manual_annot %>% apply(1, max) %>% sum())/ncol(pbmc_batch_twoSample.list[["Sample2"]]))*100)


pbmc_ClustoCell_Sample2vs_manual_annot <- as.data.frame(pbmc_ClustoCell_Sample2vs_manual_annot)
colnames(pbmc_ClustoCell_Sample2vs_manual_annot) <- c("Sample2", "Manual Annotation", "Count")

pbmc_ClustoCell_Sample2vs_manual_annot_plt <- ggplot(pbmc_ClustoCell_Sample2vs_manual_annot, 
                                                    aes(x = `Manual Annotation`, 
                                                        y = `Sample2`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_Sample2vs_manual_annot$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_Sample2vs_manual_annot_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = "c") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))

#__________________________

pbmc_ClustoCell_Sample2vs_complete_data <- table(Raw = pbmc_batch_twoSample.list[["Sample2"]]$ClustoCell_Sample2, Norm = pbmc_batch_twoSample.list[["Sample2"]]$ClustoCell)
pbmc_ClustoCell_Sample2vs_complete_data_maxOverlap <- round(((pbmc_ClustoCell_Sample2vs_complete_data %>% apply(1, max) %>% sum())/ncol(pbmc_batch_twoSample.list[["Sample2"]]))*100)


pbmc_ClustoCell_Sample2vs_complete_data <- as.data.frame(pbmc_ClustoCell_Sample2vs_complete_data)
colnames(pbmc_ClustoCell_Sample2vs_complete_data) <- c("Sample2", "Complete Data", "Count")

pbmc_ClustoCell_Sample2vs_complete_data_plt <- ggplot(pbmc_ClustoCell_Sample2vs_complete_data, 
                                                     aes(x = `Complete Data`, 
                                                         y = `Sample2`,  color = Count)) +
  geom_point(size = log10(pbmc_ClustoCell_Sample2vs_complete_data$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         pbmc_ClustoCell_Sample2vs_complete_data_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(text = element_text(size = 7),
        legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
        legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
        legend.box.spacing = unit(0, "pt"),
        title = element_text(size = 7),
        axis.title = element_text(size = 7),
        plot.tag = element_text(size = 7, face = "bold"),
        legend.title = element_text(size = 5.5, margin = margin(b = 2)),
        legend.key.width = unit(0.25, "cm"),
        legend.key.height = unit(0.5, "cm")) +
  labs(tag = NULL) + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                   barwidth = unit(0.2, "cm")))

combined_twoSample_pbmc_plt <- 
  wrap_plots(
  wrap_plots(pbmc_ClustoCell_Sample1vs_manual_annot_plt, 
             pbmc_ClustoCell_Sample1vs_complete_data_plt, 
             axes = 'collect_y', guides = "collect"),
  
  wrap_plots(pbmc_ClustoCell_Sample2vs_manual_annot_plt,
             pbmc_ClustoCell_Sample2vs_complete_data_plt,
             axes = 'collect_y', guides = "collect"),
  guides = "keep", widths = c(18, 18)
) + plot_annotation(theme = theme(plot.title = element_text(size = 9)))


saveRDS(object = combined_twoSample_pbmc_plt, file = "../../../Testing results/BatchEffect/combined_twoSample_pbmc_plt.rds")

ggsave(filename = "../../../Testing results/BatchEffect/combined_twoSample_pbmc_plt.png", 
       plot = combined_twoSample_pbmc_plt, device = "png", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/BatchEffect/combined_twoSample_pbmc_plt.pdf", 
       plot = combined_twoSample_pbmc_plt, device = "pdf", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)



```

## pbmc3k and pbmc10k combined

```{r}

# Preparing the data

## PBMC3K

pbmc <- SeuratData::LoadData("pbmc3k")
pbmc$manual_annot <- pbmc$seurat_annotations

pbmc@meta.data <- pbmc@meta.data %>% 
  dplyr::mutate(major_manual_cellTypes = case_when(
    as.character(manual_annot) %in% c("Naive CD4 T", "Memory CD4 T") ~ "CD4 T",
    as.character(manual_annot) %in% c("B") ~ "B",
    as.character(manual_annot) %in% c("CD8 T", "NK") ~ "CL",
    as.character(manual_annot) %in% c("CD14+ Mono", "FCGR3A+ Mono", "DC") ~ "MNP",
    as.character(manual_annot) %in% c("Platelet") ~ "PLT",
    .default = NA
  ))

pbmc$major_manual_cellTypes <- factor(pbmc$major_manual_cellTypes, levels = c("CD4 T", "B", "CL", "MNP", "PLT"))

# Normalizing the data
pbmc <- NormalizeData(pbmc, normalization.method = "LogNormalize", scale.factor = 10000)

# Identification of highly variable features (feature selection)
pbmc <- FindVariableFeatures(pbmc, selection.method = "vst", nfeatures = 2000)

# Scaling the data
pbmc <- ScaleData(pbmc, features = VariableFeatures(pbmc))

# Perform linear dimensional reduction
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))

# Elbow plot
ElbowPlot(pbmc)

# Cluster the cells
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)

# Run non-linear dimensional reduction (UMAP/tSNE)
pbmc <- RunUMAP(pbmc, dims = 1:10)

# Note that you can set `label = TRUE` or use the LabelClusters function to help label
# individual clusters
DimPlot(pbmc, reduction = "umap")

pbmc_DimPlot <- CellDimPlot(pbmc, reduction = "umap", label = F, pt.size = 0.2, theme = "theme_blank") + ggtitle("KNN")
pbmc_DimPlot

pbmc_DimPlot_mannual_annot <- DimPlot(pbmc, reduction = "umap", group.by = "manual_annot", label = F, pt.size = 0.2) + ggtitle("Manual Annotation")
pbmc_DimPlot_mannual_annot

# Find the clusters via ClustoCell

pbmc_ClustoCell <- clustoCell(data = pbmc, high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, seed = 121)
pbmc <- addClustoData(obj = pbmc, clustoCell = pbmc_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

#________________________________________________________________________
#________________________________________________________________________

## PBMC10k

# Install the dataset
# download.file(url = "https://seurat.nygenome.org/src/contrib/pbmcMultiome.SeuratData_0.1.4.tar.gz",
#               destfile = "Benchmarking/Test Datasets/pbmcMultiome/pbmcMultiome.SeuratData_0.1.4.tar.gz")
# 
# install.packages("Benchmarking/Test Datasets/pbmcMultiome/pbmcMultiome.SeuratData_0.1.4.tar.gz", 
#                  repos = NULL, type = "source")

# Load the pbmcMultiome dataset

# library(SeuratData)
# library(pbmcMultiome.SeuratData)
pbmc10k <- pbmcMultiome.SeuratData::pbmc.rna

pbmc10k <- subset(pbmc10k, seurat_annotations != "filtered")
pbmc10k$manual_annot <- pbmc10k$seurat_annotations

# Normalizing the data
pbmc10k <- NormalizeData(pbmc10k, normalization.method = "LogNormalize", scale.factor = 10000)

# Identification of highly variable features (feature selection)
pbmc10k <- FindVariableFeatures(pbmc10k, selection.method = "vst", nfeatures = 2000)

# Scaling the data
pbmc10k <- ScaleData(pbmc10k, features = VariableFeatures(pbmc10k))

# Perform linear dimensional reduction
pbmc10k <- RunPCA(pbmc10k, features = VariableFeatures(object = pbmc10k))

# Elbow plot
ElbowPlot(pbmc10k)

# Cluster the cells
pbmc10k <- FindNeighbors(pbmc10k, dims = 1:12)
pbmc10k <- FindClusters(pbmc10k, resolution = 0.5)

# Run non-linear dimensional reduction (UMAP/tSNE)
pbmc10k <- RunUMAP(pbmc10k, dims = 1:12)

# Note that you can set `label = TRUE` or use the LabelClusters function to help label
# individual clusters
DimPlot(pbmc10k, reduction = "umap")

pbmc10k_DimPlot <- CellDimPlot(pbmc10k, reduction = "umap", label = F, pt.size = 0.2, theme = "theme_blank") + ggtitle("KNN")
pbmc10k_DimPlot

pbmc10k_DimPlot_major_mannual_annot <- CellDimPlot(pbmc10k, reduction = "umap", group_by = "mannual_annot",
                                             theme = "theme_blank", label = F, pt.size = 0.2) + 
  ggtitle("Manual Annotation")
pbmc10k_DimPlot_major_mannual_annot

pbmc10k_DimPlot_mannual_annot <- CellDimPlot(pbmc10k, reduction = "umap", group_by = "manual_annot",
                                                   theme = "theme_blank", label = F, pt.size = 0.2) + 
  ggtitle("Manual Annotation")
pbmc10k_DimPlot_mannual_annot

# Find the clusters via ClustoCell

pbmc10k_ClustoCell <- clustoCell(data = pbmc10k, high_quantile = 0.25, low_quantile = 0.25, 
                                 subcluster_resolution_weight = 0.9,
                                 subset_to_HVG = F, seed = 121)
pbmc10k <- addClustoData(obj = pbmc10k, clustoCell = pbmc10k_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

#________________________________________________________________________

pbmc10k$MajorClustoCell <- paste("PBMC10K", pbmc10k$ClustoCell, sep = "-")
pbmc10k$SubTypeClustoCell <- paste("PBMC10K", pbmc10k$SubClustoCell, sep = "-")
pbmc10k$Sample <- "PBMC10K"

pbmc$MajorClustoCell <- paste("PBMC3K", pbmc$ClustoCell, sep = "-")
pbmc$SubTypeClustoCell <- paste("PBMC3K", pbmc$SubClustoCell, sep = "-")
pbmc$Sample <- "PBMC3K"

#________________________________________________________________________
#________________________________________________________________________

two_pbmc_combined <- merge(pbmc, pbmc10k)
two_pbmc_combined <- JoinLayers(two_pbmc_combined)

two_pbmc_combined <- NormalizeData(two_pbmc_combined, normalization.method = "LogNormalize", scale.factor = 10000)
two_pbmc_combined <- FindVariableFeatures(two_pbmc_combined, selection.method = "vst", nfeatures = 2000)
two_pbmc_combined <- ScaleData(two_pbmc_combined, features = VariableFeatures(two_pbmc_combined))
two_pbmc_combined <- RunPCA(two_pbmc_combined, features = VariableFeatures(object = two_pbmc_combined))
ElbowPlot(two_pbmc_combined)
# two_pbmc_combined <- FindNeighbors(two_pbmc_combined, dims = 1:11)
# two_pbmc_combined <- FindClusters(two_pbmc_combined, resolution = 0.5)
two_pbmc_combined <- RunUMAP(two_pbmc_combined, dims = 1:11)

CellDimPlot(two_pbmc_combined, reduction = "umap", group_by = "Sample", theme = "theme_blank", label = T, pt.size = 0.2)

two_pbmc_combined_ClustoCell <- clustoCell(data = two_pbmc_combined, high_quantile = 0.25, low_quantile = 0.25, subset_to_HVG = F, leiden_resolution = 1.5, identify_subclusters = F, seed = 121)
two_pbmc_combined <- addClustoData(obj = two_pbmc_combined, clustoCell = two_pbmc_combined_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")

CellDimPlot(two_pbmc_combined, reduction = "umap", group_by = "ClustoCell", theme = "theme_blank", label = T, pt.size = 0.2) + 
  CellDimPlot(two_pbmc_combined, reduction = "umap", group_by = "MajorClustoCell", theme = "theme_blank", label = T, pt.size = 0.2)

#_______________________

two_pbmc_combined_umap <- CellDimPlot(two_pbmc_combined, reduction = "umap", group_by = c("Sample"), label = F, pt_size = 0.000001, 
                                                   theme = "theme_blank", xlab = "", ylab = "", raster = F,
                                                   legend.position = "right", legend.direction = "vertical", palette = "npg") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 0, 5, 0),
    legend.margin = margin(0, 0, 5, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  ) +
  labs(color = NULL, x = NULL, y = NULL, tag = "d") +
  guides(color = guide_legend(override.aes = list(size = 2))) 


#_________________

### ClustoCell Merged vs Separated
two_pbmc_combined_ClustoCellMerged_vs_split <- table(ClustoCellMerged = two_pbmc_combined$ClustoCell, ClustoCellSplit = two_pbmc_combined$MajorClustoCell)
two_pbmc_combined_ClustoCellMerged_vs_split_maxOverlap <- round(((two_pbmc_combined_ClustoCellMerged_vs_split %>% apply(1, max) %>% sum())/ncol(two_pbmc_combined))*100)

two_pbmc_combined_ClustoCellMerged_vs_split <- as.data.frame(two_pbmc_combined_ClustoCellMerged_vs_split)
colnames(two_pbmc_combined_ClustoCellMerged_vs_split) <- c("Merged Batches", "Separate Batches", "Count")

two_pbmc_combined_ClustoCellMerged_vs_split$`Separate Batches` <- gsub("PBMC", "", two_pbmc_combined_ClustoCellMerged_vs_split$`Separate Batches`)

two_pbmc_combined_ClustoCellMerged_vs_split_cat_df <- two_pbmc_combined_ClustoCellMerged_vs_split |>
    distinct(`Separate Batches`) |>
    mutate(color = case_when(
      grepl(pattern = "10K", `Separate Batches`) ~ plotthis::palette_list$npg[2],
      grepl(pattern = "3K", `Separate Batches`) ~ plotthis::palette_list$npg[1]
    ))

two_pbmc_combined_ClustoCellMerged_vs_split_plt <- ggplot(two_pbmc_combined_ClustoCellMerged_vs_split, 
                                                    aes(x = `Merged Batches`, 
                                                        y = `Separate Batches`, color = Count)) +
  geom_point(size = log10(two_pbmc_combined_ClustoCellMerged_vs_split$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(x = "Merged Batches", y = "Separate Batches", title = NULL,
       subtitle = paste0("Concordance: ",
                         two_pbmc_combined_ClustoCellMerged_vs_split_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(
    text = element_text(size = 7),
    title = element_text(size = 7),
    axis.title = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    legend.title = element_text(size = 5.5, margin = margin(b = 2)),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
    legend.box.spacing = unit(0, "pt"),
    legend.key.width = unit(0.25, "cm"),
    legend.key.height = unit(0.5, "cm")) +
  labs(tag = "e") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                   barwidth = unit(0.2, "cm"))) +
  geom_tile(
    data = two_pbmc_combined_ClustoCellMerged_vs_split_cat_df,
    aes(y = `Separate Batches`, x = -0.5, fill = color),
    inherit.aes = FALSE, 
    width = 0.7,
    height = 0.6
) +
    
    scale_fill_identity(guide = "none") +                    # use literal colors for strip
    scale_y_discrete(expand = expansion(add = 0.6)) +  # leave space at bottom
    coord_cartesian(clip = "off")  +      # prevent clipping of strip
    # coord_flip() +
    labs(tag = "e") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                    barwidth = unit(0.2, "cm"))) +
    theme(
        plot.margin   = margin(b = 0)
        # axis.title.x  = element_text(margin = margin(t = 0, b = 0)),
        # axis.text.x   = element_text(margin = margin(t = 0, b = 0))
    )


two_pbmcMerged_combined_plt <- 
    wrap_plots(
      two_pbmc_combined_umap,
      two_pbmc_combined_ClustoCellMerged_vs_split_plt,
      guides = "keep", widths = c(17, 19)
               ) + plot_annotation(theme = theme(plot.title = element_text(size = 9)))


saveRDS(object = two_pbmcMerged_combined_plt, file = "../../../Testing results/BatchEffect/two_pbmcMerged_combined_plt.rds")

ggsave(filename = "../../../Testing results/BatchEffect/two_pbmcMerged_combined_plt.png", 
       plot = two_pbmcMerged_combined_plt, device = "png", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/BatchEffect/two_pbmcMerged_combined_plt.pdf", 
       plot = two_pbmcMerged_combined_plt, device = "pdf", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)


```

## Splatter (Global + gene subset)

```{r}

# Splatter (Global + gene subset)

# Only global batch effect →
# batch.facLoc > 0, batch.facScale = 0.
# 
# Only gene-specific effects (no global mean difference) →
# batch.facLoc = 0, batch.facScale > 0.
# 
# Both →
# batch.facLoc > 0, batch.facScale > 0.

library(splatter)
library(SingleCellExperiment)

# Parameters: 2 batches with gene-specific effects
set.seed(1234)
splatter_globalPlusSubset_params <- newSplatParams(batchCells = c(500, 800), nGenes = 2000,   
                                                   group.prob = c(0.1, 0.2, 0.3, 0.05, 0.35),
                                                   de.prob = 0.6,
                                                   de.facLoc = 0.6,
                                                   batch.facLoc = 0.3, # average log-fold-change across batches
                                                   batch.facScale = 0.3, # variability of gene-specific shifts
                                                   seed = 1234
                                                   ) 

splatter_globalPlusSubset_sim <- splatSimulate(params = splatter_globalPlusSubset_params, 
                                               method = "groups", batch.rmEffect = FALSE, verbose = FALSE) %>% 
  scuttle::logNormCounts() %>% 
  Seurat::as.Seurat() %>% 
  NormalizeData() %>% 
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000/5)

splatter_globalPlusSubset_sim <- ScaleData(splatter_globalPlusSubset_sim, features = VariableFeatures(splatter_globalPlusSubset_sim))
splatter_globalPlusSubset_sim <- RunPCA(splatter_globalPlusSubset_sim, features = VariableFeatures(object = splatter_globalPlusSubset_sim))
ElbowPlot(splatter_globalPlusSubset_sim)
splatter_globalPlusSubset_sim <- FindNeighbors(splatter_globalPlusSubset_sim, dims = 1:8)
splatter_globalPlusSubset_sim <- FindClusters(splatter_globalPlusSubset_sim, resolution = 0.001)
splatter_globalPlusSubset_sim <- RunUMAP(splatter_globalPlusSubset_sim, dims = 1:8)
DimPlot(splatter_globalPlusSubset_sim , reduction = "pca", group.by = c("Group", "Batch", "seurat_clusters"))

splatter_globalPlusSubset_sim_ClustoCell <- clustoCell(data = splatter_globalPlusSubset_sim, assay = "originalexp", leiden_resolution = 0.6, seed = 121)
splatter_globalPlusSubset_sim <- addClustoData(obj = splatter_globalPlusSubset_sim, clustoCell = splatter_globalPlusSubset_sim_ClustoCell, major_cluster_name = "ClustoCell", sub_cluster_name = "SubClustoCell")


#________

splatter_globalPlusSubset_major_pca <- CellDimPlot(splatter_globalPlusSubset_sim, reduction = "pca", group_by = c("Group"), label = F, pt_size = 0.000001, facet_by = "Batch", 
                                  theme = "theme_blank", xlab = "", ylab = "", raster = F,
                                  legend.position = "right", legend.direction = "vertical", palette = "npg") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(0, 0, 5, 0),
    legend.margin = margin(0, 0, 5, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt")
  ) +
  scale_color_manual(values = npg_reverse_pal) +
  labs(color = NULL, x = NULL, y = NULL, tag = "f") +
  guides(color = guide_legend(override.aes = list(size = 2))) 

#________

splatter_globalPlusSubset_sim_ClustoCell_vs_groups <- table(
                                                            Group = as.character(splatter_globalPlusSubset_sim$Group), ClustoCell = as.character(splatter_globalPlusSubset_sim$ClustoCell))

splatter_globalPlusSubset_sim_ClustoCell_vs_groups_maxOverlap <- round(((splatter_globalPlusSubset_sim_ClustoCell_vs_groups %>% 
                                                                                 apply(1, max) %>% sum())/ncol(splatter_globalPlusSubset_sim))*100)

splatter_globalPlusSubset_sim_ClustoCell_vs_groups <- as.data.frame(splatter_globalPlusSubset_sim_ClustoCell_vs_groups)
colnames(splatter_globalPlusSubset_sim_ClustoCell_vs_groups) <- c("Group", "ClustoCell", "Count")
splatter_globalPlusSubset_sim_ClustoCell_vs_groups <- splatter_globalPlusSubset_sim_ClustoCell_vs_groups %>%
  mutate(ClustoCell = as.character(ClustoCell))

splatter_globalPlusSubset_sim_ClustoCell_vs_groups_plt <- ggplot(splatter_globalPlusSubset_sim_ClustoCell_vs_groups, 
                                                                       aes(x = ClustoCell, 
                                                                           y = Group, color = Count)) +
  geom_point(size = log10(splatter_globalPlusSubset_sim_ClustoCell_vs_groups$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         splatter_globalPlusSubset_sim_ClustoCell_vs_groups_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(
    text = element_text(size = 7),
    title = element_text(size = 7),
    axis.title = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    legend.title = element_text(size = 5.5, margin = margin(b = 2)),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
    legend.box.spacing = unit(0, "pt"),
    legend.key.width = unit(0.25, "cm"),
    legend.key.height = unit(0.5, "cm")) +
  labs(tag = "g") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))

#_____________

splatter_globalPlusSubset_sim_KNN_vs_groups <- table(
                                                    Group = as.character(splatter_globalPlusSubset_sim$Group), KNN = as.character(splatter_globalPlusSubset_sim$seurat_clusters))

splatter_globalPlusSubset_sim_KNN_vs_groups_maxOverlap <- round(((splatter_globalPlusSubset_sim_KNN_vs_groups %>% 
                                                                   apply(1, max) %>% sum())/ncol(splatter_globalPlusSubset_sim))*100)

splatter_globalPlusSubset_sim_KNN_vs_groups <- as.data.frame(splatter_globalPlusSubset_sim_KNN_vs_groups)
colnames(splatter_globalPlusSubset_sim_KNN_vs_groups) <- c("Group", "KNN", "Count")
splatter_globalPlusSubset_sim_KNN_vs_groups <- splatter_globalPlusSubset_sim_KNN_vs_groups %>%
  mutate(KNN = as.character(KNN))

splatter_globalPlusSubset_sim_KNN_vs_groups_plt <- ggplot(splatter_globalPlusSubset_sim_KNN_vs_groups, 
                                                         aes(x = KNN, 
                                                             y = Group, color = Count)) +
  geom_point(size = log10(splatter_globalPlusSubset_sim_KNN_vs_groups$Count)/1.5) +
  # geom_text(aes(label = Count)) +
  scale_color_gradient(low = "white", high = scales::muted("#DC0000FF")) +
  theme_minimal() +
  labs(title = NULL,
       subtitle = paste0("Concordance: ",
                         splatter_globalPlusSubset_sim_KNN_vs_groups_maxOverlap,
                         '%')) + 
  styled_theme(legend_position = "right", style = "classic", title_align = "left", border_mode = "all", legends_to_drop = "size") +
  theme(
    text = element_text(size = 7),
    title = element_text(size = 7),
    axis.title = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    legend.title = element_text(size = 5.5, margin = margin(b = 2)),
    legend.margin = margin(0, 0, 0, 0),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot
    legend.box.spacing = unit(0, "pt"),
    legend.key.width = unit(0.25, "cm"),
    legend.key.height = unit(0.5, "cm")) +
  labs(tag = "h") + guides(color = guide_colorbar(barheight = unit(1.5, "cm"), 
                                                  barwidth = unit(0.2, "cm")))


splatter_globalPlusSubset_combined_plt <- 
      wrap_plots(
        splatter_globalPlusSubset_major_pca,
        splatter_globalPlusSubset_sim_ClustoCell_vs_groups_plt,
        splatter_globalPlusSubset_sim_KNN_vs_groups_plt, 
        guides = "keep", widths = c(18, 7, 9)
               ) + plot_annotation(theme = theme(plot.title = element_text(size = 9)))


saveRDS(object = splatter_globalPlusSubset_combined_plt, file = "../../../Testing results/BatchEffect/splatter_globalPlusSubset_combined_plt.rds")

ggsave(filename = "../../../Testing results/BatchEffect/splatter_globalPlusSubset_combined_plt.png", 
       plot = splatter_globalPlusSubset_combined_plt, device = "png", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/BatchEffect/splatter_globalPlusSubset_combined_plt.pdf", 
       plot = splatter_globalPlusSubset_combined_plt, device = "pdf", width = 6.85, height = (1.762)*1, units = "in", dpi = 300)




```

## Combine the plots

```{r}

# Combine the plots

batch_combine_plt <- 
wrap_plots(combined_twoSample_pbmc_plt,
               two_pbmcMerged_combined_plt,
               splatter_globalPlusSubset_combined_plt, heights = c(1,1.6,1))

ggsave(filename = "../../../Testing results/Batch Effect/batch_combine_plt.png", 
       plot = batch_combine_plt, device = "png", width = 6.85, height = (1.762)*2, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/Batch Effect/batch_combine_plt.pdf", 
       plot = batch_combine_plt, device = "pdf", width = 6.85, height = (1.762)*2, units = "in", dpi = 300)

```

# Combine all norm, sim, batch plots

```{r}

# Combine all norm, sim, batch plots

norm_sim_batch_combined_plt <- wrap_plots(combined_norm_res_plt, batch_combine_plt, patchwork::plot_spacer(), heights = c(0.6, 3, 2)) 

# ggsave(filename = "../../../Testing results/norm_sim_batch_combined_plt.png", 
#        plot = norm_sim_batch_combined_plt, device = "png", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/norm_sim_batch_combined_plt.pdf", 
       plot = norm_sim_batch_combined_plt, device = "pdf", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

```

# Performance of ClustoCell markers for dim reduction in the presence of batch

## Human Pancreatic Islet Cells (Panc8)

## Data preparation

The data was installed using the SeuratData R package.
  
```{r}

# Preparing the Reference dataset

# SeuratData::InstallData("panc8.SeuratData")

# Load the panc8 dataset
panc8_so <- SeuratData::LoadData("panc8")

# Removing low-quality/suspicious gene names
panc8_noisy_features <- c(
  "CXorf49+CXorf49B", "CXorf49+CXorf49B.1", "CXorf51A+CXorf51B", "CXorf51A+CXorf51B.1",
  "DKFZp434J0226", "DKFZp434L192", "DKFZp434P228", "DKFZp451B082", "DKFZp686D0853",
  "DKFZp686K1684", "DKFZp686O1327", "DKFZp779M0652", "pk", "HSPB2-C11orf52",
  "RNASEK-C17orf49", "RPL17-C18orf32", "RPL17-C18orf32.1", "TGIF2-C20orf24", "ZHX1-C8orf76"
)

panc8_so <- subset(panc8_so, features = rownames(panc8_so)[!rownames(panc8_so) %in% panc8_noisy_features])

# We need to remove the Spike-ins as well
panc8_so <- subset(panc8_so, features = rownames(panc8_so)[!grepl("^ERCC-", rownames(panc8_so))])

panc8_so@meta.data$celltype <- stringr::str_to_title(panc8_so@meta.data$celltype)
panc8_so@meta.data$tech <- stringr::str_to_title(panc8_so@meta.data$tech)

panc8_so@meta.data <- panc8_so@meta.data %>% 
  dplyr::mutate(celltype = case_match(panc8_so@meta.data$celltype,
                                      "Activated_stellate" ~ "aStellate",
                                      "Quiescent_stellate" ~ "qStellate",
                                      .default = celltype))

qs2::qs_save(object = panc8_so, file = "../../../Benchmarking/Test Datasets/panc8/panc8_so.qs2")

panc8_so <- qs2::qs_read("../../../Benchmarking/Test Datasets/panc8/panc8_so.qs2")

```

## Based on 2000 HVGs

```{r}

# Dim plot using 2000 HVGs

panc8_so <- NormalizeData(panc8_so, normalization.method = "LogNormalize", scale.factor = 10000)
panc8_so <- FindVariableFeatures(panc8_so, selection.method = "vst", nfeatures = 2000)
panc8_so <- ScaleData(panc8_so, features = VariableFeatures(panc8_so))
panc8_so <- RunPCA(panc8_so, features = VariableFeatures(object = panc8_so))
ElbowPlot(panc8_so, ndims = 50)
panc8_so <- RunUMAP(panc8_so, dims = 1:20)

#__________

## Calculate silhouette score
set.seed(121)
panc8_2000_hvg_pca_silhouette <- bluster::approxSilhouette(Seurat::Embeddings(object = panc8_so, reduction = "pca")[,1:20], 
                                                           clusters = panc8_so$celltype)


panc8_2000_hvg_pca_silhouette <- round(mean(panc8_2000_hvg_pca_silhouette[, "width"]), 3)

## Calculate trustworthiness
panc8_2000_hvg_trust_score <- compute_trust_sklearn(X_high = Seurat::Embeddings(panc8_so, "pca")[, 1:20], 
                                                    X_low = Seurat::Embeddings(panc8_so, "umap"),
                                                    k = 30L, nCells = ncol(panc8_so), seed = 121)

panc8_2000_hvg_umap <- CellDimPlot(panc8_so, reduction = "umap", group_by = "celltype", 
                                   label = F, pt_size = 2, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = -5),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 0, b = 5, l = -5),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "HVGs (n=2,000)", 
       color = NULL, x = NULL, y = NULL, tag = "i") +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2))) 

```

## Based on ClustoCell Markers

```{r}

# Dim plot based on clustoCell markers
options(future.globals.maxSize = 1e9)
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 100)

panc8_annot_markoClust <- markoClust(data = panc8_so, 
                                     cluster_labels = "celltype", identify_subclusters = F,
                                     seed = 121)

qs2::qs_save(object = panc8_annot_markoClust, 
             file = "../../../Benchmarking/Test Datasets/panc8/panc8_annot_markoClust.qs2")

panc8_annot_markoClust <- qs2::qs_read("../../../Benchmarking/Test Datasets/panc8/panc8_annot_markoClust.qs2")

panc8_markers <- getDatasetMarkers(obj = panc8_annot_markoClust, 
                                   positive_markers = T, negative_markers = F, medium_markers = F,
                                   thresh_mode = "n", pos_thresh = 25)

panc8_so <- ScaleData(panc8_so, features = panc8_markers$combined_markers)
panc8_so <- RunPCA(panc8_so, features = panc8_markers$combined_markers, reduction.name = "clusto_pca")
panc8_so <- RunUMAP(panc8_so, dims = 1:20, reduction = "clusto_pca", reduction.name = "clusto_umap")

#___________

## Calculate silhouette score
set.seed(121)
panc8_clustoCellMarkers_pca_silhouette <- bluster::approxSilhouette(Seurat::Embeddings(object = panc8_so, reduction = "clusto_pca")[,1:20], 
                                                                    clusters = panc8_so$celltype)

panc8_clustoCellMarkers_pca_silhouette <- round(mean(panc8_clustoCellMarkers_pca_silhouette[, "width"]), 3)

## Calculate trustworthiness
panc8_clustoCellMarkers_trust_score <- compute_trust_sklearn(X_high = Seurat::Embeddings(panc8_so, "clusto_pca")[, 1:20], 
                                                             X_low = Seurat::Embeddings(panc8_so, "clusto_umap"),
                                                             k = 30L, nCells = ncol(panc8_so), seed = 121)

panc8_clustoCellMarkers_umap <- CellDimPlot(panc8_so, reduction = "clusto_umap", group_by = "celltype", 
                                            label = F, pt_size = 2, 
                                            theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                            legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = -5),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 0, b = 5, l = -5),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = paste0("ClustoCell Markers (n=", 
                         length(panc8_markers$combined_markers), 
                         ")"), 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2)))

```

## Based on FinaAllMarkers Matched with ClustoCell Markers

```{r}

panc8_markers_all <- FindAllMarkers(
  object = panc8_so,
  group.by = "celltype",
  only.pos = TRUE            # only positive markers (upregulated)
)

# Matching the markers with the number of ClustoCell Markers
panc8_markers_selected <- pick_top_markers_equally(panc8_markers_all, 
                                                            total_n = length(panc8_markers$combined_markers), 
                                                            padj_cutoff = 0.05)

#_________________

panc8_so <- ScaleData(panc8_so, features = panc8_markers_selected$gene)
panc8_so <- RunPCA(panc8_so, features = panc8_markers_selected$gene, reduction.name = "matched_findAll_pca")
panc8_so <- RunUMAP(panc8_so, dims = 1:20, reduction = "matched_findAll_pca", reduction.name = "matched_findAll_umap")

## Saving the obj after preprocessing as a qs obj
qs2::qs_save(object = panc8_so, 
             file = "../../../Benchmarking/Test Datasets/panc8/panc8_so.qs2")

#___________

## Calculate silhouette score
set.seed(121)
panc8_matched_findAll_pca_silhouette <- bluster::approxSilhouette(Seurat::Embeddings(object = panc8_so, reduction = "matched_findAll_pca")[,1:20], 
                                                                           clusters = panc8_so$celltype)

panc8_matched_findAll_pca_silhouette <- round(mean(panc8_matched_findAll_pca_silhouette[, "width"]), 3)

## Calculate trustworthiness
panc8_matched_findAll_trust_score <- compute_trust_sklearn(X_high = Seurat::Embeddings(panc8_so, "matched_findAll_pca")[, 1:20], 
                                                                    X_low = Seurat::Embeddings(panc8_so, "matched_findAll_umap"),
                                                                    k = 30L, nCells = ncol(panc8_so), seed = 121)

panc8_matched_findAll_umap <- CellDimPlot(panc8_so, reduction = "matched_findAll_umap", group_by = "celltype", 
                                                   label = F, pt_size = 2, 
                                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = -5),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 0, b = 5, l = -5),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = paste0("FindAllMarkers (n=", 
                         length(panc8_markers$combined_markers), 
                         ")"), 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2)))


```

## Batch plot

```{r}

panc8_2000_hvg_batch_umap <- CellDimPlot(panc8_so, reduction = "umap", group_by = "tech", 
                                   label = F, pt_size = 1.7, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 0, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "HVGs (n=2,000)", 
       color = NULL, x = NULL, y = NULL, tag = "j") +
  guides(color = guide_legend(ncol = 1, bycol = T, override.aes = list(size = 2))) 

#______________________

panc8_clusto_batch_umap <- CellDimPlot(panc8_so, reduction = "clusto_umap", group_by = "tech", 
                                   label = F, pt_size = 1.7, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 0, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "ClustoCell Markers (n=227)", 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 1, bycol = T, override.aes = list(size = 2))) 

#______________________

panc8_findallmarker_batch_umap <- CellDimPlot(panc8_so, reduction = "matched_findAll_umap", group_by = "tech", 
                                   label = F, pt_size = 1.7, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 0, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 0, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-5, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "FindAllMarkers (n=227)", 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 1, bycol = T, override.aes = list(size = 2))) 

```

## Metrics plot

```{r}

panc8_dim_red_metrics_df <- tibble(Mode = c("2000\nHVGs", 
                                                "ClustoCell\nMarkers", 
                                                "N-matched\nFindAllMarkers"),
                                       `Silhouette` = c(panc8_2000_hvg_pca_silhouette,
                                                        panc8_clustoCellMarkers_pca_silhouette,
                                                        panc8_matched_findAll_pca_silhouette),
                                       Trustworthiness = c(panc8_2000_hvg_trust_score,
                                                           panc8_clustoCellMarkers_trust_score,
                                                           panc8_matched_findAll_trust_score)) %>% 
  pivot_longer(cols = c("Silhouette", "Trustworthiness"), 
               names_to = "Test", values_to = "Score")

panc8_dim_red_metrics_df$Mode <- factor(panc8_dim_red_metrics_df$Mode, 
                                            levels = c("2000\nHVGs",
                                                       "N-matched\nFindAllMarkers",
                                                       "ClustoCell\nMarkers"))

#________________

panc8_dim_red_metrics_vis <- 
  ggplot(panc8_dim_red_metrics_df,
         aes(x = Mode, y = Score, fill = Mode)) +
  geom_bar(stat = "identity", width = 0.7) +
  facet_wrap(~ Test, scales = "free_y") +
  scale_y_continuous(labels = fmt_num) +
  scale_fill_manual(values = c("#238A8D", "#6A51A3", "#F06C4E"), name = NULL) +
  labs(
    title = NULL,
    subtitle = NULL,
    x = NULL,
    tag = "j"
  ) +
  theme_minimal(base_size = 7) +
  theme(
    text = element_text(size = 8),
    strip.text = element_text(size = 5),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 5),
    title = element_text(size = 8),
    plot.tag = element_text(size = 7, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "right",
    legend.direction = "vertical",
    legend.margin = margin(0, 0, 5, 0),
    legend.box.margin = margin(0, 0, 0, 5),
    legend.box.spacing = unit(0, "pt"),
    legend.text = element_text(size = 4.5),
    legend.text.align = 0,
    legend.key.width  = unit(7, "pt"),
    legend.key.height  = unit(7, "pt"),
    legend.key.spacing.y = unit(3, "pt"),
    plot.margin = margin(t = 0, r = 0, b = -5, l = 0)
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(colour = NA, size = 4)
    )
  )

```

### Combining the plots of panc8

```{r}

panc8_dimPlots_combined <-
  (
    wrap_elements(panc8_2000_hvg_umap + 
        panc8_matched_findAll_umap +
        panc8_clustoCellMarkers_umap +
        plot_layout(widths = c(10, 10, 10), heights = 1, guides = "collect")) & 
       theme(legend.margin = margin(t = 0, r = 0, b = 0, l = 0))
    ) /
    (
      (wrap_elements(
        panc8_2000_hvg_batch_umap + 
        panc8_findallmarker_batch_umap +
        panc8_clusto_batch_umap +
        plot_layout(widths = c(10, 10, 10), heights = 1, guides = "collect")) & 
       theme(legend.margin = margin(t = 0, r = 0, b = 0, l = 0))) +
    wrap_elements(panc8_dim_red_metrics_vis) + 
  plot_layout(widths = c(52, 17), heights = 1, guides = "keep") & theme(plot.margin = margin(0, 0, 0, 0))
) 


ggsave(filename = "../../../Testing results/panc8/Final/panc8_dimPlots_combined.png", 
       plot = panc8_dimPlots_combined, device = "png", width = 6.85, height = (1.762)*1.5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/panc8/Final/panc8_dimPlots_combined.pdf", 
       plot = panc8_dimPlots_combined, device = "pdf", width = 6.85, height = (1.762)*1.5, units = "in", dpi = 300)

```

## ifnb

## Data preparation

Kang et al., “Multiplexed droplet single-cell RNA-sequencing using natural genetic variation.” Nat. Biotechnol. 2018
https://pmc.ncbi.nlm.nih.gov/articles/PMC5784859/

The data was downloaded from https://seurat.nygenome.org/src/contrib/ifnb.SeuratData_3.0.0.tar.gz

```{r}

# install.packages("https://seurat.nygenome.org/src/contrib/ifnb.SeuratData_3.0.0.tar.gz", repos = NULL, type = "source")
library(ifnb.SeuratData)

# Load the ifnb dataset
ifnb_so <- SeuratData::LoadData("ifnb")
ifnb_so$celltype <- ifnb_so$seurat_annotations

ifnb_so$celltype <- as.character(ifnb_so$celltype)

ifnb_so$celltype[ifnb_so$celltype == "CD4 Memory T"] <- "CD4 Mem"
ifnb_so$celltype[ifnb_so$celltype == "CD4 Naive T"] <- "CD4 Naive"
ifnb_so$celltype[ifnb_so$celltype == "CD4 Memory T"] <- "CD4 Mem"
ifnb_so$celltype[ifnb_so$celltype == "T activated"] <- "T Act."
ifnb_so$celltype[ifnb_so$celltype == "B Activated"] <- "B Act."

ifnb_so$celltype <- factor(ifnb_so$celltype)

ifnb_so@meta.data$stim <- stringr::str_to_title(ifnb_so@meta.data$stim)

qs2::qs_save(object = ifnb_so, file = "../../../Benchmarking/Test Datasets/ifnb/ifnb_so.qs2")

ifnb_so <- qs2::qs_read("../../../Benchmarking/Test Datasets/ifnb/ifnb_so.qs2")

```

## Based on 2000 HVGs

```{r}

# Dim plot using 2000 HVGs

ifnb_so <- NormalizeData(ifnb_so, normalization.method = "LogNormalize", scale.factor = 10000)
ifnb_so <- FindVariableFeatures(ifnb_so, selection.method = "vst", nfeatures = 2000)
ifnb_so <- ScaleData(ifnb_so, features = VariableFeatures(ifnb_so))
ifnb_so <- RunPCA(ifnb_so, features = VariableFeatures(object = ifnb_so))
ElbowPlot(ifnb_so, ndims = 50)
ifnb_so <- RunUMAP(ifnb_so, dims = 1:10)

#__________

## Calculate silhouette score
set.seed(121)
ifnb_2000_hvg_pca_silhouette <- bluster::approxSilhouette(Seurat::Embeddings(object = ifnb_so, reduction = "pca")[,1:10], 
                                                           clusters = ifnb_so$celltype)


ifnb_2000_hvg_pca_silhouette <- round(mean(ifnb_2000_hvg_pca_silhouette[, "width"]), 3)

## Calculate trustworthiness
ifnb_2000_hvg_trust_score <- compute_trust_sklearn(X_high = Seurat::Embeddings(ifnb_so, "pca")[, 1:10], 
                                                    X_low = Seurat::Embeddings(ifnb_so, "umap"),
                                                    k = 30L, nCells = ncol(ifnb_so), seed = 121)

ifnb_2000_hvg_umap <- CellDimPlot(ifnb_so, reduction = "umap", group_by = "celltype", 
                                   label = F, pt_size = 2, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 2, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 2, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-6, "pt"),
    legend.key.spacing.x = unit(4, "pt")
  ) +
  labs(subtitle = "HVGs (n=2,000)", 
       color = NULL, x = NULL, y = NULL, tag = "k") +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2))) 

```

## Based on ClustoCell Markers

```{r}

# Dim plot based on clustoCell markers
options(future.globals.maxSize = 1e9)
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 100)

ifnb_annot_markoClust <- markoClust(data = ifnb_so, 
                                     cluster_labels = "celltype", identify_subclusters = F,
                                     seed = 121)

qs2::qs_save(object = ifnb_annot_markoClust, 
             file = "../../../Benchmarking/Test Datasets/ifnb/ifnb_annot_markoClust.qs2")

ifnb_annot_markoClust <- qs2::qs_read("../../../Benchmarking/Test Datasets/ifnb/ifnb_annot_markoClust.qs2")

ifnb_markers <- getDatasetMarkers(obj = ifnb_annot_markoClust, 
                                   positive_markers = T, negative_markers = F, medium_markers = F,
                                   thresh_mode = "n", pos_thresh = 25)

ifnb_so <- ScaleData(ifnb_so, features = ifnb_markers$combined_markers)
ifnb_so <- RunPCA(ifnb_so, features = ifnb_markers$combined_markers, reduction.name = "clusto_pca")
ifnb_so <- RunUMAP(ifnb_so, dims = 1:10, reduction = "clusto_pca", reduction.name = "clusto_umap")

#___________

## Calculate silhouette score
set.seed(121)
ifnb_clustoCellMarkers_pca_silhouette <- bluster::approxSilhouette(Seurat::Embeddings(object = ifnb_so, reduction = "clusto_pca")[,1:10], 
                                                                    clusters = ifnb_so$celltype)

ifnb_clustoCellMarkers_pca_silhouette <- round(mean(ifnb_clustoCellMarkers_pca_silhouette[, "width"]), 3)

## Calculate trustworthiness
ifnb_clustoCellMarkers_trust_score <- compute_trust_sklearn(X_high = Seurat::Embeddings(ifnb_so, "clusto_pca")[, 1:10], 
                                                             X_low = Seurat::Embeddings(ifnb_so, "clusto_umap"),
                                                             k = 30L, nCells = ncol(ifnb_so), seed = 121)

ifnb_clustoCellMarkers_umap <- CellDimPlot(ifnb_so, reduction = "clusto_umap", group_by = "celltype", 
                                            label = F, pt_size = 2, 
                                            theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                            legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 2, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 2, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-6, "pt"),
    legend.key.spacing.x = unit(4, "pt")
  ) +
  labs(subtitle = paste0("ClustoCell Markers (n=", 
                         length(ifnb_markers$combined_markers), 
                         ")"), 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2)))

```

## Based on FinaAllMarkers Matched with ClustoCell Markers

```{r}

ifnb_markers_all <- FindAllMarkers(
  object = ifnb_so,
  group.by = "celltype",
  only.pos = TRUE            # only positive markers (upregulated)
)

# Matching the markers with the number of ClustoCell Markers
ifnb_markers_selected <- pick_top_markers_equally(ifnb_markers_all, 
                                                   total_n = length(ifnb_markers$combined_markers), 
                                                   padj_cutoff = 0.05)

#_________________

ifnb_so <- ScaleData(ifnb_so, features = ifnb_markers_selected$gene)
ifnb_so <- RunPCA(ifnb_so, features = ifnb_markers_selected$gene, reduction.name = "matched_findAll_pca")
ifnb_so <- RunUMAP(ifnb_so, dims = 1:10, reduction = "matched_findAll_pca", reduction.name = "matched_findAll_umap")

## Saving the obj after preprocessing as a qs obj
qs2::qs_save(object = ifnb_so, file = "../../../Benchmarking/Test Datasets/ifnb/ifnb_so.qs2")

#___________

## Calculate silhouette score
set.seed(121)
ifnb_matched_findAll_pca_silhouette <- bluster::approxSilhouette(Seurat::Embeddings(object = ifnb_so, reduction = "matched_findAll_pca")[,1:10], 
                                                                  clusters = ifnb_so$celltype)

ifnb_matched_findAll_pca_silhouette <- round(mean(ifnb_matched_findAll_pca_silhouette[, "width"]), 3)

## Calculate trustworthiness
ifnb_matched_findAll_trust_score <- compute_trust_sklearn(X_high = Seurat::Embeddings(ifnb_so, "matched_findAll_pca")[, 1:10], 
                                                           X_low = Seurat::Embeddings(ifnb_so, "matched_findAll_umap"),
                                                           k = 30L, nCells = ncol(ifnb_so), seed = 121)

ifnb_matched_findAll_umap <- CellDimPlot(ifnb_so, reduction = "matched_findAll_umap", group_by = "celltype", 
                                          label = F, pt_size = 2, 
                                          theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                          legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 2, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 2, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-6, "pt"),
    legend.key.spacing.x = unit(4, "pt")
  ) +
  labs(subtitle = paste0("FindAllMarkers (n=", 
                         length(ifnb_markers$combined_markers), 
                         ")"), 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 2, bycol = T, override.aes = list(size = 2)))


```

## Batch plot

```{r}

ifnb_2000_hvg_batch_umap <- CellDimPlot(ifnb_so, reduction = "umap", group_by = "stim", 
                                   label = F, pt_size = 2, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 2, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 2, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-6, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "HVGs (n=2,000)", 
       color = NULL, x = NULL, y = NULL, tag = "k") +
  guides(color = guide_legend(ncol = 1, bycol = T, override.aes = list(size = 2))) 

#______________________

ifnb_clusto_batch_umap <- CellDimPlot(ifnb_so, reduction = "clusto_umap", group_by = "stim", 
                                   label = F, pt_size = 2, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 2, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 2, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-6, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "ClustoCell Markers (n=168)", 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 1, bycol = T, override.aes = list(size = 2))) 

#______________________

ifnb_findAllMarkers_batch_umap <- CellDimPlot(ifnb_so, reduction = "matched_findAll_umap", group_by = "stim", 
                                   label = F, pt_size = 2, 
                                   theme = "theme_blank", xlab = "", ylab = "", raster = T,
                                   legend.position = "right", legend.direction = "vertical", palette = "igv") +
  # Add coord_fixed to remove internal axis padding (no expansion around UMAP points)
  coord_fixed(clip = "off", expand = FALSE) +  
  theme(
    legend.title = element_text(face = "italic", size = 9),
    text = element_text(size = 8),
    strip.text = element_text(size = 7),
    plot.tag = element_text(size = 7, face = "bold"),
    title = element_text(size = 8),
    axis.title = element_blank(),
    # NEW: Remove outer white space around entire plot (top, right, bottom, left)
    plot.margin = margin(t = 0, r = 2, b = 5, l = 2),
    plot.subtitle = element_text(size = 6.5),
    legend.margin = margin(t = 0, r = 2, b = 5, l = 2),      # Internal legend padding
    legend.box.margin = margin(0, 0, 0, 5),  # External space to plot (keep slight right for legend fit)
    legend.box.spacing = unit(0, "pt"),
    legend.key.size = unit(0.5, "cm"),
    legend.text = element_text(size = 4.5),
    legend.key.width = unit(0, "pt"),  # Minimize space between point and text
    legend.text.align = 0,
    legend.key.spacing.y = unit(-6, "pt"),
    legend.key.spacing.x = unit(-4, "pt")
  ) +
  labs(subtitle = "FindAllMarkers (n=168)", 
       color = NULL, x = NULL, y = NULL, tag = NULL) +
  guides(color = guide_legend(ncol = 1, bycol = T, override.aes = list(size = 2))) 

```

## Metrics plot

```{r}

ifnb_dim_red_metrics_df <- tibble(Mode = c("2000\nHVGs", 
                                                "ClustoCell\nMarkers", 
                                                "N-matched\nFindAllMarkers"),
                                       `Silhouette` = c(ifnb_2000_hvg_pca_silhouette,
                                                        ifnb_clustoCellMarkers_pca_silhouette,
                                                        ifnb_matched_findAll_pca_silhouette),
                                       Trustworthiness = c(ifnb_2000_hvg_trust_score,
                                                           ifnb_clustoCellMarkers_trust_score,
                                                           ifnb_matched_findAll_trust_score)) %>% 
  pivot_longer(cols = c("Silhouette", "Trustworthiness"), 
               names_to = "Test", values_to = "Score")

ifnb_dim_red_metrics_df$Mode <- factor(ifnb_dim_red_metrics_df$Mode, 
                                            levels = c("2000\nHVGs",
                                                       "N-matched\nFindAllMarkers",
                                                       "ClustoCell\nMarkers"))

#________________

ifnb_dim_red_metrics_vis <- 
  ggplot(ifnb_dim_red_metrics_df,
         aes(x = Mode, y = Score, fill = Mode)) +
  geom_bar(stat = "identity", width = 0.7) +
  facet_wrap(~ Test, scales = "free_y") +
  scale_y_continuous(labels = fmt_num) +
  scale_fill_manual(values = c("#238A8D", "#6A51A3", "#F06C4E"), name = NULL) +
  labs(
    title = NULL,
    subtitle = NULL,
    x = NULL,
    tag = "l"
  ) +
  theme_minimal(base_size = 7) +
  theme(
    text = element_text(size = 8),
    strip.text = element_text(size = 5),
    axis.title = element_blank(),
    axis.text.y = element_text(size = 5),
    title = element_text(size = 8),
    plot.tag = element_text(size = 7, face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "right",
    legend.direction = "vertical",
    legend.margin = margin(0, 0, 5, 0),
    legend.box.margin = margin(0, 0, 0, 5),
    legend.box.spacing = unit(0, "pt"),
    legend.text = element_text(size = 4.5),
    legend.text.align = 0,
    legend.key.width  = unit(7, "pt"),
    legend.key.height  = unit(7, "pt"),
    legend.key.spacing.y = unit(3, "pt"),
    plot.margin = margin(t = 0, r = 0, b = -5, l = 0)
  ) +
  guides(
    fill = guide_legend(
      override.aes = list(colour = NA, size = 4)
    )
  )

```

### Cobmining infb plots

```{r}


# ifnb_dimPlots_combined <-
#     (
#     wrap_elements(ifnb_2000_hvg_umap + 
#         ifnb_matched_findAll_umap +
#         ifnb_clustoCellMarkers_umap +
#         plot_layout(widths = c(10, 10, 10), heights = 1, guides = "collect")) & 
#        theme(legend.margin = margin(t = 0, r = 0, b = 0, l = 0))
#     ) /
#     (
#       (wrap_elements(
#         ifnb_2000_hvg_batch_umap + 
#         ifnb_findAllMarkers_batch_umap +
#         ifnb_clusto_batch_umap +
#         plot_layout(widths = c(10, 10, 10), heights = 1, guides = "collect")) & 
#        theme(legend.margin = margin(t = 0, r = 0, b = 0, l = 0))) +
#     wrap_elements(ifnb_dim_red_metrics_vis) + 
#   plot_layout(widths = c(52, 17), heights = 1, guides = "keep") & theme(plot.margin = margin(0, 0, 0, 0))
# ) 

```


## Combining the plots

```{r}

combined_plts <-
combined_norm_res_plt /

(pbmc_ClustoCell_Sample1vs_manual_annot_plt + pbmc_ClustoCell_Sample2vs_manual_annot_plt) /
    
(two_pbmc_combined_umap + two_pbmc_combined_ClustoCellMerged_vs_split_plt) /
    
(splatter_globalPlusSubset_major_pca + splatter_globalPlusSubset_sim_ClustoCell_vs_groups_plt + splatter_globalPlusSubset_sim_KNN_vs_groups_plt) /
    
    (panc8_2000_hvg_umap + panc8_dim_red_metrics_vis) /
    (ifnb_2000_hvg_umap + ifnb_dim_red_metrics_vis)

ggsave(filename = "../../../Testing results/combined_plts.png", 
       plot = combined_plts, device = "png", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

ggsave(filename = "../../../Testing results/combined_plts.pdf", 
       plot = combined_plts, device = "pdf", width = 6.85, height = (1.762)*5, units = "in", dpi = 300)

```
